### 2018-04-25 doc/paper.tex
Add skeleton of project.

\documentclass[letterpaper,10pt]{article}

\usepackage[margin=1in]{geometry}

\usepackage{amsthm,amssymb,amsmath}

\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  citecolor=ForestGreen,
  pdfinfo={
    CreationDate={D:20180425160342},
    ModDate={D:20180425160342},
  },
}

\usepackage{embedfile}
\embedfile{\jobname.tex}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\(K\)-compensated de Casteljau}
\rhead{Danny Hermes}

\renewcommand{\headrulewidth}{0pt}
\newcommand{\fl}{\operatorname{fl}}
\newcommand{\eps}{\varepsilon}

\begin{document}

\tableofcontents

\section{de Casteljau's Method}

Consider de Casteljau's method to evaluate a degree \(n\)
polynomial in Bernstein-B\'{e}zier form with control points \(p_j\):
\begin{align*}
    b_j^{(0)} &= p_j \\
    b_j^{(k)} &= (1 - x) b_j^{(k - 1)} + x b_{j + 1}^{(k - 1)} \\
    b(x) &= b_0^{(n)}.
\end{align*}

\end{document}

### 2018-04-25 nox.py

@@ -1,0 +1,75 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""This is a configuration file for running ``nox`` on this project.
+
+To determine the supported actions run ``nox --list-sessions`` from the
+project root.
+"""
+
+import os
+
+import nox
+import py.path
+
+
+NOX_DIR = os.path.abspath(os.path.dirname(__file__))
+
+
+def get_path(*names):
+    return os.path.join(NOX_DIR, *names)
+
+
+class Remove(object):
+
+    def __init__(self, prefix, extensions):
+        self.prefix = prefix
+        self.extensions = extensions
+
+    def __call__(self):
+        for extension in self.extensions:
+            path = "{}.{}".format(self.prefix, extension)
+            os.remove(path)
+
+
+def build_tex_file(session, base, new_id, extensions=()):
+    # NOTE: This assumes that ``session.chdir(get_path('doc'))``
+    #       has been called.
+    modify_id = get_path("scripts", "modify_pdf_id.py")
+
+    # Run ``pdflatex`` thrice.
+    session.run("pdflatex", base)
+    session.run("pdflatex", base)
+    session.run("pdflatex", base)
+
+    path = get_path("doc", base)
+    remove = Remove(path, extensions)
+    session.run(remove)
+    session.run("python", modify_id, "--base", path, "--id", new_id)
+
+
+@nox.session
+def build_tex(session):
+    if py.path.local.sysfind("pdflatex") is None:
+        session.skip("`pdflatex` must be installed")
+
+    # No need to create a virtualenv.
+    session.virtualenv = False
+
+    session.chdir(get_path("doc"))
+
+    build_tex_file(
+        session,
+        "paper",
+        "F092359D979FDC08931DA1922F3E123E",
+        extensions=("aux", "log", "out", "toc"),
+    )

### 2018-04-25 scripts/modify_pdf_id.py

@@ -1,0 +1,65 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Modify the ``/ID [<...> <...>]`` line in a PDF file.
+
+This is provided to make generated manuscripts (via ``pdflatex``) be
+bitwise identical across runs.
+"""
+
+from __future__ import print_function
+
+import argparse
+
+
+SPLIT_TEXT = b"\n/ID [<"
+
+
+def do_replace(path, new_id):
+    with open(path, "rb") as file_obj:
+        contents = file_obj.read()
+
+    # Assert that there is exactly one match.
+    pre, post = contents.split(SPLIT_TEXT)
+    # ID line expected to be of the form:
+    #     /ID [<...> <...>]
+    _, post = post.split(b"\n", 1)
+    # which would just leave `<...> <...>]` after the split.
+    new_id_line = u"<{}> <{}>]\n".format(new_id, new_id)
+    new_id_line = new_id_line.encode("ascii")
+
+    new_contents = pre + new_id_line + post
+    with open(path, "wb") as file_obj:
+        file_obj.write(new_contents)
+
+    print("Updated {}".format(path))
+
+
+def main():
+    description = (
+        "Modify the `/ID` property in a PDF generated by `pdflatex`."
+    )
+    parser = argparse.ArgumentParser(description=description)
+    parser.add_argument(
+        "--base", required=True, help="Base path for PDF file to be modified."
+    )
+    parser.add_argument(
+        "--id", dest="id_", required=True, help="The prescribed ID to add."
+    )
+
+    args = parser.parse_args()
+    filename = "{}.pdf".format(args.base)
+    do_replace(filename, args.id_)
+
+
+if __name__ == "__main__":
+    main()

### 2018-04-25 nox.py
Adding some "source" code.
Also adding the beginnings of a way to count floating point
operations for a given algorithm.

@@ -73,3 +73,13 @@ def build_tex(session):
         "F092359D979FDC08931DA1922F3E123E",
         extensions=("aux", "log", "out", "toc"),
     )
+
+
+@nox.session
+def flop_counts(session):
+    # No need to create a virtualenv.
+    session.virtualenv = False
+
+    env = {"PYTHONPATH": get_path("src")}
+    compute_counts = get_path("scripts", "compute_counts.py")
+    session.run("python", compute_counts, env=env)

### 2018-04-25 scripts/compute_counts.py

@@ -1,0 +1,57 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Run many experiments to confirm the flop counts for algorithms."""
+
+from __future__ import print_function
+
+import eft
+import operation_count
+
+
+def count_add_eft():
+    parent = operation_count.Computation()
+    val1 = operation_count.Float(1.5, parent)
+    val2 = operation_count.Float(0.5 + 0.5 ** 52, parent)
+    sum_, error = eft.add_eft(val1, val2)
+    assert sum_.value == 2.0
+    assert error.value == 0.5 ** 52
+    print("     add_eft(): {} flops".format(parent.count))
+
+
+def count__split():
+    parent = operation_count.Computation()
+    val = operation_count.Float(1.0 + 0.5 ** 27, parent)
+    high, low = eft._split(val)
+    assert high.value == 1.0
+    assert low.value == 0.5 ** 27
+    print("      _split(): {} flops".format(parent.count))
+
+
+def count_multiply_eft():
+    parent = operation_count.Computation()
+    val1 = operation_count.Float(1.0 + 0.5 ** 40, parent)
+    val2 = operation_count.Float(1.0 - 0.5 ** 40, parent)
+    product, error = eft.multiply_eft(val1, val2)
+    assert product.value == 1.0
+    assert error.value == -0.5 ** 80
+    print("multiply_eft(): {} flops".format(parent.count))
+
+
+def main():
+    count_add_eft()
+    count__split()
+    count_multiply_eft()
+
+
+if __name__ == "__main__":
+    main()

### 2018-04-25 src/eft.py

@@ -1,0 +1,42 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Collection of error-free transforms."""
+
+
+def add_eft(val1, val2):
+    # See: https://doi.org/10.1137/030601818
+    sum_ = val1 + val2
+    delta1 = sum_ - val1
+    error = (val1 - (sum_ - delta1)) + (val2 - delta1)
+    return sum_, error
+
+
+def _split(val):
+    # Helper for ``multiply_eft``.
+    scaled = val * 67108865.0  # 67108865 == 2^{26} + 1.
+    high_bits = scaled - (scaled - val)
+    low_bits = val - high_bits
+    return high_bits, low_bits
+
+
+def multiply_eft(val1, val2):
+    # See: https://doi.org/10.1109/TC.2008.215
+    product = val1 * val2
+    high1, low1 = _split(val1)
+    high2, low2 = _split(val2)
+    error = (
+        low1
+        * low2
+        - (((product - high1 * high2) - low1 * high2) - high1 * low2)
+    )
+    return product, error

### 2018-04-25 src/operation_count.py

@@ -1,0 +1,71 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Helpers for counting flops."""
+
+
+class Computation(object):
+    """Stateful manager of the number of flops."""
+
+    def __init__(self):
+        self.count = 0
+
+
+class Float(object):
+    """A ``float``-like type that will increment a flop count.
+
+    Args:
+        value (float): The current value.
+        computation (.Computation): The current computation
+            in progress.
+    """
+
+    def __init__(self, value, computation):
+        self.value = value
+        self.computation = computation
+
+    def _get_value(self, other):
+        if isinstance(other, Float):
+            if other.computation is not self.computation:
+                raise ValueError(
+                    "Two `Float`s being combined should have the "
+                    "same parent computation."
+                )
+            return other.value
+        elif isinstance(other, float):
+            return other
+        else:
+            return None
+
+    def __add__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.count += 1
+        return Float(self.value + value, self.computation)
+
+    def __sub__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.count += 1
+        return Float(self.value - value, self.computation)
+
+    def __mul__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.count += 1
+        return Float(self.value * value, self.computation)

### 2018-04-25 scripts/compute_counts.py
Adding some breakdown to flop counts.

@@ -25,7 +25,7 @@ def count_add_eft():
     sum_, error = eft.add_eft(val1, val2)
     assert sum_.value == 2.0
     assert error.value == 0.5 ** 52
-    print("     add_eft(): {} flops".format(parent.count))
+    print("     add_eft(): {}".format(parent.display))
 
 
 def count__split():
@@ -34,7 +34,7 @@ def count__split():
     high, low = eft._split(val)
     assert high.value == 1.0
     assert low.value == 0.5 ** 27
-    print("      _split(): {} flops".format(parent.count))
+    print("      _split(): {}".format(parent.display))
 
 
 def count_multiply_eft():
@@ -44,7 +44,7 @@ def count_multiply_eft():
     product, error = eft.multiply_eft(val1, val2)
     assert product.value == 1.0
     assert error.value == -0.5 ** 80
-    print("multiply_eft(): {} flops".format(parent.count))
+    print("multiply_eft(): {}".format(parent.display))
 
 
 def main():

### 2018-04-25 src/operation_count.py

@@ -17,7 +17,19 @@ class Computation(object):
     """Stateful manager of the number of flops."""
 
     def __init__(self):
-        self.count = 0
+        self.add_count = 0
+        self.sub_count = 0
+        self.mul_count = 0
+
+    @property
+    def count(self):
+        return self.add_count + self.sub_count + self.mul_count
+
+    @property
+    def display(self):
+        return "{:2d} flops ({:2d} add, {:2d} sub, {:2d} multiply)".format(
+            self.count, self.add_count, self.sub_count, self.mul_count
+        )
 
 
 class Float(object):
@@ -51,7 +63,7 @@ class Float(object):
         if value is None:
             return NotImplemented
 
-        self.computation.count += 1
+        self.computation.add_count += 1
         return Float(self.value + value, self.computation)
 
     def __sub__(self, other):
@@ -59,7 +71,7 @@ class Float(object):
         if value is None:
             return NotImplemented
 
-        self.computation.count += 1
+        self.computation.sub_count += 1
         return Float(self.value - value, self.computation)
 
     def __mul__(self, other):
@@ -67,5 +79,5 @@ class Float(object):
         if value is None:
             return NotImplemented
 
-        self.computation.count += 1
+        self.computation.mul_count += 1
         return Float(self.value * value, self.computation)

### 2018-04-26 scripts/compute_counts.py
Adding Horner to the operation count script.
Also adding a FMA flag to `multiply_eft` which is
one by default.

@@ -15,6 +15,7 @@
 from __future__ import print_function
 
 import eft
+import horner
 import operation_count
 
 
@@ -38,19 +39,47 @@ def count__split():
 
 
 def count_multiply_eft():
-    parent = operation_count.Computation()
-    val1 = operation_count.Float(1.0 + 0.5 ** 40, parent)
-    val2 = operation_count.Float(1.0 - 0.5 ** 40, parent)
-    product, error = eft.multiply_eft(val1, val2)
-    assert product.value == 1.0
-    assert error.value == -0.5 ** 80
-    print("multiply_eft(): {}".format(parent.display))
+    print("multiply_eft():")
+    for use_fma in (True, False):
+        parent = operation_count.Computation()
+        val1 = operation_count.Float(1.0 + 0.5 ** 40, parent)
+        val2 = operation_count.Float(1.0 - 0.5 ** 40, parent)
+        product, error = eft.multiply_eft(val1, val2, use_fma=use_fma)
+        assert product.value == 1.0
+        assert error.value == -0.5 ** 80
+        if use_fma:
+            description = "with FMA:    "
+        else:
+            description = "w / out FMA: "
+        print("  {} {}".format(description, parent.display))
+
+
+def count_horner_basic():
+    print("horner.basic() (2n):")
+    for degree in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.0, parent)
+        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
+        horner.basic(x, coeffs)
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
+def count_horner_compensated():
+    print("horner.compensated() (11n + 1):")
+    for degree in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.0, parent)
+        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
+        horner.compensated(x, coeffs)
+        print("  degree {}:     {}".format(degree, parent.display))
 
 
 def main():
     count_add_eft()
     count__split()
     count_multiply_eft()
+    count_horner_basic()
+    count_horner_compensated()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/eft.py

@@ -29,14 +29,18 @@ def _split(val):
     return high_bits, low_bits
 
 
-def multiply_eft(val1, val2):
+def multiply_eft(val1, val2, use_fma=True):
     # See: https://doi.org/10.1109/TC.2008.215
     product = val1 * val2
-    high1, low1 = _split(val1)
-    high2, low2 = _split(val2)
-    error = (
-        low1
-        * low2
-        - (((product - high1 * high2) - low1 * high2) - high1 * low2)
-    )
+    if use_fma:
+        error = val1.fma(val1, val2, -product)
+    else:
+        high1, low1 = _split(val1)
+        high2, low2 = _split(val2)
+        error = (
+            low1
+            * low2
+            - (((product - high1 * high2) - low1 * high2) - high1 * low2)
+        )
+
     return product, error

### 2018-04-26 src/horner.py

@@ -1,0 +1,63 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Performs Horner's method.
+
+Horner's method computes
+
+.. math::
+
+    p(x) = a_n x^n + \cdots a_1 x + a_0
+
+via
+
+.. math::
+
+    \begin{align*}
+    p_n &= a_n \\
+    p_k &= p_{k + 1} x + a_k \\
+    p(x) &= p_0
+    \end{align*}
+
+This module provides both the standard version and a compensated version.
+
+.. note::
+
+   This assumes throughout that ``coeffs`` is ordered from
+   :math:`a_n` to :math:`a_0`.
+"""
+
+import eft
+
+
+def basic(x, coeffs):
+    p = coeffs[0]
+    for coeff in coeffs[1:]:
+        p = p * x + coeff
+
+    return p
+
+
+def _compensated(x, coeffs):
+    p = coeffs[0]
+    e = 0.0
+    for coeff in coeffs[1:]:
+        prod, e1 = eft.multiply_eft(p, x)
+        p, e2 = eft.add_eft(prod, coeff)
+        e = x * e + (e1 + e2)
+
+    return p, e
+
+
+def compensated(x, coeffs):
+    p, e = _compensated(x, coeffs)
+    return p + e

### 2018-04-26 src/operation_count.py

@@ -13,6 +13,14 @@
 """Helpers for counting flops."""
 
 
+import fractions
+
+
+_DISPLAY_TEMPLATE = (
+    "{:3d} flops ({:2d} add, {:2d} sub, {:2d} multiply, {:2d} FMA)"
+)
+
+
 class Computation(object):
     """Stateful manager of the number of flops."""
 
@@ -20,15 +28,22 @@ class Computation(object):
         self.add_count = 0
         self.sub_count = 0
         self.mul_count = 0
+        self.fma_count = 0
 
     @property
     def count(self):
-        return self.add_count + self.sub_count + self.mul_count
+        return (
+            self.add_count + self.sub_count + self.mul_count + self.fma_count
+        )
 
     @property
     def display(self):
-        return "{:2d} flops ({:2d} add, {:2d} sub, {:2d} multiply)".format(
-            self.count, self.add_count, self.sub_count, self.mul_count
+        return _DISPLAY_TEMPLATE.format(
+            self.count,
+            self.add_count,
+            self.sub_count,
+            self.mul_count,
+            self.fma_count,
         )
 
 
@@ -74,6 +89,9 @@ class Float(object):
         self.computation.sub_count += 1
         return Float(self.value - value, self.computation)
 
+    def __neg__(self):
+        return Float(-self.value, self.computation)
+
     def __mul__(self, other):
         value = self._get_value(other)
         if value is None:
@@ -81,3 +99,15 @@ class Float(object):
 
         self.computation.mul_count += 1
         return Float(self.value * value, self.computation)
+
+    def fma(self, val1, val2, val3):
+        for val in (val1, val2, val3):
+            if not isinstance(val, Float):
+                raise TypeError("Only `Float` allowed in fma", val)
+
+        self.computation.fma_count += 1
+        frac1 = fractions.Fraction(val1.value)
+        frac2 = fractions.Fraction(val2.value)
+        frac3 = fractions.Fraction(val3.value)
+        result = float(frac1 * frac2 + frac3)
+        return Float(result, self.computation)

### 2018-04-26 scripts/compute_counts.py
Adding start of de Casteljau for op counts.

@@ -14,6 +14,7 @@
 
 from __future__ import print_function
 
+import de_casteljau
 import eft
 import horner
 import operation_count
@@ -74,12 +75,23 @@ def count_horner_compensated():
         print("  degree {}:     {}".format(degree, parent.display))
 
 
+def count_de_casteljau_basic():
+    print("de_casteljau.basic() ((3n^2 + 9n + 8)/2):")
+    for degree in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.0, parent)
+        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
+        de_casteljau.basic(x, coeffs)
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
 def main():
     count_add_eft()
     count__split()
     count_multiply_eft()
     count_horner_basic()
     count_horner_compensated()
+    count_de_casteljau_basic()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/de_casteljau.py

@@ -1,0 +1,53 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Performs de Casteljau's method.
+
+de Casteljau's method evaluates a function in Bernstein-Bezier form
+
+.. math::
+
+    p(s) = p_0 (1 - s)^n + \cdots + p_j \binom{n}{j} (1 - s)^{n - j} s^j +
+        \cdots + p_n s^n
+
+by progressively reducing the control points
+
+.. math::
+
+    \begin{align*}
+    p_j^{(0)} &= p_j \\
+    p_j^{(k + 1)} &= (1 - s) p_j^{(k)} + p_{j + 1}^{(k)} \\
+    p(s) &= p_0^{(n)}.
+    \end{align*}
+
+This module provides both the standard version and a compensated version.
+
+.. note::
+
+   This assumes throughout that ``coeffs`` is ordered from
+   :math:`p_n` to :math:`p_0`.
+"""
+
+
+def basic(s, coeffs):
+    r = 1.0 - s
+
+    degree = len(coeffs) - 1
+    pk = list(coeffs)
+    for d in range(degree, 0, -1):
+        new_pk = []
+        for i in range(d):
+            new_pk.append(r * pk[i] + s * pk[i + 1])
+        # Update the "current" values.
+        pk = new_pk
+
+    return pk[0]

### 2018-04-26 src/operation_count.py

@@ -89,6 +89,14 @@ class Float(object):
         self.computation.sub_count += 1
         return Float(self.value - value, self.computation)
 
+    def __rsub__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.sub_count += 1
+        return Float(value - self.value, self.computation)
+
     def __neg__(self):
         return Float(-self.value, self.computation)
 

### 2018-04-26 scripts/compute_counts.py
Adding some verification for algorithms.

@@ -59,9 +59,10 @@ def count_horner_basic():
     print("horner.basic() (2n):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
-        x = operation_count.Float(0.0, parent)
-        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
-        horner.basic(x, coeffs)
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.basic(x, coeffs)
+        assert p.value == 2.0 ** (degree + 1) - 1
         print("  degree {}:     {}".format(degree, parent.display))
 
 
@@ -69,9 +70,10 @@ def count_horner_compensated():
     print("horner.compensated() (11n + 1):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
-        x = operation_count.Float(0.0, parent)
-        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
-        horner.compensated(x, coeffs)
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.compensated(x, coeffs)
+        assert p.value == 2.0 ** (degree + 1) - 1
         print("  degree {}:     {}".format(degree, parent.display))
 
 
@@ -79,9 +81,13 @@ def count_de_casteljau_basic():
     print("de_casteljau.basic() ((3n^2 + 9n + 8)/2):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
-        x = operation_count.Float(0.0, parent)
-        coeffs = (operation_count.Float(0.0, parent),) * (degree + 1)
-        de_casteljau.basic(x, coeffs)
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = de_casteljau.basic(x, coeffs)
+        assert p.value == 0.5 ** degree
         print("  degree {}:     {}".format(degree, parent.display))
 
 

### 2018-04-26 scripts/compute_counts.py
Fixing flop count for de Casteljau.

@@ -78,7 +78,7 @@ def count_horner_compensated():
 
 
 def count_de_casteljau_basic():
-    print("de_casteljau.basic() ((3n^2 + 9n + 8)/2):")
+    print("de_casteljau.basic() ((3n^2 + 3n + 2)/2):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)

### 2018-04-26 scripts/compute_counts.py
Adding compensated de Casteljau to flop counts.

@@ -20,6 +20,9 @@ import horner
 import operation_count
 
 
+SEPARATOR = "-" * 80
+
+
 def count_add_eft():
     parent = operation_count.Computation()
     val1 = operation_count.Float(1.5, parent)
@@ -91,13 +94,34 @@ def count_de_casteljau_basic():
         print("  degree {}:     {}".format(degree, parent.display))
 
 
+def count_de_casteljau_compensated():
+    print("de_casteljau.compensated() (9n^2 + 9n + 7):")
+    for degree in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = de_casteljau.compensated(x, coeffs)
+        assert p.value == 0.5 ** degree
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
 def main():
     count_add_eft()
+    print(SEPARATOR)
     count__split()
+    print(SEPARATOR)
     count_multiply_eft()
+    print(SEPARATOR)
     count_horner_basic()
+    print(SEPARATOR)
     count_horner_compensated()
+    print(SEPARATOR)
     count_de_casteljau_basic()
+    print(SEPARATOR)
+    count_de_casteljau_compensated()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/de_casteljau.py

@@ -38,16 +38,50 @@ This module provides both the standard version and a compensated version.
 """
 
 
+import eft
+
+
 def basic(s, coeffs):
     r = 1.0 - s
 
     degree = len(coeffs) - 1
     pk = list(coeffs)
-    for d in range(degree, 0, -1):
+    for k in range(degree):
         new_pk = []
-        for i in range(d):
-            new_pk.append(r * pk[i] + s * pk[i + 1])
+        for j in range(degree - k):
+            new_pk.append(r * pk[j] + s * pk[j + 1])
         # Update the "current" values.
         pk = new_pk
 
     return pk[0]
+
+
+def _compensated(s, coeffs):
+    r, rho = eft.add_eft(1.0, -s)
+
+    degree = len(coeffs) - 1
+    pk = list(coeffs)
+    ek = [0.0] * (degree + 1)
+    for k in range(degree):
+        new_pk = []
+        new_ek = []
+
+        for j in range(degree - k):
+            prod1, pi = eft.multiply_eft(r, pk[j])
+            prod2, sigma = eft.multiply_eft(s, pk[j + 1])
+            sum_, beta = eft.add_eft(prod1, prod2)
+            new_pk.append(sum_)
+            # Now update the error.
+            w = pi + sigma + beta + pk[j] * rho
+            new_ek.append(w + s * ek[j + 1] + r * ek[j])
+
+        # Update the "current" values.
+        pk = new_pk
+        ek = new_ek
+
+    return pk[0], ek[0]
+
+
+def compensated(x, coeffs):
+    p, e = _compensated(x, coeffs)
+    return p + e

### 2018-04-26 src/operation_count.py

@@ -17,7 +17,7 @@ import fractions
 
 
 _DISPLAY_TEMPLATE = (
-    "{:3d} flops ({:2d} add, {:2d} sub, {:2d} multiply, {:2d} FMA)"
+    "{:3d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:2d} FMA)"
 )
 
 
@@ -80,6 +80,14 @@ class Float(object):
 
         self.computation.add_count += 1
         return Float(self.value + value, self.computation)
+
+    def __radd__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.add_count += 1
+        return Float(value + self.value, self.computation)
 
     def __sub__(self, other):
         value = self._get_value(other)

### 2018-04-26 scripts/compute_counts.py
Adding the "expanded" (K=3) de Casteljau algorithm.

@@ -81,7 +81,7 @@ def count_horner_compensated():
 
 
 def count_de_casteljau_basic():
-    print("de_casteljau.basic() ((3n^2 + 3n + 2)/2):")
+    print("de_casteljau.basic() ((3n^2 + 3n + 2) / 2 = 3 T_n + 1):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -95,7 +95,7 @@ def count_de_casteljau_basic():
 
 
 def count_de_casteljau_compensated():
-    print("de_casteljau.compensated() (9n^2 + 9n + 7):")
+    print("de_casteljau.compensated() (9n^2 + 9n + 7 = 18 T_n + 7):")
     for degree in range(1, 9 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -104,6 +104,20 @@ def count_de_casteljau_compensated():
             for k in range(degree + 1)
         )
         p = de_casteljau.compensated(x, coeffs)
+        assert p.value == 0.5 ** degree
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
+def count_de_casteljau_compensated3():
+    print("de_casteljau.compensated3() ((59n^2 + 59n + 16) / 2 = 59 T_n + 8):")
+    for degree in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = de_casteljau.compensated3(x, coeffs)
         assert p.value == 0.5 ** degree
         print("  degree {}:     {}".format(degree, parent.display))
 
@@ -122,6 +136,8 @@ def main():
     count_de_casteljau_basic()
     print(SEPARATOR)
     count_de_casteljau_compensated()
+    print(SEPARATOR)
+    count_de_casteljau_compensated3()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/de_casteljau.py

@@ -82,6 +82,62 @@ def _compensated(s, coeffs):
     return pk[0], ek[0]
 
 
-def compensated(x, coeffs):
-    p, e = _compensated(x, coeffs)
+def compensated(s, coeffs):
+    p, e = _compensated(s, coeffs)
     return p + e
+
+
+def _compensated3(s, coeffs):
+    r, rho = eft.add_eft(1.0, -s)
+
+    degree = len(coeffs) - 1
+    pk = list(coeffs)
+    eAk = [0.0] * (degree + 1)
+    eBk = [0.0] * (degree + 1)
+    for k in range(degree):
+        new_pk = []
+        new_eAk = []
+        new_eBk = []
+
+        for j in range(degree - k):
+            # Update the "level 0" stuff.
+            P1, pi1 = eft.multiply_eft(r, pk[j])
+            P2, pi2 = eft.multiply_eft(s, pk[j + 1])
+            S3, sigma3 = eft.add_eft(P1, P2)
+            new_pk.append(S3)
+            # Update the "level 1" stuff.
+            S4, sigma4 = eft.add_eft(pi1, pi2)
+            S5, sigma5 = eft.add_eft(S4, sigma3)
+            P6, pi6 = eft.multiply_eft(rho, pk[j])
+            wA, sigma7 = eft.add_eft(S5, P6)
+            P8, pi8 = eft.multiply_eft(s, eAk[j + 1])
+            S9, sigma9 = eft.add_eft(P8, wA)
+            P10, pi10 = eft.multiply_eft(r, eAk[j])
+            S11, sigma11 = eft.add_eft(S9, P10)
+            new_eAk.append(S11)
+            # Update the "level 2" stuff.
+            wB = (
+                sigma4
+                + sigma5
+                + pi6
+                + sigma7
+                + pi8
+                + sigma9
+                + pi10
+                + sigma11
+                + rho
+                * eAk[j]
+            )
+            new_eBk.append(wB + s * eBk[j + 1] + r * eBk[j])
+
+        # Update the "current" values.
+        pk = new_pk
+        eAk = new_eAk
+        eBk = new_eBk
+
+    return pk[0], eAk[0], eBk[0]
+
+
+def compensated3(s, coeffs):
+    p, eA, eB = _compensated3(s, coeffs)
+    return (p + eA) + eB

### 2018-04-26 src/operation_count.py

@@ -17,7 +17,7 @@ import fractions
 
 
 _DISPLAY_TEMPLATE = (
-    "{:3d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:2d} FMA)"
+    "{:4d} flops ({:3d} add, {:4d} sub, {:3d} multiply, {:3d} FMA)"
 )
 
 
@@ -117,13 +117,15 @@ class Float(object):
         return Float(self.value * value, self.computation)
 
     def fma(self, val1, val2, val3):
-        for val in (val1, val2, val3):
-            if not isinstance(val, Float):
-                raise TypeError("Only `Float` allowed in fma", val)
+        float1 = self._get_value(val1)
+        float2 = self._get_value(val2)
+        float3 = self._get_value(val3)
+        if None in (float1, float2, float3):
+            raise TypeError("Only `Float` or `float` allowed in fma")
 
         self.computation.fma_count += 1
-        frac1 = fractions.Fraction(val1.value)
-        frac2 = fractions.Fraction(val2.value)
-        frac3 = fractions.Fraction(val3.value)
+        frac1 = fractions.Fraction(float1)
+        frac2 = fractions.Fraction(float2)
+        frac3 = fractions.Fraction(float3)
         result = float(frac1 * frac2 + frac3)
         return Float(result, self.computation)

### 2018-04-26 scripts/compute_counts.py
Adding `SumK` from original EFT paper.

@@ -58,6 +58,30 @@ def count_multiply_eft():
         print("  {} {}".format(description, parent.display))
 
 
+def count__vec_sum():
+    print("_vec_sum() (6(|p| - 1)):")
+    for size_p in range(1, 9 + 1):
+        parent = operation_count.Computation()
+        p = [operation_count.Float(1.0, parent)] * size_p
+        eft._vec_sum(p)
+
+        assert p[size_p - 1].value == float(size_p)
+        print("  |p| = {}:     {}".format(size_p, parent.display))
+
+
+def count_sum_k():
+    print("sum_k() ((6K - 5)(|p| - 1)):")
+    for k in (2, 3, 4, 5):
+        print("  K = {}".format(k))
+        for size_p in range(1, 9 + 1):
+            parent = operation_count.Computation()
+            p = [operation_count.Float(1.0, parent)] * size_p
+            total = eft.sum_k(p, k)
+
+            assert total.value == float(size_p)
+            print("    |p| = {}:    {}".format(size_p, parent.display))
+
+
 def count_horner_basic():
     print("horner.basic() (2n):")
     for degree in range(1, 9 + 1):
@@ -128,6 +152,10 @@ def main():
     count__split()
     print(SEPARATOR)
     count_multiply_eft()
+    print(SEPARATOR)
+    count__vec_sum()
+    print(SEPARATOR)
+    count_sum_k()
     print(SEPARATOR)
     count_horner_basic()
     print(SEPARATOR)

### 2018-04-26 src/eft.py

@@ -44,3 +44,26 @@ def multiply_eft(val1, val2, use_fma=True):
         )
 
     return product, error
+
+
+def _vec_sum(p):
+    # See: https://doi.org/10.1137/030601818
+    # Helper for ``sum_k``.
+    # NOTE: This modifies ``p`` in place.
+    n = len(p)
+    for i in range(1, n):
+        p[i], p[i - 1] = add_eft(p[i], p[i - 1])
+
+
+def sum_k(p, k):
+    # See: https://doi.org/10.1137/030601818
+    p = list(p)  # Make a copy to be modified.
+
+    for _ in range(k - 1):
+        _vec_sum(p)
+
+    result = p[0]
+    for p_val in p[1:]:
+        result += p_val
+
+    return result

### 2018-04-26 scripts/compute_counts.py
Dropping the "multi" flop counts from 9 to 5.

@@ -60,7 +60,7 @@ def count_multiply_eft():
 
 def count__vec_sum():
     print("_vec_sum() (6(|p| - 1)):")
-    for size_p in range(1, 9 + 1):
+    for size_p in range(1, 5 + 1):
         parent = operation_count.Computation()
         p = [operation_count.Float(1.0, parent)] * size_p
         eft._vec_sum(p)
@@ -73,7 +73,7 @@ def count_sum_k():
     print("sum_k() ((6K - 5)(|p| - 1)):")
     for k in (2, 3, 4, 5):
         print("  K = {}".format(k))
-        for size_p in range(1, 9 + 1):
+        for size_p in range(1, 5 + 1):
             parent = operation_count.Computation()
             p = [operation_count.Float(1.0, parent)] * size_p
             total = eft.sum_k(p, k)
@@ -84,7 +84,7 @@ def count_sum_k():
 
 def count_horner_basic():
     print("horner.basic() (2n):")
-    for degree in range(1, 9 + 1):
+    for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(2.0, parent)
         coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
@@ -95,7 +95,7 @@ def count_horner_basic():
 
 def count_horner_compensated():
     print("horner.compensated() (11n + 1):")
-    for degree in range(1, 9 + 1):
+    for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(2.0, parent)
         coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
@@ -106,7 +106,7 @@ def count_horner_compensated():
 
 def count_de_casteljau_basic():
     print("de_casteljau.basic() ((3n^2 + 3n + 2) / 2 = 3 T_n + 1):")
-    for degree in range(1, 9 + 1):
+    for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
         coeffs = tuple(
@@ -120,7 +120,7 @@ def count_de_casteljau_basic():
 
 def count_de_casteljau_compensated():
     print("de_casteljau.compensated() (9n^2 + 9n + 7 = 18 T_n + 7):")
-    for degree in range(1, 9 + 1):
+    for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
[…]

### 2018-04-26 scripts/compute_counts.py
Adding Horner compensated (K=4) to flop counts.

@@ -104,6 +104,28 @@ def count_horner_compensated():
         print("  degree {}:     {}".format(degree, parent.display))
 
 
+def count_horner_compensated3():
+    print("horner.compensated3() (32n + 46, n >= 2):")
+    for degree in range(2, 6 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.compensated3(x, coeffs)
+        assert p.value == 2.0 ** (degree + 1) - 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
+def count_horner_compensated4():
+    print("horner.compensated4() (72n + 138, n >= 3):")
+    for degree in range(3, 7 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.compensated4(x, coeffs)
+        assert p.value == 2.0 ** (degree + 1) - 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
 def count_de_casteljau_basic():
     print("de_casteljau.basic() ((3n^2 + 3n + 2) / 2 = 3 T_n + 1):")
     for degree in range(1, 5 + 1):
@@ -160,6 +182,10 @@ def main():
     count_horner_basic()
     print(SEPARATOR)
     count_horner_compensated()
+    print(SEPARATOR)
+    count_horner_compensated3()
+    print(SEPARATOR)
+    count_horner_compensated4()
     print(SEPARATOR)
     count_de_casteljau_basic()
     print(SEPARATOR)

### 2018-04-26 src/horner.py

@@ -40,6 +40,9 @@ import eft
 
 
 def basic(x, coeffs):
+    if not coeffs:
+        return 0.0
+
     p = coeffs[0]
     for coeff in coeffs[1:]:
         p = p * x + coeff
@@ -48,16 +51,67 @@ def basic(x, coeffs):
 
 
 def _compensated(x, coeffs):
+    if not coeffs:
+        return 0.0, [], []
+
     p = coeffs[0]
-    e = 0.0
+    e_pi = []
+    e_sigma = []
     for coeff in coeffs[1:]:
         prod, e1 = eft.multiply_eft(p, x)
         p, e2 = eft.add_eft(prod, coeff)
-        e = x * e + (e1 + e2)
+        e_pi.append(e1)
+        e_sigma.append(e2)
 
-    return p, e
+    return p, e_pi, e_sigma
 
 
 def compensated(x, coeffs):
-    p, e = _compensated(x, coeffs)
+    p, e_pi, e_sigma = _compensated(x, coeffs)
+
+    # Compute the error via standard Horner's.
+    e = 0.0
+    for e1, e2 in zip(e_pi, e_sigma):
+        e = x * e + (e1 + e2)
+
     return p + e
+
+
+def compensated3(x, coeffs):
+    h1, p2, p3 = _compensated(x, coeffs)
+    h2, p4, p5 = _compensated(x, p2)
+    h3, p6, p7 = _compensated(x, p3)
+
+    # Use standard Horner from here.
+    h4 = basic(x, p4)
+    h5 = basic(x, p5)
+    h6 = basic(x, p6)
+    h7 = basic(x, p7)
+
+    # Now use 3-fold summation.
+    p = [h1, h2, h3, h4, h5, h6, h7]
+    return eft.sum_k(p, 3)
+
+
+def compensated4(x, coeffs):
+    h1, p2, p3 = _compensated(x, coeffs)
+    h2, p4, p5 = _compensated(x, p2)
+    h3, p6, p7 = _compensated(x, p3)
+    h4, p8, p9 = _compensated(x, p4)
+    h5, p10, p11 = _compensated(x, p5)
+    h6, p12, p13 = _compensated(x, p6)
+    h7, p14, p15 = _compensated(x, p7)
+
+    # Use standard Horner from here.
+    h8 = basic(x, p8)
+    h9 = basic(x, p9)
+    h10 = basic(x, p10)
[…]

### 2018-04-26 src/operation_count.py

@@ -17,7 +17,7 @@ import fractions
 
 
 _DISPLAY_TEMPLATE = (
-    "{:4d} flops ({:3d} add, {:4d} sub, {:3d} multiply, {:3d} FMA)"
+    "{:3d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:2d} FMA)"
 )
 
 

### 2018-04-26 scripts/compute_counts.py
Adding de Casteljau compensated (K=4) to flop counts.
I have determined that there will be

- FMA: (3K - 4)Tn flops
- Mul: (3K - 1)Tn

and the `Sub`|`Add` flops are

- K = 2 -->  4 Tn + 4 |  7 Tn + 3
- K = 3 --> 24 Tn + 4 | 22 Tn + 4
- K = 4 --> 64 Tn + 4 | 47 Tn + 5

I am **close** to "finding" a formula.

@@ -168,6 +168,20 @@ def count_de_casteljau_compensated3():
         print("  degree {}:     {}".format(degree, parent.display))
 
 
+def count_de_casteljau_compensated4():
+    print("de_casteljau.compensated4() (65n^2 + 65n + 9 = 130 T_n + 9):")
+    for degree in range(1, 5 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = de_casteljau.compensated4(x, coeffs)
+        assert p.value == 0.5 ** degree
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
 def main():
     count_add_eft()
     print(SEPARATOR)
@@ -192,6 +206,8 @@ def main():
     count_de_casteljau_compensated()
     print(SEPARATOR)
     count_de_casteljau_compensated3()
+    print(SEPARATOR)
+    count_de_casteljau_compensated4()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/de_casteljau.py

@@ -141,3 +141,82 @@ def _compensated3(s, coeffs):
 def compensated3(s, coeffs):
     p, eA, eB = _compensated3(s, coeffs)
     return (p + eA) + eB
+
+
+def _compensated4(s, coeffs):
+    r, rho = eft.add_eft(1.0, -s)
+
+    degree = len(coeffs) - 1
+    pk = list(coeffs)
+    eAk = [0.0] * (degree + 1)
+    eBk = [0.0] * (degree + 1)
+    eCk = [0.0] * (degree + 1)
+    for k in range(degree):
+        new_pk = []
+        new_eAk = []
+        new_eBk = []
+        new_eCk = []
+
+        for j in range(degree - k):
+            # Update the "level 0" stuff.
+            P1, pi1 = eft.multiply_eft(r, pk[j])
+            P2, pi2 = eft.multiply_eft(s, pk[j + 1])
+            S3, sigma3 = eft.add_eft(P1, P2)
+            new_pk.append(S3)
+            # Update the "level 1" stuff.
+            S4, sigma4 = eft.add_eft(pi1, pi2)
+            S5, sigma5 = eft.add_eft(S4, sigma3)
+            P6, pi6 = eft.multiply_eft(rho, pk[j])
+            wA, sigma7 = eft.add_eft(S5, P6)
+            P8, pi8 = eft.multiply_eft(s, eAk[j + 1])
+            S9, sigma9 = eft.add_eft(P8, wA)
+            P10, pi10 = eft.multiply_eft(r, eAk[j])
+            S11, sigma11 = eft.add_eft(S9, P10)
+            new_eAk.append(S11)
+            # Update the "level 2" stuff.
+            S12, sigma12 = eft.add_eft(sigma4, sigma5)
+            S14, sigma14 = eft.add_eft(S12, pi6)
+            S15, sigma15 = eft.add_eft(S14, sigma7)
+            S16, sigma16 = eft.add_eft(S15, pi8)
+            S17, sigma17 = eft.add_eft(S16, sigma9)
+            S18, sigma18 = eft.add_eft(S17, pi10)
+            S19, sigma19 = eft.add_eft(S18, sigma11)
+            P13, pi13 = eft.multiply_eft(rho, eAk[j])
+            wB, sigma20 = eft.add_eft(S19, P13)
+            P21, pi21 = eft.multiply_eft(s, eBk[j + 1])
[…]

### 2018-04-26 src/operation_count.py

@@ -17,7 +17,7 @@ import fractions
 
 
 _DISPLAY_TEMPLATE = (
-    "{:3d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:2d} FMA)"
+    "{:4d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:3d} FMA)"
 )
 
 

### 2018-04-26 scripts/compute_counts.py
Adding de Casteljau compensated (K=5) to flop counts.
It seems that the four flop types have the following counts

- FMA: (3K - 4)Tn
- Mul: (3K - 1)Tn
- Sub: (10K^2 - 30K + 24)Tn + 4
- Add: (5K^2 - 10K + 7)Tn + K + 1
------------------
- Tot: (15K^2 - 34K + 26)Tn + K + 5

@@ -66,7 +66,7 @@ def count__vec_sum():
         eft._vec_sum(p)
 
         assert p[size_p - 1].value == float(size_p)
-        print("  |p| = {}:     {}".format(size_p, parent.display))
+        print("  |p| = {}:      {}".format(size_p, parent.display))
 
 
 def count_sum_k():
@@ -182,6 +182,24 @@ def count_de_casteljau_compensated4():
         print("  degree {}:     {}".format(degree, parent.display))
 
 
+def count_de_casteljau_compensated5():
+    msg = (
+        "de_casteljau.compensated5() ((231n^2 + 231n + 20) / 2 "
+        "= 231 T_n + 10):"
+    )
+    print(msg)
+    for degree in range(1, 5 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = de_casteljau.compensated5(x, coeffs)
+        assert p.value == 0.5 ** degree
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
 def main():
     count_add_eft()
     print(SEPARATOR)
@@ -208,6 +226,8 @@ def main():
     count_de_casteljau_compensated3()
     print(SEPARATOR)
     count_de_casteljau_compensated4()
+    print(SEPARATOR)
+    count_de_casteljau_compensated5()
 
 
 if __name__ == "__main__":

### 2018-04-26 src/de_casteljau.py

@@ -61,25 +61,25 @@ def _compensated(s, coeffs):
 
     degree = len(coeffs) - 1
     pk = list(coeffs)
-    ek = [0.0] * (degree + 1)
+    eAk = [0.0] * (degree + 1)
     for k in range(degree):
         new_pk = []
-        new_ek = []
+        new_eAk = []
 
         for j in range(degree - k):
-            prod1, pi = eft.multiply_eft(r, pk[j])
-            prod2, sigma = eft.multiply_eft(s, pk[j + 1])
-            sum_, beta = eft.add_eft(prod1, prod2)
-            new_pk.append(sum_)
+            P1, pi1 = eft.multiply_eft(r, pk[j])
+            P2, pi2 = eft.multiply_eft(s, pk[j + 1])
+            S3, sigma3 = eft.add_eft(P1, P2)
+            new_pk.append(S3)
             # Now update the error.
-            w = pi + sigma + beta + pk[j] * rho
-            new_ek.append(w + s * ek[j + 1] + r * ek[j])
+            wA = pi1 + pi2 + sigma3 + rho * pk[j]
+            new_eAk.append(wA + s * eAk[j + 1] + r * eAk[j])
 
         # Update the "current" values.
         pk = new_pk
-        ek = new_ek
+        eAk = new_eAk
 
-    return pk[0], ek[0]
+    return pk[0], eAk[0]
 
 
 def compensated(s, coeffs):
@@ -106,10 +106,10 @@ def _compensated3(s, coeffs):
             S3, sigma3 = eft.add_eft(P1, P2)
             new_pk.append(S3)
             # Update the "level 1" stuff.
-            S4, sigma4 = eft.add_eft(pi1, pi2)
-            S5, sigma5 = eft.add_eft(S4, sigma3)
+            S4, sigma4 = eft.add_eft(pi1, pi2)  # wA = pi1 + pi2 + ...
+            S5, sigma5 = eft.add_eft(S4, sigma3)  # ... + sigma3 + ...
             P6, pi6 = eft.multiply_eft(rho, pk[j])
-            wA, sigma7 = eft.add_eft(S5, P6)
+            wA, sigma7 = eft.add_eft(S5, P6)  # ... + rho * pk[j]
             P8, pi8 = eft.multiply_eft(s, eAk[j + 1])
             S9, sigma9 = eft.add_eft(P8, wA)
             P10, pi10 = eft.multiply_eft(r, eAk[j])
[…]

### 2018-04-26 src/operation_count.py

@@ -17,7 +17,7 @@ import fractions
 
 
 _DISPLAY_TEMPLATE = (
-    "{:4d} flops ({:3d} add, {:3d} sub, {:3d} multiply, {:3d} FMA)"
+    "{:4d} flops ({:4d} add, {:4d} sub, {:3d} multiply, {:3d} FMA)"
 )
 
 

### 2018-04-26 doc/paper.tex
Add remark about flop count in PDF.
Adding a bibliography with some quotes.

@@
 \usepackage{amsthm,amssymb,amsmath}
+%% H/T: https://tex.stackexchange.com/a/202168/32270
+\usepackage{textcomp}
 
+\usepackage[usenames, dvipsnames]{color}
 \usepackage{hyperref}
@@
     b_j^{(0)} &= p_j \\
-    b_j^{(k)} &= (1 - x) b_j^{(k - 1)} + x b_{j + 1}^{(k - 1)} \\
-    b(x) &= b_0^{(n)}.
+    b_j^{(k)} &= (1 - s) b_j^{(k - 1)} + s b_{j + 1}^{(k - 1)} \\
+    b(s) &= b_0^{(n)}.
 \end{align*}
 
+\subsection{Selection of Test Cases}
+
+From \cite{Delgado2015}
+
+\begin{quote}
+  We can observe that, in this case, the algorithm with a good
+  behavior everywhere is the de Casteljau algorithm
+\end{quote}
+
+\noindent In the same paper (when referring to \cite{Bezerra2013}):
+
+\begin{quote}
+  assuming that all control points are positive. This assumption avoided
+  ill-conditioned polynomials. In this section, we shall show that this is
+  a natural assumption in Computer Aided Geometric Design (from now on,
+  C.A.G.D.) and that it permits to assure high relative precision for the
+  evaluation through a large family of representations in C.A.G.D.
+\end{quote}
+
+\noindent From the same author, in \cite{Mainar2005}:
+
+\begin{quote}
+  Let us observe that in this case, the de Casteljau algorithm presents
+  better stability properties for the evaluation near the roots. In fact,
+  the de Casteljau algorithm has good behaviour even when using simple
+  precision, although the running error bound is not so accurate in points
+  close to the roots.
+\end{quote}
+
+\subsection{\texorpdfstring{\(K\)}{K}-Fold Error Filtering}
+
+Empirically, it seems the process takes
+\[(15K^2 - 34K + 26)T_n + K + 5\]
+flops to evaluate a degree \(n\) polynomial. (Here \(T_n\) is the
+\(n\)th triangular number.)
+
+\section{Bogus Section for Refs}
+
+Here they are, for now
+\begin{itemize}
+  \item Compensated Horner (\(K = 2\)) (\cite{langlois_et_al:DSP:2006:442})
+  \item Compensated de Casteljau (\cite{Jiang2010})
+  \item Newton with compensated Horner (\cite{Graillat2008})
+  \item \(K\)-fold Sum (\cite{Ogita2005})
+  \item \(K\)-fold Horner (\cite{Graillat2009})
+\end{itemize}
+
+\bibliography{paper}
+\bibliographystyle{alpha}
+
 \end{document}

### 2018-04-26 nox.py

@@ -41,15 +41,22 @@ class Remove(object):
             os.remove(path)
 
 
-def build_tex_file(session, base, new_id, extensions=()):
+def build_tex_file(session, base, new_id, extensions=(), with_bibtex=False):
     # NOTE: This assumes that ``session.chdir(get_path('doc'))``
     #       has been called.
     modify_id = get_path("scripts", "modify_pdf_id.py")
 
-    # Run ``pdflatex`` thrice.
-    session.run("pdflatex", base)
-    session.run("pdflatex", base)
-    session.run("pdflatex", base)
+    if with_bibtex:
+        session.run("pdflatex", base)
+        session.run("bibtex", base)
+        session.run("pdflatex", base)
+        session.run("bibtex", base)
+        session.run("pdflatex", base)
+    else:
+        # Run ``pdflatex`` thrice.
+        session.run("pdflatex", base)
+        session.run("pdflatex", base)
+        session.run("pdflatex", base)
 
     path = get_path("doc", base)
     remove = Remove(path, extensions)
@@ -62,6 +69,9 @@ def build_tex(session):
     if py.path.local.sysfind("pdflatex") is None:
         session.skip("`pdflatex` must be installed")
 
+    if py.path.local.sysfind("bibtex") is None:
+        session.skip("`bibtex` must be installed")
+
     # No need to create a virtualenv.
     session.virtualenv = False
 
@@ -72,6 +82,7 @@ def build_tex(session):
         "paper",
         "F092359D979FDC08931DA1922F3E123E",
         extensions=("aux", "log", "out", "toc"),
+        with_bibtex=True,
     )
 
 

### 2018-04-26 doc/paper.tex
Adding page / context to the three included quotes.

@@
 
-From \cite{Delgado2015}
+From \cite{Delgado2015} (end of Section 3):
 
@@
 
-\noindent In the same paper (when referring to \cite{Bezerra2013}):
+\noindent In the same paper (when referring to \cite{Bezerra2013} at the
+beginning of Section 2):
 
@@
 
-\noindent From the same author, in \cite{Mainar2005}:
+\noindent From the same author, in \cite{Mainar2005} (towards the
+end of Section 5, at the bottom of page 109):
 

### 2018-04-26 doc/paper.tex
Adding section about condition numbers.

@@
 \renewcommand{\headrulewidth}{0pt}
-\newcommand{\fl}{\operatorname{fl}}
+\newcommand{\cond}[1]{\operatorname{cond}\left(#1\right)}
+\newcommand{\fl}[1]{\operatorname{fl}\left(#1\right)}
 \newcommand{\eps}{\varepsilon}
@@
 \end{align*}
+
+\subsection{Condition Number}
+
+For a polynomial \(p(x)\) in the power basis, we have
+(\cite{langlois_et_al:DSP:2006:442}):
+\[\cond{p(x)} = \frac{\widetilde{p}\left(\left|x\right|\right)}{
+  \left|p(x)\right|} = \frac{\sum_j \left|a_j\right| \left|x\right|^j}{
+  \left|p(x)\right|}.\]
+In particular, this means that if \(x \geq 0\) and each \(a_j \geq 0\)
+we must necessarily have \(\cond{p(x)} = 1\). To see an example in use,
+consider \(p(x) = (x - 1)^n\) and input values of the form \(x = 1 + \delta\)
+(with \(\left|\delta\right| \ll 1\)). Since \(a_j = \binom{n}{j} (-1)^{n - j}\)
+we have \(\widetilde{p}(x) = (x + 1)^n\) hence
+\[\cond{p\left(1 + \delta\right)} = \frac{(2 + \delta)^n}{
+  \left|\delta\right|^n} = \left|1 + \frac{2}{\delta}\right|^n.\]
+As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
+
+For a polynomial \(p(s)\) in Bernstein form, we have (\cite{Jiang2010}):
+\[\cond{p(s)} = \frac{\widetilde{p}\left(s\right)}{
+  \left|p(s)\right|} = \frac{\sum_j \left|p_j\right| \left|b_{j, n}(s)\right|}{
+  \left|p(s)\right|}.\]
+The Bernstein form is suited for \(s \in \left[0, 1\right]\), which means
+\(b_{j, n}(s) \geq 0\) typically. If \(s \in \left[0, 1\right]\) and each
+\(p_j \geq 0\) we must necessarily have \(\cond{p(s)} = 1\). To see an
+example in use, consider
+\[p(s) = (1 - 2s)^n = \left[(1 - s) - s\right]^n = \sum_j \binom{n}{j}
+(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j b_{j, n}(s)\]
+and input values of the form \(x = \frac{1}{2} + \delta\)
+(with \(\left|\delta\right| \ll \frac{1}{2}\)). Since \(p_j = (-1)^j\)
+we have \(\widetilde{p}(s) = \left[(1 - s) + s\right]^n = 1\)
+\[\cond{p\left(\frac{1}{2} + \delta\right)} = \frac{1}{
+  \left|2\delta\right|^n}.\]
+As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
 

### 2018-04-26 doc/paper.tex
Implementing the generic K-fold compensated Horner's.
Also adding the full flop counts to the write-up.

@@
 
-Empirically, it seems the process takes
+After implementing for \(K = 2, 3, \ldots, 12\) and instrumenting all
+relevant floating point operations, the \(K\)-fold Horner requires
+\[(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right)\]
+flops to evaluate a degree \(n\) polynomial. As a comparison, the
+non-compensated form of Horner requires \(2n\) flops. Of these,
+\(\left(2^{K - 1} - 1\right)n - 2^{K - 1}(K - 3) - 2\) are
+FMA (fused-multiply-add) instructions.
+
+After implementing for \(K = 2, 3, 4, 5\) and instrumenting all relevant
+floating point operations, the \(K\)-fold de Casteljau requires
 \[(15K^2 - 34K + 26)T_n + K + 5\]
 flops to evaluate a degree \(n\) polynomial. (Here \(T_n\) is the
-\(n\)th triangular number.)
+\(n\)th triangular number.) As a comparison, the non-compensated form of
+de Casteljau requires \(3 T_n + 1\) flops. Of these, \((3K - 4)T_n\) are
+FMA instructions. On hardware that doesn't support FMA,
+every FMA will be exchanged for 10 \(\ominus\)'s and 6 \(\otimes\)'s so the
+count will increase by \((10 + 6 - 1)(3K - 4)T_n\).
 

### 2018-04-26 scripts/compute_counts.py

@@ -103,6 +103,17 @@ def count_horner_compensated():
         assert p.value == 2.0 ** (degree + 1) - 1
         print("  degree {}:     {}".format(degree, parent.display))
 
+    # NOTE: This is **the same** as ``horner.compensated()`` but uses
+    #       a different algorithm.
+    print("horner.compensated_k(..., 2) (12n + 10):")
+    for degree in range(1, 5 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.compensated_k(x, coeffs, 2)
+        assert p.value == 2.0 ** (degree + 1) - 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
 
 def count_horner_compensated3():
     print("horner.compensated3() (32n + 46, n >= 2):")
@@ -114,14 +125,45 @@ def count_horner_compensated3():
         assert p.value == 2.0 ** (degree + 1) - 1
         print("  degree {}:     {}".format(degree, parent.display))
 
+    print("horner.compensated_k(..., 3) (32n + 46, n >= 2):")
+    for degree in range(2, 6 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(2.0, parent)
+        coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
+        p = horner.compensated_k(x, coeffs, 3)
+        assert p.value == 2.0 ** (degree + 1) - 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
 
 def count_horner_compensated4():
-    print("horner.compensated4() (72n + 138, n >= 3):")
+    print("horner.compensated_k(..., 4) (72n + 138, n >= 3):")
     for degree in range(3, 7 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(2.0, parent)
         coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
-        p = horner.compensated4(x, coeffs)
+        p = horner.compensated_k(x, coeffs, 4)
+        assert p.value == 2.0 ** (degree + 1) - 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
+def count_horner_compensated5():
+    print("horner.compensated_k(..., 5) (152n + 350, n >= 4):")
+    for degree in range(4, 8 + 1):
+        parent = operation_count.Computation()
[…]

### 2018-04-26 src/horner.py

@@ -93,25 +93,20 @@ def compensated3(x, coeffs):
     return eft.sum_k(p, 3)
 
 
-def compensated4(x, coeffs):
-    h1, p2, p3 = _compensated(x, coeffs)
-    h2, p4, p5 = _compensated(x, p2)
-    h3, p6, p7 = _compensated(x, p3)
-    h4, p8, p9 = _compensated(x, p4)
-    h5, p10, p11 = _compensated(x, p5)
-    h6, p12, p13 = _compensated(x, p6)
-    h7, p14, p15 = _compensated(x, p7)
+def compensated_k(x, coeffs, k):
+    h = {}
+    p = {1: coeffs}
 
-    # Use standard Horner from here.
-    h8 = basic(x, p8)
-    h9 = basic(x, p9)
-    h10 = basic(x, p10)
-    h11 = basic(x, p11)
-    h12 = basic(x, p12)
-    h13 = basic(x, p13)
-    h14 = basic(x, p14)
-    h15 = basic(x, p15)
+    # First, "filter" off the errors from the interior
+    # polynomials.
+    for i in range(1, 2 ** (k - 1)):
+        h[i], p[2 * i], p[2 * i + 1] = _compensated(x, p[i])
 
-    # Now use 4-fold summation.
-    p = [h1, h2, h3, h4, h5, h6, h7, h8, h9, h10, h11, h12, h13, h14, h15]
-    return eft.sum_k(p, 4)
+    # Then use standard Horner for the leaf polynomials.
+    for i in range(2 ** (k - 1), 2 ** k):
+        h[i] = basic(x, p[i])
+
+    # Now use K-fold summation on everything in ``h`` (but keep the
+    # order).
+    to_sum = [h[i] for i in range(1, 2 ** k)]
+    return eft.sum_k(to_sum, k)

### 2018-04-26 doc/paper.tex
Adding note about asymptotic behavior of flop counts.
Also adding a note about when the Horner count applies.
Added some base algorithm blocks.

@@
 \usepackage{textcomp}
+%% H/T: https://tex.stackexchange.com/a/215651/32270
+\usepackage[ruled]{algorithm2e}
 
@@
 \tableofcontents
+
+\section{IEEE Stuff}
+
+\begin{algorithm}[H]
+  \textit{function} \(\left[x, y\right] = \mathtt{TwoSum}(a, b)\) \\
+  \quad \(x = a \oplus b\) \\
+  \quad \(z = x \ominus a\) \\
+  \quad \(y = (a \ominus (x \ominus z)) \oplus (b \ominus z)\)
+
+  \caption{\textit{EFT of the sum of two floating point numbers.}}
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \textit{function} \(\left[x, y\right] = \mathtt{Split}(a)\) \\
+  \quad \(z = a \otimes (2^r + 1)\) \\
+  \quad \(x = z \ominus (z \ominus a)\) \\
+  \quad \(y = a \ominus x\)
+
+  \caption{\textit{Splitting of a floating point number into two parts.}}
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \textit{function} \(\left[x, y\right] = \mathtt{TwoProd}(a, b)\) \\
+  \quad \(x = a \otimes b\) \\
+  \quad \(\left[a_h, a_l\right] = \mathtt{Split}(a)\) \\
+  \quad \(\left[b_h, b_l\right] = \mathtt{Split}(b)\) \\
+  \quad \(y = a_l \otimes b_l \ominus (((x \ominus a_h \otimes b_h)
+      \ominus a_l \otimes b_h) \ominus a_h \otimes b_l)\)
+
+  \caption{\textit{EFT of the product of two floating point numbers.}}
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \textit{function} \(\left[x, y\right] = \mathtt{TwoProdFMA}(a, b)\) \\
+  \quad \(x = a \otimes b\) \\
+  \quad \(y = \mathtt{FMA}(a, b, -x)\)
+
+  \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
+\end{algorithm}
 
@@
 relevant floating point operations, the \(K\)-fold Horner requires
-\[(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right)\]
-flops to evaluate a degree \(n\) polynomial. As a comparison, the
+\[(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right) =
+\mathcal{O}\left((n + K)2^K\right)\]
+flops to evaluate a degree \(n\) polynomial (this only applies when
+\(n \geq K - 1\)). As a comparison, the
 non-compensated form of Horner requires \(2n\) flops. Of these,
@@
 floating point operations, the \(K\)-fold de Casteljau requires
-\[(15K^2 - 34K + 26)T_n + K + 5\]
+\[(15K^2 - 34K + 26)T_n + K + 5 =
+\mathcal{O}\left(n^2 K^2\right)\]
 flops to evaluate a degree \(n\) polynomial. (Here \(T_n\) is the

### 2018-04-26 doc/paper.tex
Added pseudocode for basic de Casteljau.
Also changing approach to algorithm in LaTeX.
Adding compensated de Casteljau as an algorithm.
Also adding the forgotten `res` and `n` variables.
Adding an abstract.

@@
 \usepackage{textcomp}
-%% H/T: https://tex.stackexchange.com/a/215651/32270
-\usepackage[ruled]{algorithm2e}
+%% H/T: https://tex.stackexchange.com/a/56877/32270
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 
+\begin{abstract}
+In computer aided geometric design a polynomial is usually represented in
+Bernstein form. This paper presents a family of compensated algorithms to
+accurately evaluate a polynomial in Bernstein form with floating point
+coefficients. The principle is to apply error-free transformations to
+improve the traditional de Casteljau algorithm. At each stage of computation,
+round-off error is passed on to first order errors, then to second order
+errors, and so on. After the computation has been ``filtered'' \(K\)-times
+via this process, the resulting output is as accurate as the de Casteljau
+algorithm performed in \(K\) times the working precision. Forward error
+analysis and numerical experiments illustrate the accuracy of this family
+of algorithms.
+\end{abstract}
+
 \tableofcontents
@@
 \begin{algorithm}[H]
-  \textit{function} \(\left[x, y\right] = \mathtt{TwoSum}(a, b)\) \\
-  \quad \(x = a \oplus b\) \\
-  \quad \(z = x \ominus a\) \\
-  \quad \(y = (a \ominus (x \ominus z)) \oplus (b \ominus z)\)
+  \caption{\textit{EFT of the sum of two floating point numbers.}}
 
-  \caption{\textit{EFT of the sum of two floating point numbers.}}
+  \begin{algorithmic}
+    \Function{\(\left[x, y\right] = \mathtt{TwoSum}\)}{$a, b$}
+      \State \(x = a \oplus b\)
+      \State \(z = x \ominus a\)
+      \State \(y = (a \ominus (x \ominus z)) \oplus (b \ominus z)\)
+    \EndFunction
+  \end{algorithmic}
 \end{algorithm}
@@
 \begin{algorithm}[H]
-  \textit{function} \(\left[x, y\right] = \mathtt{Split}(a)\) \\
-  \quad \(z = a \otimes (2^r + 1)\) \\
-  \quad \(x = z \ominus (z \ominus a)\) \\
-  \quad \(y = a \ominus x\)
+  \caption{\textit{Splitting of a floating point number into two parts.}}
 
-  \caption{\textit{Splitting of a floating point number into two parts.}}
+  \begin{algorithmic}
+    \Function{\(\left[x, y\right] = \mathtt{Split}\)}{$a$}
+      \State \(z = a \otimes (2^r + 1)\)
+      \State \(x = z \ominus (z \ominus a)\)
+      \State \(y = a \ominus x\)
+    \EndFunction
+  \end{algorithmic}
 \end{algorithm}
@@
 \begin{algorithm}[H]
-  \textit{function} \(\left[x, y\right] = \mathtt{TwoProd}(a, b)\) \\
-  \quad \(x = a \otimes b\) \\
-  \quad \(\left[a_h, a_l\right] = \mathtt{Split}(a)\) \\
-  \quad \(\left[b_h, b_l\right] = \mathtt{Split}(b)\) \\
-  \quad \(y = a_l \otimes b_l \ominus (((x \ominus a_h \otimes b_h)
+  \caption{\textit{EFT of the product of two floating point numbers.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[x, y\right] = \mathtt{TwoProd}\)}{$a, b$}
+      \State \(x = a \otimes b\)
+      \State \(\left[a_h, a_l\right] = \mathtt{Split}(a)\)
+      \State \(\left[b_h, b_l\right] = \mathtt{Split}(b)\)
+      \State \(y = a_l \otimes b_l \ominus (((x \ominus a_h \otimes b_h)
           \ominus a_l \otimes b_h) \ominus a_h \otimes b_l)\)
-
-  \caption{\textit{EFT of the product of two floating point numbers.}}
+    \EndFunction
+  \end{algorithmic}
 \end{algorithm}
@@
 \begin{algorithm}[H]
-  \textit{function} \(\left[x, y\right] = \mathtt{TwoProdFMA}(a, b)\) \\
-  \quad \(x = a \otimes b\) \\
-  \quad \(y = \mathtt{FMA}(a, b, -x)\)
+  \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
 
-  \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
+  \begin{algorithmic}
+    \Function{\(\left[x, y\right] = \mathtt{TwoProdFMA}\)}{$a, b$}
+      \State \(x = a \otimes b\)
+      \State \(y = \mathtt{FMA}(a, b, -x)\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \caption{\textit{de Casteljau algorithm for polynomial evaluation.}}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$p, s$}
+      \State \(n = \texttt{length}(p) - 1\)
+      \State \(r = 1 \ominus s\)
+      \\
+      \For{\(j = 0, \ldots, n\)}
+        \State \(b_j^{(n)} = p_j\)
+      \EndFor
+      \\
+      \For{\(k = n - 1, \ldots, 0\)}
+        \For{\(j = 0, \ldots, k\)}
+          \State \(b_j^{(k)} = \left(r \otimes b_j^{(k + 1)}\right) \oplus
+              \left(s \otimes b_{j + 1}^{(k + 1)}\right)\)
+        \EndFor
+      \EndFor
+      \\
+      \State \(\mathtt{res} = b_0^{(0)}\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \caption{\textit{Compensated de Casteljau algorithm for polynomial evaluation.}}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$p, s$}
+      \State \(n = \texttt{length}(p) - 1\)
+      \State \(\left[r, \rho\right] = \mathtt{TwoSum}(1, -s)\)
+      \\
+      \For{\(j = 0, \ldots, n\)}
+        \State \(b_j^{(n)} = p_j\)
+        \State \(e_j^{(n)} = 0\)
+      \EndFor
+      \\
+      \For{\(k = n - 1, \ldots, 0\)}
+        \For{\(j = 0, \ldots, k\)}
+          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
+              r, b_j^{(k + 1)}\right)\)
+          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
+              s, b_{j + 1}^{(k + 1)}\right)\)
+          \State \(\left[b_j^{(k)}, \sigma_3\right] = \mathtt{TwoSum}(
+              P_1, P_2)\)
+          \State \(w = \pi_1 \oplus \pi_2 \oplus \sigma_3 \oplus
+              \left(\rho \otimes b_j^{(k + 1)}\right)\)
+          \State \(e_j^{(k)} = w \oplus \left(s \otimes e_{j + 1}^{(k + 1)}
+              \right) \oplus \left(r \otimes e_j^{(k + 1)}\right)\)
+        \EndFor
+      \EndFor
+      \\
+      \State \(\mathtt{res} = b_0^{(0)} \oplus e_0^{(0)}\)
+    \EndFunction
+  \end{algorithmic}
 \end{algorithm}

### 2018-04-27 src/eft.py
Adding an example of **compensated** de Casteljau failing.

@@ -13,6 +13,9 @@
 """Collection of error-free transforms."""
 
 
+import fractions
+
+
 def add_eft(val1, val2):
     # See: https://doi.org/10.1137/030601818
     sum_ = val1 + val2
@@ -29,11 +32,25 @@ def _split(val):
     return high_bits, low_bits
 
 
+def _fma(val1, val2, val3):
+    if (
+        isinstance(val1, float)
+        and isinstance(val2, float)
+        and isinstance(val3, float)
+    ):
+        frac1 = fractions.Fraction(val1)
+        frac2 = fractions.Fraction(val2)
+        frac3 = fractions.Fraction(val3)
+        return float(frac1 * frac2 + frac3)
+    else:
+        return val1.fma(val1, val2, val3)
+
+
 def multiply_eft(val1, val2, use_fma=True):
     # See: https://doi.org/10.1109/TC.2008.215
     product = val1 * val2
     if use_fma:
-        error = val1.fma(val1, val2, -product)
+        error = _fma(val1, val2, -product)
     else:
         high1, low1 = _split(val1)
         high2, low2 = _split(val2)

### 2018-04-27 doc/paper.tex
Using `P/pi` and `S/sigma` notation for EFTs.
Adding an example of **compensated** de Casteljau failing.
Updating comment about "bad" outcome for compensated de Casteljau.
Adding more to parenthetical phrase about u^4 term.

@@
 \usepackage{algpseudocode}
+%% H/T: https://tex.stackexchange.com/a/244805
+\usepackage{booktabs,siunitx}
 
@@
   \begin{algorithmic}
-    \Function{\(\left[x, y\right] = \mathtt{TwoSum}\)}{$a, b$}
-      \State \(x = a \oplus b\)
-      \State \(z = x \ominus a\)
-      \State \(y = (a \ominus (x \ominus z)) \oplus (b \ominus z)\)
+    \Function{\(\left[S, \sigma\right] = \mathtt{TwoSum}\)}{$a, b$}
+      \State \(S = a \oplus b\)
+      \State \(z = S \ominus a\)
+      \State \(\sigma = (a \ominus (S \ominus z)) \oplus (b \ominus z)\)
     \EndFunction
@@
   \begin{algorithmic}
-    \Function{\(\left[x, y\right] = \mathtt{Split}\)}{$a$}
+    \Function{\(\left[h, \ell\right] = \mathtt{Split}\)}{$a$}
       \State \(z = a \otimes (2^r + 1)\)
-      \State \(x = z \ominus (z \ominus a)\)
-      \State \(y = a \ominus x\)
+      \State \(h = z \ominus (z \ominus a)\)
+      \State \(\ell = a \ominus h\)
     \EndFunction
@@
   \begin{algorithmic}
-    \Function{\(\left[x, y\right] = \mathtt{TwoProd}\)}{$a, b$}
-      \State \(x = a \otimes b\)
-      \State \(\left[a_h, a_l\right] = \mathtt{Split}(a)\)
-      \State \(\left[b_h, b_l\right] = \mathtt{Split}(b)\)
-      \State \(y = a_l \otimes b_l \ominus (((x \ominus a_h \otimes b_h)
-          \ominus a_l \otimes b_h) \ominus a_h \otimes b_l)\)
+    \Function{\(\left[P, \pi\right] = \mathtt{TwoProd}\)}{$a, b$}
+      \State \(P = a \otimes b\)
+      \State \(\left[a_h, a_{\ell}\right] = \mathtt{Split}(a)\)
+      \State \(\left[b_h, b_{\ell}\right] = \mathtt{Split}(b)\)
+      \State \(\pi = a_{\ell} \otimes b_{\ell} \ominus (((P \ominus
+          a_h \otimes b_h)
+          \ominus a_{\ell} \otimes b_h) \ominus a_h \otimes b_{\ell})\)
     \EndFunction
@@
   \begin{algorithmic}
-    \Function{\(\left[x, y\right] = \mathtt{TwoProdFMA}\)}{$a, b$}
-      \State \(x = a \otimes b\)
-      \State \(y = \mathtt{FMA}(a, b, -x)\)
+    \Function{\(\left[P, \pi\right] = \mathtt{TwoProdFMA}\)}{$a, b$}
+      \State \(P = a \otimes b\)
+      \State \(\pi = \mathtt{FMA}(a, b, -P)\)
     \EndFunction
@@
 \begin{align*}
-    b_j^{(0)} &= p_j \\
-    b_j^{(k)} &= (1 - s) b_j^{(k - 1)} + s b_{j + 1}^{(k - 1)} \\
-    b(s) &= b_0^{(n)}.
+    b_j^{(n)} &= p_j \\
+    b_j^{(k)} &= (1 - s) b_j^{(k + 1)} + s b_{j + 1}^{(k + 1)} \\
+    b(s) &= b_0^{(0)}.
 \end{align*}
@@
 As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
+
+\subsection{Example of Compensated de Casteljau Failing}
+
+Consider
+\[p(s) = (4s - 3)^3 (8s + 7) = (-189) b_{0, 4}(s) + (-54) b_{1, 4}(s) +
+57 b_{2, 4}(s) + (-32) b_{3, 4}(s) + 15 b_{4, 4}(s)\]
+at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
+have \(p(s) = 2^{38.6} u^3 + \mathcal{O}\left(u^4\right)\).
+However, using the compensated de Casteljau algorithm results in
+\(\widehat{p} + \widehat{e} = 0\):
+
+\begin{center}
+  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
+    \toprule
+    k & j & b_j^{(k)} & e_j^{(k)} \\
+    \midrule
+    3 & 0 & -87\frac{3}{4} + 108032u & -32u \\
+    3 & 1 & 29\frac{1}{4} + 88768u & 32u \\
+    3 & 2 & -9\frac{3}{4} - 71200u & 0 \\
+    3 & 3 & 3\frac{1}{4} + 37600u & 0 \\
+    \midrule
+    2 & 0 & 187200u & -15360000u^2 \\
+    2 & 1 & -62408u & 8u - 128000000u^2 \\
+    2 & 2 & 20800u & 87040000u^2 \\
+    \midrule
+    1 & 0 & -6u - 199688192u^2 & \boxed{6u - 99831808u^2} \\
+    1 & 1 & -2u + 66568192u^2  & \boxed{2u + 33271808u^2} \\
+    \midrule
+    0 & 0 & -3u + 7296u^2 & \boxed{3u - 7296u^2} \\
+    \bottomrule
+  \end{tabular}
+\end{center}
+This is due to dropped terms: \(e_0^{(1)}\) drops \(-90112000000u^3\),
+\(e_1^{(1)}\) drops \(172032000000u^3\), \(e_0^{(0)}\) drops
+\(319488000000u^3\) (that's relative to the \textbf{incorrect} \(e_j^{(1)}\)
+terms, if compared to the correct terms there'd also be a \(u^4\) term
+dropped).
 

### 2018-04-27 doc/paper.tex
Updating some notation.
- Switching out `e_j` for `\Delta b_j` (I want to refer to
  the higher order errors in similar ways)
- Dropping `p` for a bold `b` when referring to the polynomial
  in functions
- Dropping usage of `p_j` as Bernstein coefficients
- Changing Bernstein notation from lowercase (`b_{j, n}`) to
  uppercase (`B_{j, n}`)

@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$p, s$}
-      \State \(n = \texttt{length}(p) - 1\)
+    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$\mathbf{b}, s$}
+      \State \(n = \texttt{length}(b) - 1\)
       \State \(r = 1 \ominus s\)
@@
       \For{\(j = 0, \ldots, n\)}
-        \State \(b_j^{(n)} = p_j\)
+        \State \(b_j^{(n)} = b_j\)
       \EndFor
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$p, s$}
-      \State \(n = \texttt{length}(p) - 1\)
+    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$\mathbf{b}, s$}
+      \State \(n = \texttt{length}(\mathbf{b}) - 1\)
       \State \(\left[r, \rho\right] = \mathtt{TwoSum}(1, -s)\)
@@
       \For{\(j = 0, \ldots, n\)}
-        \State \(b_j^{(n)} = p_j\)
-        \State \(e_j^{(n)} = 0\)
+        \State \(b_j^{(n)} = b_j\)
+        \State \(\Delta b_j^{(n)} = 0\)
       \EndFor
@@
               \left(\rho \otimes b_j^{(k + 1)}\right)\)
-          \State \(e_j^{(k)} = w \oplus \left(s \otimes e_{j + 1}^{(k + 1)}
-              \right) \oplus \left(r \otimes e_j^{(k + 1)}\right)\)
+          \State \(\Delta b_j^{(k)} = w \oplus \left(s \otimes
+              \Delta b_{j + 1}^{(k + 1)}
+              \right) \oplus \left(r \otimes \Delta b_j^{(k + 1)}\right)\)
         \EndFor
@@
       \\
-      \State \(\mathtt{res} = b_0^{(0)} \oplus e_0^{(0)}\)
+      \State \(\mathtt{res} = b_0^{(0)} \oplus \Delta b_0^{(0)}\)
     \EndFunction
@@
 Consider de Casteljau's method to evaluate a degree \(n\)
-polynomial in Bernstein-B\'{e}zier form with control points \(p_j\):
+polynomial in Bernstein-B\'{e}zier form with control points \(b_j\):
 \begin{align*}
-    b_j^{(n)} &= p_j \\
+    b_j^{(n)} &= b_j \\
     b_j^{(k)} &= (1 - s) b_j^{(k + 1)} + s b_{j + 1}^{(k + 1)} \\
@@
 \[\cond{p(s)} = \frac{\widetilde{p}\left(s\right)}{
-  \left|p(s)\right|} = \frac{\sum_j \left|p_j\right| \left|b_{j, n}(s)\right|}{
+  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| \left|B_{j, n}(s)\right|}{
   \left|p(s)\right|}.\]
 The Bernstein form is suited for \(s \in \left[0, 1\right]\), which means
-\(b_{j, n}(s) \geq 0\) typically. If \(s \in \left[0, 1\right]\) and each
-\(p_j \geq 0\) we must necessarily have \(\cond{p(s)} = 1\). To see an
+\(B_{j, n}(s) \geq 0\) typically. If \(s \in \left[0, 1\right]\) and each
+\(b_j \geq 0\) we must necessarily have \(\cond{p(s)} = 1\). To see an
 example in use, consider
 \[p(s) = (1 - 2s)^n = \left[(1 - s) - s\right]^n = \sum_j \binom{n}{j}
-(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j b_{j, n}(s)\]
+(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j B_{j, n}(s)\]
 and input values of the form \(x = \frac{1}{2} + \delta\)
-(with \(\left|\delta\right| \ll \frac{1}{2}\)). Since \(p_j = (-1)^j\)
+(with \(\left|\delta\right| \ll \frac{1}{2}\)). Since \(b_j = (-1)^j\)
 we have \(\widetilde{p}(s) = \left[(1 - s) + s\right]^n = 1\)
@@
 Consider
-\[p(s) = (4s - 3)^3 (8s + 7) = (-189) b_{0, 4}(s) + (-54) b_{1, 4}(s) +
-57 b_{2, 4}(s) + (-32) b_{3, 4}(s) + 15 b_{4, 4}(s)\]
+\[p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
+57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)\]
 at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
@@
 However, using the compensated de Casteljau algorithm results in
-\(\widehat{p} + \widehat{e} = 0\):
+\(b_0^{(0)} + \Delta b_0^{(0)} = 0\):
 
@@
     \toprule
-    k & j & b_j^{(k)} & e_j^{(k)} \\
+    k & j & b_j^{(k)} & \Delta b_j^{(k)} \\
     \midrule
@@
 \end{center}
-This is due to dropped terms: \(e_0^{(1)}\) drops \(-90112000000u^3\),
-\(e_1^{(1)}\) drops \(172032000000u^3\), \(e_0^{(0)}\) drops
-\(319488000000u^3\) (that's relative to the \textbf{incorrect} \(e_j^{(1)}\)
-terms, if compared to the correct terms there'd also be a \(u^4\) term
-dropped).
+This is due to dropped terms: \(\Delta b_0^{(1)}\) drops \(-90112000000u^3\),
+\(\Delta b_1^{(1)}\) drops \(172032000000u^3\), \(\Delta b_0^{(0)}\) drops
+\(319488000000u^3\) (that's relative to the \textbf{incorrect}
+\(\Delta b_j^{(1)}\) terms, if compared to the correct terms there'd also
+be a \(u^4\) term dropped).
 

### 2018-04-27 doc/paper.tex
Updating "failed compensation" table.
This way it includes the two columns needed to get the
answer to full precision.

(I verified this with `de_casteljau._compensated4()` and it
matches the values **exactly** at every step.)
Dropping `2^{38.6}` in favor of exact/compact expression.
Changing `Delta` to `partial`.
Also renaming the "K-fold" section to "operation counts"
and reserving the "K-fold" section for future use.
Adding diagram of the error filtering process.

@@
 }
+
+%% H/T: https://tex.stackexchange.com/a/99051/32270
+\PassOptionsToPackage{usenames,dvipsnames}{xcolor}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+\tikzstyle{box} = [
+  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
+  text centered, draw=black]
+\tikzstyle{empty-box} = [
+  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
+  text width=0.05\textwidth]
+\tikzstyle{empty-box-vert} = [
+  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
+  text centered, text height=0.001\textwidth]
+\tikzstyle{line} = [draw, -latex']
 
@@
         \State \(b_j^{(n)} = b_j\)
-        \State \(\Delta b_j^{(n)} = 0\)
+        \State \(\partial b_j^{(n)} = 0\)
       \EndFor
@@
               \left(\rho \otimes b_j^{(k + 1)}\right)\)
-          \State \(\Delta b_j^{(k)} = w \oplus \left(s \otimes
-              \Delta b_{j + 1}^{(k + 1)}
-              \right) \oplus \left(r \otimes \Delta b_j^{(k + 1)}\right)\)
+          \State \(\partial b_j^{(k)} = w \oplus \left(s \otimes
+              \partial b_{j + 1}^{(k + 1)}
+              \right) \oplus \left(r \otimes \partial b_j^{(k + 1)}\right)\)
         \EndFor
@@
       \\
-      \State \(\mathtt{res} = b_0^{(0)} \oplus \Delta b_0^{(0)}\)
+      \State \(\mathtt{res} = b_0^{(0)} \oplus \partial b_0^{(0)}\)
     \EndFunction
@@
 at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
-have \(p(s) = 2^{38.6} u^3 + \mathcal{O}\left(u^4\right)\).
+have \(p(s) = 13(3200u)^3 + \mathcal{O}\left(u^4\right)\).
 However, using the compensated de Casteljau algorithm results in
-\(b_0^{(0)} + \Delta b_0^{(0)} = 0\):
+\(b_0^{(0)} + \partial b_0^{(0)} = 0\):
 
 \begin{center}
-  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
+  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
     \toprule
-    k & j & b_j^{(k)} & \Delta b_j^{(k)} \\
+    k & j & b_j^{(k)} & \partial b_j^{(k)} & \partial^2 b_j^{(k)} & \partial^3 b_j^{(k)} \\
     \midrule
-    3 & 0 & -87\frac{3}{4} + 108032u & -32u \\
-    3 & 1 & 29\frac{1}{4} + 88768u & 32u \\
-    3 & 2 & -9\frac{3}{4} - 71200u & 0 \\
-    3 & 3 & 3\frac{1}{4} + 37600u & 0 \\
+    3 & 0 & -87\frac{3}{4} + 108032u & -32u & 0 & 0 \\
+    3 & 1 & 29\frac{1}{4} + 88768u & 32u & 0 & 0 \\
+    3 & 2 & -9\frac{3}{4} - 71200u & 0 & 0 & 0 \\
+    3 & 3 & 3\frac{1}{4} + 37600u & 0 & 0 & 0 \\
     \midrule
-    2 & 0 & 187200u & -15360000u^2 \\
-    2 & 1 & -62408u & 8u - 128000000u^2 \\
-    2 & 2 & 20800u & 87040000u^2 \\
+    2 & 0 & 187200u & -15360000u^2 & 0 & 0 \\
+    2 & 1 & -62408u & 8u - 128000000u^2 & 0 & 0 \\
+    2 & 2 & 20800u & 87040000u^2 & 0 & 0 \\
     \midrule
-    1 & 0 & -6u - 199688192u^2 & \boxed{6u - 99831808u^2} \\
-    1 & 1 & -2u + 66568192u^2  & \boxed{2u + 33271808u^2} \\
+    1 & 0 & -6u - 199688192u^2 & 6u - 99831808u^2 & -90112000000u^3 & 0 \\
+    1 & 1 & -2u + 66568192u^2  & 2u + 33271808u^2 & 172032000000u^3 & 0 \\
     \midrule
-    0 & 0 & -3u + 7296u^2 & \boxed{3u - 7296u^2} \\
+    0 & 0 & -3u + 7296u^2 & 3u - 7296u^2 & 13 (3200u)^3 + 3048(512u)^4 & 962(128u)^4 \\
     \bottomrule
@@
 \end{center}
-This is due to dropped terms: \(\Delta b_0^{(1)}\) drops \(-90112000000u^3\),
-\(\Delta b_1^{(1)}\) drops \(172032000000u^3\), \(\Delta b_0^{(0)}\) drops
-\(319488000000u^3\) (that's relative to the \textbf{incorrect}
-\(\Delta b_j^{(1)}\) terms, if compared to the correct terms there'd also
-be a \(u^4\) term dropped).
 
@@
 \subsection{\texorpdfstring{\(K\)}{K}-Fold Error Filtering}
+
+\begin{center}
+\begin{tikzpicture}[node distance = 0.2\textwidth]
+    \node [box] (bj) {\(b_j\)};
+    \node [box, below of=bj] (bn0) {\(b_j^{(n)}\)};
+    \node [box, right of=bn0] (dbn0) {\(\partial b_j^{(n)}\)};
+    \node [box, right of=dbn0] (ddbn0) {\(\partial^2 b_j^{(n)}\)};
+    \node [box, right of=ddbn0] (dddbn0) {\(\partial^3 b_j^{(n)}\)};
+    %%
+    \node [box, below of=bn0] (bn1) {\(b_j^{(n - 1)}\)};
+    \node [box, below of=dbn0] (dbn1) {\(\partial b_j^{(n - 1)}\)};
+    \node [box, below of=ddbn0] (ddbn1) {\(\partial^2 b_j^{(n - 1)}\)};
+    \node [box, below of=dddbn0] (dddbn1) {\(\partial^3 b_j^{(n - 1)}\)};
+    %%
+    \node [box, below of=bn1] (bn2) {\(b_j^{(n - 2)}\)};
+    \node [box, below of=dbn1] (dbn2) {\(\partial b_j^{(n - 2)}\)};
+    \node [box, below of=ddbn1] (ddbn2) {\(\partial^2 b_j^{(n - 2)}\)};
+    \node [box, below of=dddbn1] (dddbn2) {\(\partial^3 b_j^{(n - 2)}\)};
+    %%
+    \path [line] (bn0) -- node[above] {\(w_1^{(n)}\)} (dbn0);
+    \path [line] (dbn0) -- node[above] {\(w_2^{(n)}\)} (ddbn0);
+    \path [line] (ddbn0) -- node[above] {\(w_3^{(n)}\)} (dddbn0);
+    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
+    \path [line] (dddbn0) -- node[above] {\(w_4^{(n)}\)} (right0);
+    %%
+    \path [line] (bn1) -- node[above] {\(w_1^{(n - 1)}\)} (dbn1);
+    \path [line] (dbn1) -- node[above] {\(w_2^{(n - 1)}\)} (ddbn1);
+    \path [line] (ddbn1) -- node[above] {\(w_3^{(n - 1)}\)} (dddbn1);
+    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
+    \path [line] (dddbn1) -- node[above] {\(w_4^{(n - 1)}\)} (right1);
+    %%
+    \path [line] (bn2) -- node[above] {\(w_1^{(n - 2)}\)} (dbn2);
+    \path [line] (dbn2) -- node[above] {\(w_2^{(n - 2)}\)} (ddbn2);
+    \path [line] (ddbn2) -- node[above] {\(w_3^{(n - 2)}\)} (dddbn2);
+    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
+    \path [line] (dddbn2) -- node[above] {\(w_4^{(n - 2)}\)} (right2);
+    %%
+    \path [line] (bj) -- (bn0);
+    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
+    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
+    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
+    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
+    %%
+    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
+    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
+    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
+    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
+    %%
+    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
+    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
+    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
+    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
+    %%
+    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
+    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
+    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
+    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
+\end{tikzpicture}
+\end{center}
+
+\subsection{Operation Counts}
 

### 2018-04-27 nox.py
Adding a do-nothing `nox -s make_images` session.
This just adds the `matplotlibrc` file needed and the
hard-coded dependencies for a pinned version of `matplotlib`.

@@ -23,6 +23,19 @@ import py.path
 
 
 NOX_DIR = os.path.abspath(os.path.dirname(__file__))
+SINGLE_INTERP = "python3.6"
+DEPS = {
+    "matplotlib": (
+        "cycler == 0.10.0",
+        "kiwisolver == 1.0.1",
+        "matplotlib == 2.2.2",
+        "numpy == 1.14.2",
+        "pyparsing == 2.2.0",
+        "python-dateutil == 2.7.2",
+        "pytz == 2018.4",
+        "six == 1.11.0",
+    )
+}
 
 
 def get_path(*names):
@@ -94,3 +107,12 @@ def flop_counts(session):
     env = {"PYTHONPATH": get_path("src")}
     compute_counts = get_path("scripts", "compute_counts.py")
     session.run("python", compute_counts, env=env)
+
+
+@nox.session
+def make_images(session):
+    session.interpreter = SINGLE_INTERP
+    # Install all dependencies.
+    session.install(*DEPS["matplotlib"])
+    # Run the script(s).
+    env = {"MATPLOTLIBRC": get_path("images")}

### 2018-04-28 doc/paper.tex
Added a hacky introduction and some placeholder sections.
This introduction is largely modeled after and borrowed from
the two primary papers (K-fold Horner and compensated de Casteljau).

Also turned all `\[...\]` math blocks into `equation` groups so
that they get a number and can be referenced.

@@
 \newcommand{\eps}{\varepsilon}
+\newcommand{\mach}{\mathbf{u}}
 
@@
 round-off error is passed on to first order errors, then to second order
-errors, and so on. After the computation has been ``filtered'' \(K\)-times
-via this process, the resulting output is as accurate as the de Casteljau
+errors, and so on. After the computation has been ``filtered'' \((K - 1)\)
+times via this process, the resulting output is as accurate as the de Casteljau
 algorithm performed in \(K\) times the working precision. Forward error
@@
 \tableofcontents
+
+\section{Introduction}
+
+In computer aided geometric design, polynomials are usually expressed in
+Bernstein form. Polynomials in this form are usually evaluated by the
+de Casteljau algorithm. This algorithm has a round-off error bound
+which grows only linearly with degree, even though the number of
+arithmetic operations grows quadratically. Nevertheless the de Casteljau
+algorithm returns results arbitrarily less accurate than the working
+precision \(\mach\) when evaluating \(p(s)\) is ill-conditioned.
+The relative accuracy of the computed
+evaluation with the de Casteljau algorithm (\texttt{DeCasteljau}) satisfies
+(\cite{Mainar1999}) the following a priori bound:
+\begin{equation}\label{de-casteljau-error}
+  \frac{\left|p(s) - \mathtt{DeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq
+  \cond{p, s} \times \mathcal{O}(\mach).
+\end{equation}
+In the right-hand side of this inequality, \(\mach\) is the computing
+precision and the condition number \(\cond{p, s} \geq 1\) only depends
+on \(s\) and the Bernstein coefficients of \(p\) --- its expression will
+be given further.
+
+For ill-conditioned problems, such as evaluating \(p(s)\) near a
+multiple root, the condition number may be arbitrarily large, i.e.
+\(\cond{p, s} > 1 / \mach\), in
+which case most or all of the computed digits will be incorrect.
+In some cases, even the order of magnitude of the computed value
+of \(p(s)\) can be incorrect.
+
+To address ill-conditioned problems, error-free transformations can
+be applied in \textit{compensated algorithms} to account for roundoff.
+Error-free transformations were studied in great detail in \cite{Ogita2005}
+and open a large number of applications.
+In \cite{langlois_et_al:DSP:2006:442}, a compensated Horner's algorithm was
+described to evaluate a polynomial in the monomial basis. In \cite{Jiang2010},
+a similar method was described to perform a compensated version of the de
+Casteljau algorithm. In both cases, the \(\cond{p, s}\) factor is moved
+from \(\mach\) to \(\mach^2\) and the computed value is as accurate
+as if the computations were done in twice the working precision. For example,
+the compensated de Casteljau algorithm (\texttt{CompDeCasteljau}) satisfies
+\begin{equation}\label{de-casteljau-2-error}
+  \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
+    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
+    \mathcal{O}\left(\mach^2\right).
+\end{equation}
+For problems with \(\cond{p, s} < 1 / \mach^2\), the relative error
+is \(\mach\), i.e. accurate to full precision, aside from rounding to the
+nearest floating point number.
+
+In \cite{Graillat2009}, the authors built on the compensated Horner's
+algorithm to produce a method for evaluating a polynomial as if
+the computations were done in \(K\) times the working precision for
+any \(K \geq 2\). This result motivates this paper, though the
+approach there is somewhat different than ours. They perform each computation
+with error-free transformations and interpret the errors as coefficients of new
+polynomials. They then evaluate the error polynomials, which (recursively)
+generate second order error polynomials and so on. This recursive property
+causes the number of operations to grow exponentially in \(K\). Here, we
+instead have a fixed number of error groups, each corresponding to roundoff
+from the group above it. For example, when
+\((1 - s) b_j^{(n)} + s b_{j + 1}^{(n)}\) is computed in floating point, any
+error is filtered down to the error group below it.
+
+As in~\eqref{de-casteljau-error}, the accuracy of the compensated
+result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
+polynomial evaluations. We describe how to defer rounding into progressively
+smaller error groups and improve the accuracy of the computed result by a
+factor of \(\mach\) for every error group added. So we derive
+\texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
+that satisfies the following a prior bound for any arbitrary integer \(K\):
+\begin{equation}
+  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s)\right|}{
+    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
+    \mathcal{O}\left(\mach^K\right).
+\end{equation}
+This means that the computed value with \texttt{CompDeCasteljauK} is now
+as accurate as the result of the de Casteljau algorithm performed in
+\(K\) times the working precision with a final rounding back to this
+working precision.
+
+The paper is organized as follows. Section~\ref{notation} introduces some
+basic notations and results about floating point arithmetic, error-free
+transformations and the de Casteljau algorithm. In Section~\ref{compensated-2},
+the compensated algorithm for polynomial evaluation from \cite{Jiang2010} is
+reviewed and notation is established for the expansion. In
+Section~\ref{compensated-k}, the \(K\) compensated algorithm is provided and
+a forward error analysis is performed. Finally, in Section~\ref{numerical} we
+give numerical tests to illustrate the practical efficiency of our algorithms.
+
+\section{Basic notation and results}\label{notation}
+
+Placeholder
+
+\section{Compensated de Casteljau}\label{compensated-2}
+
+Placeholder
+
+\section{\texorpdfstring{\(K\)}{K} Compensated de Casteljau}\label{compensated-k}
+
+Placeholder
+
+\section{Numerical experiments}\label{numerical}
+
+Placeholder
 
@@
 (\cite{langlois_et_al:DSP:2006:442}):
-\[\cond{p(x)} = \frac{\widetilde{p}\left(\left|x\right|\right)}{
+\begin{equation}
+\cond{p(x)} = \frac{\widetilde{p}\left(\left|x\right|\right)}{
   \left|p(x)\right|} = \frac{\sum_j \left|a_j\right| \left|x\right|^j}{
-  \left|p(x)\right|}.\]
+  \left|p(x)\right|}.
+\end{equation}
 In particular, this means that if \(x \geq 0\) and each \(a_j \geq 0\)
@@
 we have \(\widetilde{p}(x) = (x + 1)^n\) hence
-\[\cond{p\left(1 + \delta\right)} = \frac{(2 + \delta)^n}{
-  \left|\delta\right|^n} = \left|1 + \frac{2}{\delta}\right|^n.\]
+\begin{equation}
+\cond{p\left(1 + \delta\right)} = \frac{(2 + \delta)^n}{
+  \left|\delta\right|^n} = \left|1 + \frac{2}{\delta}\right|^n.
+\end{equation}
 As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
@@
 For a polynomial \(p(s)\) in Bernstein form, we have (\cite{Jiang2010}):
-\[\cond{p(s)} = \frac{\widetilde{p}\left(s\right)}{
+\begin{equation}
+\cond{p(s)} = \frac{\widetilde{p}\left(s\right)}{
   \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| \left|B_{j, n}(s)\right|}{
-  \left|p(s)\right|}.\]
+  \left|p(s)\right|}.
+\end{equation}
 The Bernstein form is suited for \(s \in \left[0, 1\right]\), which means
@@
 example in use, consider
-\[p(s) = (1 - 2s)^n = \left[(1 - s) - s\right]^n = \sum_j \binom{n}{j}
-(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j B_{j, n}(s)\]
+\begin{equation}
+p(s) = (1 - 2s)^n = \left[(1 - s) - s\right]^n = \sum_j \binom{n}{j}
+(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j B_{j, n}(s)
+\end{equation}
 and input values of the form \(x = \frac{1}{2} + \delta\)
@@
 we have \(\widetilde{p}(s) = \left[(1 - s) + s\right]^n = 1\)
-\[\cond{p\left(\frac{1}{2} + \delta\right)} = \frac{1}{
-  \left|2\delta\right|^n}.\]
+\begin{equation}
+\cond{p\left(\frac{1}{2} + \delta\right)} = \frac{1}{
+  \left|2\delta\right|^n}.
+\end{equation}
 As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
@@
 Consider
-\[p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
-57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)\]
+\begin{equation}
+p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
+57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)
+\end{equation}
 at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
@@
 
-\subsection{\texorpdfstring{\(K\)}{K}-Fold Error Filtering}
+\subsection{\texorpdfstring{\(K\)}{K} Error Filtering}
 
@@
 relevant floating point operations, the \(K\)-fold Horner requires
-\[(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right) =
-\mathcal{O}\left((n + K)2^K\right)\]
+\begin{equation}
+(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right) =
+\mathcal{O}\left((n + K)2^K\right)
+\end{equation}
 flops to evaluate a degree \(n\) polynomial (this only applies when
@@
 floating point operations, the \(K\)-fold de Casteljau requires
-\[(15K^2 - 34K + 26)T_n + K + 5 =
-\mathcal{O}\left(n^2 K^2\right)\]
+\begin{equation}
+(15K^2 - 34K + 26)T_n + K + 5 =
+\mathcal{O}\left(n^2 K^2\right)
+\end{equation}
 flops to evaluate a degree \(n\) polynomial. (Here \(T_n\) is the

### 2018-04-28 doc/paper.tex
Filling out the notation section.
This is (for now) just a copy of the referenced sections.

@@
 \usepackage{booktabs,siunitx}
+%% H/T: https://tex.stackexchange.com/a/4217/32270
+\usepackage{mathtools}
+
+\theoremstyle{definition}
+\newtheorem{theorem}{Theorem}
 
@@
 
-To address ill-conditioned problems, error-free transformations can
+To address ill-conditioned problems, error-free transformations (EFT) can
 be applied in \textit{compensated algorithms} to account for roundoff.
@@
 The paper is organized as follows. Section~\ref{notation} introduces some
-basic notations and results about floating point arithmetic, error-free
+basic notation and results about floating point arithmetic, error-free
 transformations and the de Casteljau algorithm. In Section~\ref{compensated-2},
@@
 
-Placeholder
+Throughout this paper we assume that the computation in floating point
+arithmetic obeys the model
+\begin{equation}
+  a \star b = \fl{a \circ b} = (a \circ b)(1 + \eps_1) =
+  (a \circ b) / (1 + \eps_2)
+\end{equation}
+where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
+\in \left\{+, -, \times, \div\right\}\) and \(\left|\eps_1\right|,
+\left|\eps_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
+and \(\star\) represents the floating point computation, e.g.
+\(a \oplus b = \fl{a + b}\). We also assume that the computed result of
+\(\alpha \in \mathbf{R}\) in floating point arithmetic is denoted by
+\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
+all floating point numbers (see \cite{Higham2002} for more details).
+Following \cite{Higham2002}, we will use the following classic properties in
+error analysis.
+
+\begin{enumerate}
+  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
+      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
+  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
+      n \mach / (1 - n \mach)\),
+  \item \((1 + \theta_k)(1 + \theta_j) \leq (1 + \theta_{k + j})\),
+  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}\),
+  \item \((1 + \mach)^j \leq 1 / (1 - j \mach)\).
+\end{enumerate}
+
+Now let us introduce some results concerning error-free transformation (EFT).
+For a pair floating point numbers \(a, b \in \mathbf{F}\), there exists
+a floating point number \(y\) satisfying \(a \circ b = x + y\) where
+\(x = \fl{a \circ b}\) and \(\circ \in \left\{+, -, \times\right\}\).
+The transformation \((a, b) \longrightarrow (x, y)\) is regarded as an
+error-free transformation. The error-free transformation algorithms of the
+sum and product of two floating point numbers used later in this paper are
+the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
+algorithm by Dekker \cite{Dekker1971}, respectively. The following theorem
+exhibits the important properties of \texttt{TwoSum} and \texttt{TwoProd}
+algorithms (see \cite{Ogita2005}).
+
+\begin{theorem}[\cite{Ogita2005}]
+For \(a, b \in \mathbf{F}\) and \(x, y \in \mathbf{F}\), \texttt{TwoSum} and
+\texttt{TwoProd} satisfy
+\begin{itemize}
+  \item \(\left[x, y\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
+      \(x + y = a + b\), \(y \leq \mach \left|x\right|\),
+      \(y \leq \mach \left|a + b\right|\)
+  \item \(\left[x, y\right] = \mathtt{TwoProd}(a, b)\),
+      \(x = \fl{a \times b}\), \(x + y = a \times b\),
+      \(y \leq \mach \left|x\right|\),
+      \(y \leq \mach \left|a \times b\right|\).
+\end{itemize}
+\end{theorem}
+
+Next, we recall the de Casteljau algorithm:
+
+\begin{algorithm}[H]
+  \caption{\textit{de Casteljau algorithm for polynomial evaluation.}}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$\mathbf{b}, s$}
+      \State \(n = \texttt{length}(b) - 1\)
+      \State \(r = 1 \ominus s\)
+      \\
+      \For{\(j = 0, \ldots, n\)}
+        \State \(b_j^{(n)} = b_j\)
+      \EndFor
+      \\
+      \For{\(k = n - 1, \ldots, 0\)}
+        \For{\(j = 0, \ldots, k\)}
+          \State \(b_j^{(k)} = \left(r \otimes b_j^{(k + 1)}\right) \oplus
+              \left(s \otimes b_{j + 1}^{(k + 1)}\right)\)
+        \EndFor
+      \EndFor
+      \\
+      \State \(\mathtt{res} = b_0^{(0)}\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{theorem}[\cite{Mainar1999}]
+If \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) and \(\mathtt{DeCasteljau}(p, s)\)
+is the value computed by the de Casteljau algorithm then
+\begin{equation}
+\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{2n}
+\sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
+\end{equation}
+\end{theorem}
+
+The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
+b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
+\cite{Mainar1999}, \cite{Farouki1987}):
+\begin{equation}
+\cond{p, s} = \frac{\widetilde{p}\left(s\right)}{
+  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
+  \left|p(s)\right|},
+\end{equation}
+where \(B_{j, n} \geq 0\).
 
@@
       \State \(\pi = \mathtt{FMA}(a, b, -P)\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-  \caption{\textit{de Casteljau algorithm for polynomial evaluation.}}
-
-  \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$\mathbf{b}, s$}
-      \State \(n = \texttt{length}(b) - 1\)
-      \State \(r = 1 \ominus s\)
-      \\
-      \For{\(j = 0, \ldots, n\)}
-        \State \(b_j^{(n)} = b_j\)
-      \EndFor
-      \\
-      \For{\(k = n - 1, \ldots, 0\)}
-        \For{\(j = 0, \ldots, k\)}
-          \State \(b_j^{(k)} = \left(r \otimes b_j^{(k + 1)}\right) \oplus
-              \left(s \otimes b_{j + 1}^{(k + 1)}\right)\)
-        \EndFor
-      \EndFor
-      \\
-      \State \(\mathtt{res} = b_0^{(0)}\)
     \EndFunction

### 2018-04-28 doc/paper.tex
Adding write-up to section about existing algorithm.
Though I highly seasoned the original compensated de Casteljau
paper with my own notation.

@@
 \begin{theorem}[\cite{Ogita2005}]
-For \(a, b \in \mathbf{F}\) and \(x, y \in \mathbf{F}\), \texttt{TwoSum} and
-\texttt{TwoProd} satisfy
+For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
+\texttt{TwoSum} and \texttt{TwoProd} satisfy
 \begin{itemize}
-  \item \(\left[x, y\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
-      \(x + y = a + b\), \(y \leq \mach \left|x\right|\),
-      \(y \leq \mach \left|a + b\right|\)
-  \item \(\left[x, y\right] = \mathtt{TwoProd}(a, b)\),
-      \(x = \fl{a \times b}\), \(x + y = a \times b\),
-      \(y \leq \mach \left|x\right|\),
-      \(y \leq \mach \left|a \times b\right|\).
+  \item \(\left[S, \sigma\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
+      \(S + \sigma = a + b\), \(\sigma \leq \mach \left|S\right|\),
+      \(\sigma \leq \mach \left|a + b\right|\)
+  \item \(\left[P, \pi\right] = \mathtt{TwoProd}(a, b)\),
+      \(P = \fl{a \times b}\), \(P + \pi = a \times b\),
+      \(\pi \leq \mach \left|P\right|\),
+      \(\pi \leq \mach \left|a \times b\right|\).
 \end{itemize}
+The letters \(\sigma\) and \(\pi\) are used to indicate that the
+errors came from sum and product, respectively.
 \end{theorem}
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$\mathbf{b}, s$}
+    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$b, s$}
       \State \(n = \texttt{length}(b) - 1\)
-      \State \(r = 1 \ominus s\)
+      \State \(\widehat{r} = 1 \ominus s\)
       \\
       \For{\(j = 0, \ldots, n\)}
-        \State \(b_j^{(n)} = b_j\)
+        \State \(\widehat{b}_j^{(n)} = b_j\)
       \EndFor
@@
         \For{\(j = 0, \ldots, k\)}
-          \State \(b_j^{(k)} = \left(r \otimes b_j^{(k + 1)}\right) \oplus
-              \left(s \otimes b_{j + 1}^{(k + 1)}\right)\)
+          \State \(\widehat{b}_j^{(k)} = \left(
+              r \otimes \widehat{b}_j^{(k + 1)}\right) \oplus
+              \left(s \otimes \widehat{b}_{j + 1}^{(k + 1)}\right)\)
         \EndFor
@@
       \\
-      \State \(\mathtt{res} = b_0^{(0)}\)
+      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
     \EndFunction
@@
 
-Placeholder
+In this section we review the compensated de Casteljau algorithm
+from \cite{Jiang2010}. In order to track the local errors at
+each update step, we use four EFTs:
+\begin{align}
+\left[\widehat{r}, \rho\right] &= \mathtt{TwoSum}(1, -s) \\
+\left[P_1, \pi_1\right] &= \mathtt{TwoProd}\left(\widehat{r}, \widehat{b}_j^{(k + 1)}\right) \\
+\left[P_2, \pi_2\right] &= \mathtt{TwoProd}\left(s, \widehat{b}_{j + 1}^{(k + 1)}\right) \\
+\left[\widehat{b}_j^{(k)}, \sigma_3\right] &= \mathtt{TwoSum}(P_1, P_2)
+\end{align}
+With these, we can exactly describe the local error between the exact
+update and computed update:
+\begin{gather}
+\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \\
+(1 - s) \cdot \widehat{b}_j^{(k + 1)} +
+  s \cdot \widehat{b}_{j + 1}^{(k + 1)} =
+\widehat{b}_j^{(k)} + \ell_j^{(k)}.
+\end{gather}
+
+\begin{figure}[H]
+\centering
+\begin{tikzpicture}[node distance = 0.2\textwidth]
+  \node (b0) {\(\widehat{b}^{(n)}\)}
+    child {node (b1) {\(\widehat{b}^{(n - 1)}\)}
+      child {node (b2) {\(\widehat{b}^{(n - 2)}\)}
+        child {node (bdots) {\(\vdots\)}
+          child {node (bn) {\(\widehat{b}^{(0)}\)}}
+          child {node (ln) {\(\ell^{(0)}\)}}
+        }
+        child {node (ldots) {\(\vdots\)}}
+      }
+      child {node (l2) {\(\ell^{(n - 2)}\)}}
+    }
+    child {node (l1) {\(\ell^{(n - 1)}\)}};
+  %%
+  \path (b1) -- (l1) node [midway] {\(+\)};
+  \path (b2) -- (l2) node [midway] {\(+\)};
+  \path (bn) -- (ln) node [midway] {\(+\)};
+  %% eb == exact b
+  \node [right of=b0] (eb0) {\(b^{(n)}\)}
+    child {node (eb1) {\(b^{(n - 1)}\)}
+      child {node (eb2) {\(b^{(n - 2)}\)}
+        child {node (ebdots) {\(\vdots\)}
+          child {node (ebn) {\(b^{(0)}\)}}
+        }
+      }
+    };
+\end{tikzpicture}
+\caption{Local round-off errors}
+\end{figure}
+
+\noindent By defining the global errors at each step
+\begin{equation}
+  \partial b_j^{(k)} = b_j^{(k)} - \widehat{b}_j^{(k)}
+\end{equation}
+we can see that the local errors accumulate in
+\(\partial b^{(k)}\):
+\begin{equation}
+  \partial b_j^{(k)} = (1 - s) \cdot \partial b_j^{(k + 1)} + s \cdot
+  \partial b_{j + 1}^{(k + 1)} + \ell_j^{(k)}.
+\end{equation}
+Hence, when computed in exact arithmetic, we have
+\begin{equation}
+  p(s) = \widehat{b}_0^{(0)} + \partial b_0^{(0)}.
+\end{equation}
+
+The idea behind the compensated de Casteljau algorithm
+is to compute both the local error and the updates of the global
+error with floating point operations:
+
+\begin{algorithm}[H]
+  \caption{\textit{Compensated de Casteljau algorithm for polynomial evaluation.}}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$b, s$}
+      \State \(n = \texttt{length}(b) - 1\)
+      \State \(\left[\widehat{r}, \rho\right] = \mathtt{TwoSum}(1, -s)\)
+      \\
+      \For{\(j = 0, \ldots, n\)}
+        \State \(\widehat{b}_j^{(n)} = b_j\)
+        \State \(\widehat{\partial b}_j^{(n)} = 0\)
+      \EndFor
+      \\
+      \For{\(k = n - 1, \ldots, 0\)}
+        \For{\(j = 0, \ldots, k\)}
+          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
+              \widehat{r}, \widehat{b}_j^{(k + 1)}\right)\)
+          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
+              s, \widehat{b}_{j + 1}^{(k + 1)}\right)\)
+          \State \(\left[\widehat{b}_j^{(k)}, \sigma_3\right] =
+              \mathtt{TwoSum}(P_1, P_2)\)
+          \State \(\widehat{\ell}_j^{(k)} = \pi_1 \oplus \pi_2 \oplus \sigma_3
+              \oplus \left(\rho \otimes \widehat{b}_j^{(k + 1)}\right)\)
+          \State \(\widehat{\partial b}_j^{(k)} =
+              \left(\widehat{r} \otimes
+              \widehat{\partial b}_j^{(k + 1)}\right) \oplus
+              \left(s \otimes \widehat{\partial b}_{j + 1}^{(k + 1)}
+              \right) \oplus
+              \widehat{\ell}_j^{(k)}\)
+        \EndFor
+      \EndFor
+      \\
+      \State \(\mathtt{res} = \widehat{b}_0^{(0)} \oplus
+          \widehat{\partial b}_0^{(0)}\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent  When comparing this computed error to the exact error, the
+difference depends only on \(s\) and the Bernstein
+coefficients of \(p\):
+
+\begin{theorem}[\cite{Jiang2010}]
+  If no underflow occurs and \(n \geq 2\), then
+  \begin{equation}
+    \left|\partial b_0^{(0)} - \widehat{\partial b}_0^{(0)}\right| \leq
+    2 \gamma_{3n + 1} \gamma_{3(n - 1)}
+    \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
+  \end{equation}
+\end{theorem}
+
+\noindent Using the bound on the error in \(\partial b\), the algorithm can
+be shown to be as accurate as if the computations were done in twice
+the working precision:
+
+\begin{theorem}[\cite{Jiang2010}]
+  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
+  \begin{equation}
+    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq \mach +
+    2 \gamma_{3n}^2 \cond{p, s}.
+  \end{equation}
+\end{theorem}
 
@@
 
-\begin{algorithm}[H]
-  \caption{\textit{Compensated de Casteljau algorithm for polynomial evaluation.}}
-
-  \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$\mathbf{b}, s$}
-      \State \(n = \texttt{length}(\mathbf{b}) - 1\)
-      \State \(\left[r, \rho\right] = \mathtt{TwoSum}(1, -s)\)
-      \\
-      \For{\(j = 0, \ldots, n\)}
-        \State \(b_j^{(n)} = b_j\)
-        \State \(\partial b_j^{(n)} = 0\)
-      \EndFor
-      \\
-      \For{\(k = n - 1, \ldots, 0\)}
-        \For{\(j = 0, \ldots, k\)}
-          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
-              r, b_j^{(k + 1)}\right)\)
-          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
-              s, b_{j + 1}^{(k + 1)}\right)\)
-          \State \(\left[b_j^{(k)}, \sigma_3\right] = \mathtt{TwoSum}(
-              P_1, P_2)\)
-          \State \(w = \pi_1 \oplus \pi_2 \oplus \sigma_3 \oplus
-              \left(\rho \otimes b_j^{(k + 1)}\right)\)
-          \State \(\partial b_j^{(k)} = w \oplus \left(s \otimes
-              \partial b_{j + 1}^{(k + 1)}
-              \right) \oplus \left(r \otimes \partial b_j^{(k + 1)}\right)\)
-        \EndFor
-      \EndFor
-      \\
-      \State \(\mathtt{res} = b_0^{(0)} \oplus \partial b_0^{(0)}\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
 \section{de Casteljau's Method}
@@
 polynomial in Bernstein-B\'{e}zier form with control points \(b_j\):
-\begin{align*}
+\begin{align}
     b_j^{(n)} &= b_j \\
@@
     b(s) &= b_0^{(0)}.
-\end{align*}
+\end{align}
 

### 2018-04-28 doc/paper.tex
Adding beginning of description of K-fold algorithm.

@@
 
-Placeholder
+In order to raise from twice the working precision to \(K\) times the
+working precision, we continue using EFTs when computing
+\(\widehat{\partial b}^{(k)}\). By tracking the round-off from each
+floating point evaluation via an EFT, we can form a cascade of global errors:
+\begin{align}
+  b_j^{(k)} &= \widehat{b}_j^{(k)} + \partial b_j^{(k)} \\
+  \partial b_j^{(k)} &= \widehat{\partial b}_j^{(k)} + \partial^2 b_j^{(k)} \\
+  \partial^2 b_j^{(k)} &= \widehat{\partial^2 b}_j^{(k)} +
+  \partial^3 b_j^{(k)} \\
+  & \vdots \nonumber
+\end{align}
+In the same way local error can be tracked when updating
+\(\widehat{b}_j^{(k)}\), it can be tracked for updates that happen down
+the cascade:
+\begin{alignat}{4}
+  (1 - s) \cdot \widehat{b}_j^{(k + 1)} &+
+  s \cdot \widehat{b}_{j + 1}^{(k + 1)} &&  &&=
+  \widehat{b}_j^{(k)} &&+ \ell_{1, j}^{(k)} \\
+  (1 - s) \cdot \widehat{\partial b}_j^{(k + 1)} &+
+  s \cdot \widehat{\partial b}_{j + 1}^{(k + 1)} &&+ \ell_{1, j}^{(k)} &&=
+  \widehat{\partial b}_j^{(k)} &&+ \ell_{2, j}^{(k)} \\
+  (1 - s) \cdot \widehat{\partial^2 b}_j^{(k + 1)} &+
+  s \cdot \widehat{\partial^2 b}_{j + 1}^{(k + 1)} &&+ \ell_{2, j}^{(k)} &&=
+  \widehat{\partial^2 b}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
+  &  &&  &&\vdots && \nonumber
+\end{alignat}
+
+In \texttt{CompDeCasteljau}, after a single stage of error filtering we
+``give up'' and use \(\widehat{\partial b}^{(k)}\) instead of
+\(\partial b^{(k)}\) (and don't keep around any information about the
+round-off error). In order to obtain results that are as accurate as if
+computed in \(K\) times the working precision, we must continue filtering
+errors down \((K - 1)\) times, so that at the final level we compute
+\(\widehat{\partial^{K - 1} b}^{(k)}\) and don't worry about round-off
+error.
+
+\begin{figure}[H]
+\centering
+\begin{tikzpicture}[node distance = 0.2\textwidth]
+    \node [box] (bj) {\(b\)};
+    \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
+    \node [box, right of=bn0] (dbn0) {\(\widehat{\partial b}^{(n)}\)};
+    \node [box, right of=dbn0] (ddbn0) {\(\widehat{\partial^2 b}^{(n)}\)};
+    \node [box, right of=ddbn0] (dddbn0) {\(\widehat{\partial^3 b}^{(n)}\)};
+    %%
+    \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
+    \node [box, below of=dbn0] (dbn1) {\(\widehat{\partial b}^{(n - 1)}\)};
+    \node [box, below of=ddbn0] (ddbn1) {\(\widehat{\partial^2 b}^{(n - 1)}\)};
+    \node [box, below of=dddbn0] (dddbn1) {\(\widehat{\partial^3 b}^{(n - 1)}\)};
+    %%
+    \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
+    \node [box, below of=dbn1] (dbn2) {\(\widehat{\partial b}^{(n - 2)}\)};
+    \node [box, below of=ddbn1] (ddbn2) {\(\widehat{\partial^2 b}^{(n - 2)}\)};
+    \node [box, below of=dddbn1] (dddbn2) {\(\widehat{\partial^3 b}^{(n - 2)}\)};
+    %%
+    \path [line] (bn0) -- node[above] {\(\widehat{\ell}_1^{(n)}\)} (dbn0);
+    \path [line] (dbn0) -- node[above] {\(\widehat{\ell}_2^{(n)}\)} (ddbn0);
+    \path [line] (ddbn0) -- node[above] {\(\widehat{\ell}_3^{(n)}\)} (dddbn0);
+    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
+    \path [line] (dddbn0) -- node[above] {\(\widehat{\ell}_4^{(n)}\)} (right0);
+    %%
+    \path [line] (bn1) -- node[above] {\(\widehat{\ell}_1^{(n - 1)}\)} (dbn1);
+    \path [line] (dbn1) -- node[above] {\(\widehat{\ell}_2^{(n - 1)}\)} (ddbn1);
+    \path [line] (ddbn1) -- node[above] {\(\widehat{\ell}_3^{(n - 1)}\)} (dddbn1);
+    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
+    \path [line] (dddbn1) -- node[above] {\(\widehat{\ell}_4^{(n - 1)}\)} (right1);
+    %%
+    \path [line] (bn2) -- node[above] {\(\widehat{\ell}_1^{(n - 2)}\)} (dbn2);
+    \path [line] (dbn2) -- node[above] {\(\widehat{\ell}_2^{(n - 2)}\)} (ddbn2);
+    \path [line] (ddbn2) -- node[above] {\(\widehat{\ell}_3^{(n - 2)}\)} (dddbn2);
+    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
+    \path [line] (dddbn2) -- node[above] {\(\widehat{\ell}_4^{(n - 2)}\)} (right2);
+    %%
+    \path [line] (bj) -- (bn0);
+    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
+    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
+    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
+    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
+    %%
+    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
+    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
+    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
+    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
+    %%
+    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
+    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
+    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
+    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
+    %%
+    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
+    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
+    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
+    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
+\end{tikzpicture}
+\caption{Filtering errors}
+\end{figure}
 
@@
 
-\subsection{\texorpdfstring{\(K\)}{K} Error Filtering}
-
-\begin{center}
-\begin{tikzpicture}[node distance = 0.2\textwidth]
-    \node [box] (bj) {\(b_j\)};
-    \node [box, below of=bj] (bn0) {\(b_j^{(n)}\)};
-    \node [box, right of=bn0] (dbn0) {\(\partial b_j^{(n)}\)};
-    \node [box, right of=dbn0] (ddbn0) {\(\partial^2 b_j^{(n)}\)};
-    \node [box, right of=ddbn0] (dddbn0) {\(\partial^3 b_j^{(n)}\)};
-    %%
-    \node [box, below of=bn0] (bn1) {\(b_j^{(n - 1)}\)};
-    \node [box, below of=dbn0] (dbn1) {\(\partial b_j^{(n - 1)}\)};
-    \node [box, below of=ddbn0] (ddbn1) {\(\partial^2 b_j^{(n - 1)}\)};
-    \node [box, below of=dddbn0] (dddbn1) {\(\partial^3 b_j^{(n - 1)}\)};
-    %%
-    \node [box, below of=bn1] (bn2) {\(b_j^{(n - 2)}\)};
-    \node [box, below of=dbn1] (dbn2) {\(\partial b_j^{(n - 2)}\)};
-    \node [box, below of=ddbn1] (ddbn2) {\(\partial^2 b_j^{(n - 2)}\)};
-    \node [box, below of=dddbn1] (dddbn2) {\(\partial^3 b_j^{(n - 2)}\)};
-    %%
-    \path [line] (bn0) -- node[above] {\(w_1^{(n)}\)} (dbn0);
-    \path [line] (dbn0) -- node[above] {\(w_2^{(n)}\)} (ddbn0);
-    \path [line] (ddbn0) -- node[above] {\(w_3^{(n)}\)} (dddbn0);
-    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
-    \path [line] (dddbn0) -- node[above] {\(w_4^{(n)}\)} (right0);
-    %%
-    \path [line] (bn1) -- node[above] {\(w_1^{(n - 1)}\)} (dbn1);
-    \path [line] (dbn1) -- node[above] {\(w_2^{(n - 1)}\)} (ddbn1);
-    \path [line] (ddbn1) -- node[above] {\(w_3^{(n - 1)}\)} (dddbn1);
-    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
-    \path [line] (dddbn1) -- node[above] {\(w_4^{(n - 1)}\)} (right1);
-    %%
-    \path [line] (bn2) -- node[above] {\(w_1^{(n - 2)}\)} (dbn2);
-    \path [line] (dbn2) -- node[above] {\(w_2^{(n - 2)}\)} (ddbn2);
-    \path [line] (ddbn2) -- node[above] {\(w_3^{(n - 2)}\)} (dddbn2);
-    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
-    \path [line] (dddbn2) -- node[above] {\(w_4^{(n - 2)}\)} (right2);
-    %%
-    \path [line] (bj) -- (bn0);
-    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
-    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
-    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
-    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
-    %%
-    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
-    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
-    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
-    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
-    %%
-    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
-    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
-    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
-    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
-    %%
-    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
-    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
-    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
-    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
-\end{tikzpicture}
-\end{center}
-
 \subsection{Operation Counts}

### 2018-04-29 doc/paper.tex
Adding explanation of how K-compensated de Casteljau works.
The description is very raw and needs cleaning, but this is
a start.

Also re-formatted the "Example of Compensated de Casteljau Failing"
section to use `\mach` instead of `u` and to put a hat over the
computed values.

@@
 \begin{gather}
-\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \\
+\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \label{ell-j} \\
 (1 - s) \cdot \widehat{b}_j^{(k + 1)} +
@@
 ``give up'' and use \(\widehat{\partial b}^{(k)}\) instead of
-\(\partial b^{(k)}\) (and don't keep around any information about the
+\(\partial b^{(k)}\) (without keeping around any information about the
 round-off error). In order to obtain results that are as accurate as if
@@
 error.
+
+When computing \(\widehat{\partial^F b}\) (i.e. the error after
+\(F\) stages of filtering)
+there will be several sources of round-off. In particular, there will be
+\begin{itemize}
+\item errors when computing \(\widehat{\ell}_{F, j}^{(k)}\) from the
+  terms in \(\ell_{F, j}^{(k)}\)
+\item an error
+for the ``missing'' \(\rho \cdot \widehat{\partial^F b}_j^{(k + 1)}\) in
+\((1 - s) \cdot \widehat{\partial^F b}_j^{(k + 1)}\)
+\item an error from the product
+  \(\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\)
+\item an error from the product
+  \(s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\)
+\item two errors from the two \(\oplus\) when combining the three
+  terms in
+  \(\left[\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\right] \oplus
+  \left[s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right] \oplus
+  \widehat{\ell}_{F, j}^{(k)}\)
+\end{itemize}
+For example, in~\eqref{ell-j}:
+\begin{equation}
+\ell_{1, j}^{(k)} = \underbrace{\pi_1}_{P_1 = \widehat{r} \otimes
+    \widehat{b}_j^{(k + 1)}} +
+\underbrace{\pi_2}_{P_2 = s \otimes \widehat{b}_{j + 1}^{(k + 1)}} +
+\underbrace{\sigma_3}_{P_1 \oplus P_2} +
+\underbrace{\rho \cdot \widehat{b}_j^{(k + 1)}}_{(1 - s)
+  \widehat{b}_j^{(k + 1)}}
+\end{equation}
+After each stage, we'll always have
+\[\ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
+\widehat{\partial^{F - 1} b}_j^{(k + 1)}\]
+where the terms \(e_1, \ldots, e_M\) come from using \texttt{TwoSum} and
+\texttt{TwoProd} when computing \(\widehat{\partial^{F - 1} b}_j^{(k)}\)
+and the \(\rho\) term comes from the roundoff
+in \(1 \ominus s\) when multiplying \((1 - s)\) by
+\(\widehat{\partial^{F - 1} b}_j^{(k + 1)}\). With this in mind, we
+can define an EFT (\texttt{LocalErrorEFT}) that computes
+\(\widehat{\ell}\) and tracks all round-off errors generated in
+the process:
+
+\begin{algorithm}[H]
+  \caption{\textit{EFT for computing the local error.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[\eta, \widehat{\ell}\right] =
+        \mathtt{LocalErrorEFT}\)}{$e, \rho, \delta b$}
+      \State \(M = \texttt{length}(e)\)
+      \\
+      \State \(\left[\widehat{\ell}, \eta_1\right] =
+          \mathtt{TwoSum}(e_1, e_2)\)
+      \For{\(j = 3, \ldots, M\)}
+        \State \(\left[\widehat{\ell}, \eta_{j - 1}\right] =
+            \mathtt{TwoSum}\left(\widehat{\ell}, e_j\right)\)
+      \EndFor
+      \\
+      \State \(\left[P, \eta_M\right] =
+          \mathtt{TwoProd}\left(\rho, \delta b\right)\)
+      \State \(\left[\widehat{\ell}, \eta_{M + 1}\right] =
+          \mathtt{TwoSum}\left(\widehat{\ell}, P\right)\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent With this EFT helper in place, we can perform \((K - 1)\)
+error filtrations:
+
+\begin{algorithm}[H]
+  \caption{\(K\)-\textit{compensated de Casteljau algorithm.}}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$b, s$}
+      \State \(n = \texttt{length}(b) - 1\)
+      \State \(\left[\widehat{r}, \rho\right] = \mathtt{TwoSum}(1, -s)\)
+      \\
+      \For{\(j = 0, \ldots, n\)}
+        \State \(\widehat{b}_j^{(n)} = b_j\)
+        \For{\(F = 1, \ldots, K - 1\)}
+          \State \(\widehat{\partial^F b}_j^{(n)} = 0\)
+        \EndFor
+      \EndFor
+      \\
+      \For{\(k = n - 1, \ldots, 0\)}
+        \For{\(j = 0, \ldots, k\)}
+          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
+              \widehat{r}, \widehat{b}_j^{(k + 1)}\right)\)
+          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
+              s, \widehat{b}_{j + 1}^{(k + 1)}\right)\)
+          \State \(\left[\widehat{b}_j^{(k)}, \sigma_3\right] =
+              \mathtt{TwoSum}(P_1, P_2)\)
+          \\
+          \State \(e = \left[\pi_1, \pi_2, \sigma_3\right]\)
+          \State \(\delta b = \widehat{b}_j^{(k + 1)}\)
+          \\
+          \For{\(F = 1, \ldots, K - 2\)}
+            \State \(\left[\eta, \widehat{\ell}\right] =
+                \mathtt{LocalErrorEFT}(e, \rho, \delta b)\)
+            \State \(M = \texttt{length}(\eta)\)
+            \\
+            \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
+                \widehat{r}, \widehat{\partial^F b}_j^{(k + 1)}\right)\)
+            \State \(\left[P_2, \eta_{M + 2}\right] = \mathtt{TwoProd}\left(
+                s, \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right)\)
+            \State \(\left[S_3, \eta_{M + 3}\right] =
+                \mathtt{TwoSum}(P_1, P_2)\)
+            \State \(\left[\widehat{\partial^F b}_j^{(k)}, \eta_{M + 4}\right]
+                = \mathtt{TwoSum}\left(S_3, \widehat{\ell}\right)\)
+            \\
+            \State \(e = \eta\)
+            \State \(\delta b = \widehat{\partial^F b}_j^{(k + 1)}\)
+          \EndFor
+          \\
+          \State \(\widehat{\ell} =
+                \mathtt{LocalError}(e, \rho, \delta b)\)
+          \State \(\widehat{\partial^{K - 1} b}_j^{(k)} =
+              \left(\widehat{r} \otimes
+              \widehat{\partial^{K - 1} b}_j^{(k + 1)}\right) \oplus
+              \left(s \otimes \widehat{\partial^{K - 1} b}_{j + 1}^{(k + 1)}
+              \right) \oplus
+              \widehat{\ell}\)
+        \EndFor
+      \EndFor
+      \\
+      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
+      \For{\(F = 1, \ldots, K - 1\)}
+        \State \(\mathtt{res} = \mathtt{res} \oplus
+            \widehat{\partial^F b}_0^{(0)}\)
+      \EndFor
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
 
@@
 \end{equation}
-at the point \(s = \frac{3}{4} + 800 u\). In exact arithmetic, we
-have \(p(s) = 13(3200u)^3 + \mathcal{O}\left(u^4\right)\).
+at the point \(s = \frac{3}{4} + 800 \mach\). In exact arithmetic, we
+have \(p(s) = 13(3200\mach)^3 + \mathcal{O}\left(\mach^4\right)\).
 However, using the compensated de Casteljau algorithm results in
-\(b_0^{(0)} + \partial b_0^{(0)} = 0\):
+\(\widehat{b}_0^{(0)} \oplus \widehat{\partial b}_0^{(0)} = 0\):
 
@@
     \toprule
-    k & j & b_j^{(k)} & \partial b_j^{(k)} & \partial^2 b_j^{(k)} & \partial^3 b_j^{(k)} \\
+    k & j & \widehat{b}_j^{(k)} & \widehat{\partial b}_j^{(k)} & \widehat{\partial^2 b}_j^{(k)} & \widehat{\partial^3 b}_j^{(k)} \\
     \midrule
-    3 & 0 & -87\frac{3}{4} + 108032u & -32u & 0 & 0 \\
-    3 & 1 & 29\frac{1}{4} + 88768u & 32u & 0 & 0 \\
-    3 & 2 & -9\frac{3}{4} - 71200u & 0 & 0 & 0 \\
-    3 & 3 & 3\frac{1}{4} + 37600u & 0 & 0 & 0 \\
+    3 & 0 & -87\frac{3}{4} + 108032\mach & -32\mach & 0 & 0 \\
+    3 & 1 & 29\frac{1}{4} + 88768\mach & 32\mach & 0 & 0 \\
+    3 & 2 & -9\frac{3}{4} - 71200\mach & 0 & 0 & 0 \\
+    3 & 3 & 3\frac{1}{4} + 37600\mach & 0 & 0 & 0 \\
     \midrule
-    2 & 0 & 187200u & -15360000u^2 & 0 & 0 \\
-    2 & 1 & -62408u & 8u - 128000000u^2 & 0 & 0 \\
-    2 & 2 & 20800u & 87040000u^2 & 0 & 0 \\
+    2 & 0 & 187200\mach & -15360000\mach^2 & 0 & 0 \\
+    2 & 1 & -62408\mach & 8\mach - 128000000\mach^2 & 0 & 0 \\
+    2 & 2 & 20800\mach & 87040000\mach^2 & 0 & 0 \\
     \midrule
-    1 & 0 & -6u - 199688192u^2 & 6u - 99831808u^2 & -90112000000u^3 & 0 \\
-    1 & 1 & -2u + 66568192u^2  & 2u + 33271808u^2 & 172032000000u^3 & 0 \\
+    1 & 0 & -6\mach - 199688192\mach^2 & 6\mach - 99831808\mach^2 & -90112000000\mach^3 & 0 \\
+    1 & 1 & -2\mach + 66568192\mach^2  & 2\mach + 33271808\mach^2 & 172032000000\mach^3 & 0 \\
     \midrule
-    0 & 0 & -3u + 7296u^2 & 3u - 7296u^2 & 13 (3200u)^3 + 3048(512u)^4 & 962(128u)^4 \\
+    0 & 0 & -3\mach + 7296\mach^2 & 3\mach - 7296\mach^2 & 13 (3200\mach)^3 + 3048(512\mach)^4 & 962(128\mach)^4 \\
     \bottomrule

### 2018-04-29 scripts/error_against_cond.py
Implementing the "fully generic" ``_compensated_k`` for de Casteljau.
I have used `r bj + s b{j+1} + ell` rather than the
`ell + s b{j+1} + r bj` that was used before, so the values
themselves have changed slightly.

@@ -79,25 +79,25 @@ def main(filename=None):
         cond_nums.append(float(exact_cond))
 
         # Compute the forward error for uncompensated de Casteljau.
-        b1, db, ddb, dddb = de_casteljau._compensated4(s, BEZIER_COEFFS)
-        exact_b1 = F(b1)
+        b, db, d2b, d3b = de_casteljau._compensated_k(s, BEZIER_COEFFS, 4)
+        exact_b1 = F(b)
         exact_forward_err1 = abs((exact_b1 - exact_p) / exact_p)
         forward_errs1.append(float(exact_forward_err1))
 
         # Compute the forward error for compensated de Casteljau.
-        b2 = b1 + db
+        b2 = b + db
         exact_b2 = F(b2)
         exact_forward_err2 = abs((exact_b2 - exact_p) / exact_p)
         forward_errs2.append(float(exact_forward_err2))
 
         # Compute the forward error for K-compensated de Casteljau (K=3).
-        b3 = b2 + ddb
+        b3 = b2 + d2b
         exact_b3 = F(b3)
         exact_forward_err3 = abs((exact_b3 - exact_p) / exact_p)
         forward_errs3.append(float(exact_forward_err3))
 
         # Compute the forward error for K-compensated de Casteljau (K=3).
-        b4 = b3 + dddb
+        b4 = b3 + d3b
         exact_b4 = F(b4)
         exact_forward_err4 = abs((exact_b4 - exact_p) / exact_p)
         forward_errs4.append(float(exact_forward_err4))

### 2018-04-29 src/de_casteljau.py

@@ -56,276 +56,104 @@ def basic(s, coeffs):
     return pk[0]
 
 
-def _compensated(s, coeffs):
+def local_error(errors, rho, delta_b):
+    r"""Compute :math:`\ell` from a list of errors.
+
+    This assumes, but does not check, that there are at least two
+    ``errors``.
+    """
+    num_errs = len(errors)
+
+    l_hat = errors[0] + errors[1]
+    for j in range(2, num_errs):
+        l_hat += errors[j]
+
+    l_hat += rho * delta_b
+
+    return l_hat
+
+
+def local_error_eft(errors, rho, delta_b):
+    r"""Perform an error-free transformation for computing :math:`\ell`.
+
+    This assumes, but does not check, that there are at least two
+    ``errors``.
+    """
+    num_errs = len(errors)
+    new_errors = [None] * (num_errs + 1)
+
+    l_hat, new_errors[0] = eft.add_eft(errors[0], errors[1])
+    for j in range(2, num_errs):
+        l_hat, new_errors[j - 1] = eft.add_eft(l_hat, errors[j])
+
+    prod, new_errors[num_errs - 1] = eft.multiply_eft(rho, delta_b)
+    l_hat, new_errors[num_errs] = eft.add_eft(l_hat, prod)
+
+    return new_errors, l_hat
+
+
+def _compensated_k(s, coeffs, K):
     r, rho = eft.add_eft(1.0, -s)
 
     degree = len(coeffs) - 1
-    pk = list(coeffs)
-    eAk = [0.0] * (degree + 1)
+    bk = {0: list(coeffs)}
+    # NOTE: This will be shared, but is read only.
+    all_zero = (0.0,) * (degree + 1)
+    for F in range(1, K - 1 + 1):
+        bk[F] = all_zero
+
     for k in range(degree):
-        new_pk = []
-        new_eAk = []
+        new_bk = {F: [] for F in range(K - 1 + 1)}
 
         for j in range(degree - k):
-            P1, pi1 = eft.multiply_eft(r, pk[j])
-            P2, pi2 = eft.multiply_eft(s, pk[j + 1])
+            # Update the "level 0" stuff.
+            P1, pi1 = eft.multiply_eft(r, bk[0][j])
+            P2, pi2 = eft.multiply_eft(s, bk[0][j + 1])
             S3, sigma3 = eft.add_eft(P1, P2)
-            new_pk.append(S3)
-            # Now update the error.
[…]

### 2018-04-30 doc/paper.tex
Adding the error plot to the main document.
Removing the references that are actually used from "bogus" section.
Adding the original JLCS10 plot to introduction.
Also

- Changing `linkcolor` to `MidnightBlue` (the default `red`
  was too foreboding)
- Changing from `darkgrid` to `white` theme for the plots
  (this results in a lot less going on)
- Adding `de_casteljau.jlcs10_compensated` as a distinct
  implementation. The key difference being the order of operations
  in computing `r b1 + s b2 + l_hat`.
- Updating section labels from `{foo}` to `{sec:foo}`

@@
   urlcolor=blue,
+  linkcolor=MidnightBlue,
   citecolor=ForestGreen,
@@
 is \(\mach\), i.e. accurate to full precision, aside from rounding to the
-nearest floating point number.
+nearest floating point number. Figure~\ref{fig:jlcs-10} shows this shift
+in relative error from \texttt{DeCasteljau} to \texttt{CompDeCasteljau}.
+
+\begin{figure}
+  \includegraphics[width=0.8125\textwidth]{../images/jlcs10_plot.pdf}
+  \centering
+  \caption{Evaluation of \(p(s) = (s - 1)\left(s - 3/4\right)^7\)
+    represented in Bernstein form.}
+  \label{fig:jlcs-10}
+\end{figure}
 
@@
 result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
-polynomial evaluations. We describe how to defer rounding into progressively
+polynomial evaluations. For example, as the condition number grows in
+Figure~\ref{fig:jlcs-10}, some points have relative error exactly equal to
+\(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
+a complete failure to evaluate the order of magnitude of \(p(s)\).
+We describe how to defer rounding into progressively
 smaller error groups and improve the accuracy of the computed result by a
@@
 
-The paper is organized as follows. Section~\ref{notation} introduces some
+The paper is organized as follows. Section~\ref{sec:notation} introduces some
 basic notation and results about floating point arithmetic, error-free
-transformations and the de Casteljau algorithm. In Section~\ref{compensated-2},
+transformations and the de Casteljau algorithm. In
+Section~\ref{sec:compensated-2},
 the compensated algorithm for polynomial evaluation from \cite{Jiang2010} is
 reviewed and notation is established for the expansion. In
-Section~\ref{compensated-k}, the \(K\) compensated algorithm is provided and
-a forward error analysis is performed. Finally, in Section~\ref{numerical} we
+Section~\ref{sec:compensated-k}, the \(K\) compensated algorithm is provided and
+a forward error analysis is performed. Finally, in Section~\ref{sec:numerical} we
 give numerical tests to illustrate the practical efficiency of our algorithms.
 
-\section{Basic notation and results}\label{notation}
+\section{Basic notation and results}\label{sec:notation}
 
@@
 
-\section{Compensated de Casteljau}\label{compensated-2}
+\section{Compensated de Casteljau}\label{sec:compensated-2}
 
@@
 
-\section{\texorpdfstring{\(K\)}{K} Compensated de Casteljau}\label{compensated-k}
+\section{\texorpdfstring{\(K\)}{K} Compensated de Casteljau}\label{sec:compensated-k}
 
@@
 
-\section{Numerical experiments}\label{numerical}
+\section{Numerical experiments}\label{sec:numerical}
 
-Placeholder
+\begin{figure}[H]
+  \includegraphics[width=0.8125\textwidth]{../images/de_casteljau_rel_error.pdf}
+  \centering
+  \caption{Accuracy of evaluation of \(p(s) = (s - 1)\left(s - 3/4\right)^7\)
+    represented in Bernstein form.}
+\end{figure}
 
@@
 \begin{itemize}
-  \item Compensated Horner (\(K = 2\)) (\cite{langlois_et_al:DSP:2006:442})
-  \item Compensated de Casteljau (\cite{Jiang2010})
   \item Newton with compensated Horner (\cite{Graillat2008})
-  \item \(K\)-fold Sum (\cite{Ogita2005})
-  \item \(K\)-fold Horner (\cite{Graillat2009})
 \end{itemize}

### 2018-04-30 scripts/error_against_cond.py

@@ -191,7 +191,7 @@ def main(filename=None):
     # Set "nice" ticks.
     ax.set_xticks([10.0 ** n for n in range(5, 65 + 10, 10)])
     ax.set_yticks([10.0 ** n for n in range(-18, 0 + 2, 2)])
-    # Set special ``xticks`` for ``1/u``, ``1/u^2`` and ``1/u^3``.
+    # Set special ``xticks`` for ``1/u^k``.
     u_xticks = []
     u_xticklabels = []
     for exponent in (1, 2, 3, 4):
@@ -237,6 +237,170 @@ def main(filename=None):
         plt.close(figure)
 
 
+def main_jlcs10(filename=None):
+    """This recreates the plot from `JLCS10`_.
+
+    .. _JLCS10: https://doi.org/10.1016/j.camwa.2010.05.021
+
+    Note that it is essentially a copy of :func:`main` (and much of the
+    code could be shared, but it isn't worth the effort).
+
+    This seeks to recreate the plot from the original paper, but to show
+    what happens as the exponent on ``1.3`` decreases from ``-44`` down to
+    ``-64``. In particular, it shows that the compensated de Casteljau
+    algorithm produces exactly zero.
+    """
+    # n = 8
+    gamma2n = 16 * U / (1 - 16 * U)
+    bound_coeff1 = float(gamma2n)
+    gamma3n = 24 * U / (1 - 24 * U)
+    bound_coeff2 = float(2 * gamma3n ** 2)
+
+    cond_nums = []
+    forward_errs1 = []
+    forward_errs2 = []
+    for j in range(-5, -64 - 1, -1):
+        s = 0.75 - 1.3 ** j
+        exact_s = F(s)
+
+        # Compute the condition number.
+        exact_p = (exact_s - 1) * (4 * exact_s - 3) ** 7 / 16384
+        # p_tilde(s) = SUM_j |b_j| B_{j, 8}(s) = (s - 1) (s/2 - 3/4)^7
+        exact_p_tilde = (exact_s - 1) * (2 * exact_s - 3) ** 7 / 16384
+        exact_cond = abs(exact_p_tilde / exact_p)
+        cond_nums.append(float(exact_cond))
+
+        # Compute the forward error for uncompensated de Casteljau.
+        b, db = de_casteljau.jlcs10_compensated(s, BEZIER_COEFFS)
+        exact_b1 = F(b)
+        exact_forward_err1 = abs((exact_b1 - exact_p) / exact_p)
[…]

### 2018-04-30 src/de_casteljau.py

@@ -56,6 +56,51 @@ def basic(s, coeffs):
     return pk[0]
 
 
+def jlcs10_compensated(s, coeffs):
+    r"""Performs the compensated de Casteljau algorithm.
+
+    .. _JLCS10: https://doi.org/10.1016/j.camwa.2010.05.021
+
+    This is similar to :func:`compensated` below, but the order of
+    operations exactly matches the `JLCS10`_ paper. In particular,
+    :math:`\widehat{\partial b}_j^{(k)}` is computed as
+
+    .. math::
+
+        \widehat{ell}_{i, j}^{(k)} \oplus \left[s \otimes
+            \widehat{\partial b}_{j + 1}^{(k + 1)}\right] \oplus
+            \left[\widehat{r} \otimes \widehat{\partial b}_j^{(k + 1)}\right]
+
+    and the code in :func:`_compensated_k` uses the "typical" order
+
+    .. math::
+
+        \left[\widehat{r} \otimes \widehat{\partial b}_j^{(k + 1)}\right]
+            \oplus \left[s \otimes \widehat{\partial b}_{j + 1}^{(k + 1)}
+            \right] \oplus \widehat{ell}_{i, j}^{(k)}.
+    """
+    r, rho = eft.add_eft(1.0, -s)
+
+    degree = len(coeffs) - 1
+    bk = {0: list(coeffs), 1: (0.0,) * (degree + 1)}
+    for k in range(degree):
+        new_bk = {0: [], 1: []}
+
+        for j in range(degree - k):
+            P1, pi1 = eft.multiply_eft(r, bk[0][j])
+            P2, pi2 = eft.multiply_eft(s, bk[0][j + 1])
+            S3, sigma3 = eft.add_eft(P1, P2)
+            new_bk[0].append(S3)
+            # Now update the error.
+            wA = pi1 + pi2 + sigma3 + rho * bk[0][j]
+            new_bk[1].append(wA + s * bk[1][j + 1] + r * bk[1][j])
+
+        # Update the "current" values.
+        bk = new_bk
+
+    return bk[0][0], bk[1][0]
+
+
 def local_error(errors, rho, delta_b):
     r"""Compute :math:`\ell` from a list of errors.
 

### 2018-04-30 nox.py
Adding a "smooth drawing" plot for comparison.
I shrunk the width of the interval by a factor of 5, from
[0.75 - 5e-5, 0.75 + 5e-5] to [0.75 - 1e-5, 0.75 + 1e-5].

@@ -129,3 +129,5 @@ def make_images(session):
     }
     script = get_path("scripts", "error_against_cond.py")
     session.run("python", script, env=env)
+    script = get_path("scripts", "smooth_drawing.py")
+    session.run("python", script, env=env)

### 2018-04-30 scripts/error_against_cond.py

@@ -18,13 +18,13 @@ has :math:`\widetilde{p}(s) = (s - 1) \left(\frac{s}{2} -
 """
 
 import fractions
-import os
 
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn
 
 import de_casteljau
+import file_utils
 
 
 F = fractions.Fraction
@@ -41,18 +41,6 @@ BEZIER_COEFFS = (
     -1.0 / 131072.0,
     0.0,
 )
-
-
-def get_path(filename):
-    """Get a file path in the ``images/`` directory.
-
-    This assumes the script is currently in the ``scripts/``
-    directory.
-    """
-    curr_dir = os.path.abspath(os.path.dirname(__file__))
-    root_dir = os.path.dirname(curr_dir)
-    images_dir = os.path.join(root_dir, "images")
-    return os.path.join(images_dir, filename)
 
 
 def main(filename=None):
@@ -231,7 +219,7 @@ def main(filename=None):
     if filename is None:
         plt.show()
     else:
-        path = get_path(filename)
+        path = file_utils.get_path(filename)
         figure.savefig(path, bbox_inches="tight")
         print("Saved {}".format(filename))
         plt.close(figure)
@@ -394,7 +382,7 @@ def main_jlcs10(filename=None):
     if filename is None:
         plt.show()
     else:
-        path = get_path(filename)
+        path = file_utils.get_path(filename)
         figure.savefig(path, bbox_inches="tight")
         print("Saved {}".format(filename))
         plt.close(figure)

### 2018-04-30 scripts/smooth_drawing.py

@@ -1,0 +1,96 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Compare how "smooth" plotted values look.
+
+This will contrast three different implementations (``K = 1, 2, 3``)
+and show that more accuracy produces smoother plots.
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn
+
+import de_casteljau
+import file_utils
+
+
+# p(s) = (s - 1) (s - 3/4)^7
+BEZIER_COEFFS = (
+    2187.0 / 16384.0,
+    -5103.0 / 131072.0,
+    729.0 / 65536.0,
+    -405.0 / 131072.0,
+    27.0 / 32768.0,
+    -27.0 / 131072.0,
+    3.0 / 65536.0,
+    -1.0 / 131072.0,
+    0.0,
+)
+ROOT = 0.75
+DELTA_S = 1e-5
+NUM_POINTS = 400
+
+
+def main(filename=None):
+    s_vals = np.linspace(ROOT - DELTA_S, ROOT + DELTA_S, NUM_POINTS)
+
+    evaluated1 = []
+    evaluated2 = []
+    evaluated3 = []
+
+    for s in s_vals:
+        b, db, d2b = de_casteljau._compensated_k(s, BEZIER_COEFFS, 3)
+        evaluated1.append(b)
+        b2 = b + db
+        evaluated2.append(b2)
+        b3 = b2 + d2b
+        evaluated3.append(b3)
+
+    figure, (ax1, ax2, ax3) = plt.subplots(1, 3, sharex=True)
+    ax1.plot(s_vals, evaluated1)
+    ax2.plot(s_vals, evaluated2)
+    ax3.plot(s_vals, evaluated3)
+
+    # Since ``sharex=True``, ticks only need to be set once.
+    ax1.set_xticks(
+        [
+            ROOT - DELTA_S,
+            ROOT - 0.5 * DELTA_S,
+            ROOT,
+            ROOT + 0.5 * DELTA_S,
[…]

### 2018-04-30 src/file_utils.py

@@ -1,0 +1,27 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Helper(s) for dealing with files."""
+
+import os
+
+
+def get_path(filename):
+    """Get a file path in the ``images/`` directory.
+
+    This assumes the script is currently in the ``src/``
+    directory.
+    """
+    curr_dir = os.path.abspath(os.path.dirname(__file__))
+    root_dir = os.path.dirname(curr_dir)
+    images_dir = os.path.join(root_dir, "images")
+    return os.path.join(images_dir, filename)

### 2018-04-30 doc/paper.tex
Adding a "smooth drawing" plot for comparison.
I shrunk the width of the interval by a factor of 5, from
[0.75 - 5e-5, 0.75 + 5e-5] to [0.75 - 1e-5, 0.75 + 1e-5].
Adding plot showing de Casteljau > Horner.
Also

- Putting new plot in the paper's intro
- Adding `u = 2^{-53}`
- Updating the early introduction of `\ell_j` to just call
  it `\ell_{1, j}`
- Adding a note that `CompDeCasteljau{p, s} = 0` is extra bad
  for root-finding
- Putting the "floating" section with EFT implementations into
  an actual appendix (and referencing the appendix from the theorem
  about sum and product EFTs)

@@
 which grows only linearly with degree, even though the number of
-arithmetic operations grows quadratically. Nevertheless the de Casteljau
+arithmetic operations grows quadratically. The Bernstein basis is
+optimally suited (\cite{Farouki1987}) for polynomial evaluation; it is
+is typically more accurate than the monomial basis, for example in
+Figure~\ref{fig:horner-inferior} evaluation via Horner's method produces
+a jagged curve for points near a tripl root, but the de Casteljau algorithm
+produces a smooth curve. Nevertheless the de Casteljau
 algorithm returns results arbitrarily less accurate than the working
@@
 be given further.
+
+\begin{figure}
+  \includegraphics[width=0.9375\textwidth]{../images/horner_inferior.pdf}
+  \centering
+  \caption{Comparing Horner's method to the de Casteljau method for
+    evaluating \(p(s) = (2s - 1)^3\) in the neighborhood of its
+    multiple root \(1/2\).}
+  \label{fig:horner-inferior}
+\end{figure}
 
@@
 
-In \cite{Graillat2009}, the authors built on the compensated Horner's
+In \cite{Graillat2009}, the authors generalized the compensated Horner's
 algorithm to produce a method for evaluating a polynomial as if
@@
 \(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
-a complete failure to evaluate the order of magnitude of \(p(s)\).
+a complete failure to evaluate the order of magnitude of \(p(s)\). For
+root-finding problems \(\mathtt{CompDeCasteljau}(p, s) = 0\) when
+\(p(s) \neq 0\) can cause premature convergence and incorrect results.
 We describe how to defer rounding into progressively
@@
 as accurate as the result of the de Casteljau algorithm performed in
-\(K\) times the working precision with a final rounding back to this
+\(K\) times the working precision with a final rounding back to the
 working precision.
@@
 and \(\star\) represents the floating point computation, e.g.
-\(a \oplus b = \fl{a + b}\). We also assume that the computed result of
+\(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
+\(\mach = 2^{-53}\).) We also assume that the computed result of
 \(\alpha \in \mathbf{R}\) in floating point arithmetic is denoted by
@@
 
-\begin{theorem}[\cite{Ogita2005}]
+\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
 For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
@@
 The letters \(\sigma\) and \(\pi\) are used to indicate that the
-errors came from sum and product, respectively.
+errors came from sum and product, respectively. See the appendix
+in Section~\ref{sec:appendix-eft} for implementation details.
 \end{theorem}
@@
 \begin{gather}
-\ell_j^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \label{ell-j} \\
+\ell_{1, j}^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \label{ell-j} \\
 (1 - s) \cdot \widehat{b}_j^{(k + 1)} +
   s \cdot \widehat{b}_{j + 1}^{(k + 1)} =
-\widehat{b}_j^{(k)} + \ell_j^{(k)}.
+\widehat{b}_j^{(k)} + \ell_{1, j}^{(k)}.
 \end{gather}
@@
   \partial b_j^{(k)} = (1 - s) \cdot \partial b_j^{(k + 1)} + s \cdot
-  \partial b_{j + 1}^{(k + 1)} + \ell_j^{(k)}.
+  \partial b_{j + 1}^{(k + 1)} + \ell_{1, j}^{(k)}.
 \end{equation}
@@
               \mathtt{TwoSum}(P_1, P_2)\)
-          \State \(\widehat{\ell}_j^{(k)} = \pi_1 \oplus \pi_2 \oplus \sigma_3
-              \oplus \left(\rho \otimes \widehat{b}_j^{(k + 1)}\right)\)
+          \State \(\widehat{\ell}_{1, j}^{(k)} = \pi_1 \oplus \pi_2 \oplus
+              \sigma_3 \oplus \left(\rho \otimes
+              \widehat{b}_j^{(k + 1)}\right)\)
           \State \(\widehat{\partial b}_j^{(k)} =
@@
               \right) \oplus
-              \widehat{\ell}_j^{(k)}\)
+              \widehat{\ell}_{1, j}^{(k)}\)
         \EndFor
@@
 
-\begin{figure}[H]
+Figure~\ref{fig:smooth-drawing} and Figure~\ref{fig:compensated-k}
+show how things improve.
+
+\begin{figure}
+  \includegraphics[width=0.9375\textwidth]{../images/de_casteljau_smooth_drawing.pdf}
+  \centering
+  \caption{Evaluation of \(p(s) = (s - 1)\left(s - 3/4\right)^7\)
+    in the neighborhood of its multiple root \(3/4\).}
+  \label{fig:smooth-drawing}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=0.8125\textwidth]{../images/de_casteljau_rel_error.pdf}
@@
     represented in Bernstein form.}
+  \label{fig:compensated-k}
 \end{figure}
-
-\section{IEEE Stuff}
-
-\begin{algorithm}[H]
-  \caption{\textit{EFT of the sum of two floating point numbers.}}
-
-  \begin{algorithmic}
-    \Function{\(\left[S, \sigma\right] = \mathtt{TwoSum}\)}{$a, b$}
-      \State \(S = a \oplus b\)
-      \State \(z = S \ominus a\)
-      \State \(\sigma = (a \ominus (S \ominus z)) \oplus (b \ominus z)\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-  \caption{\textit{Splitting of a floating point number into two parts.}}
-
-  \begin{algorithmic}
-    \Function{\(\left[h, \ell\right] = \mathtt{Split}\)}{$a$}
-      \State \(z = a \otimes (2^r + 1)\)
-      \State \(h = z \ominus (z \ominus a)\)
-      \State \(\ell = a \ominus h\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-  \caption{\textit{EFT of the product of two floating point numbers.}}
-
-  \begin{algorithmic}
-    \Function{\(\left[P, \pi\right] = \mathtt{TwoProd}\)}{$a, b$}
-      \State \(P = a \otimes b\)
-      \State \(\left[a_h, a_{\ell}\right] = \mathtt{Split}(a)\)
-      \State \(\left[b_h, b_{\ell}\right] = \mathtt{Split}(b)\)
-      \State \(\pi = a_{\ell} \otimes b_{\ell} \ominus (((P \ominus
-          a_h \otimes b_h)
-          \ominus a_{\ell} \otimes b_h) \ominus a_h \otimes b_{\ell})\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-  \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
-
-  \begin{algorithmic}
-    \Function{\(\left[P, \pi\right] = \mathtt{TwoProdFMA}\)}{$a, b$}
-      \State \(P = a \otimes b\)
-      \State \(\pi = \mathtt{FMA}(a, b, -P)\)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
 
@@
 
+\section{Appendix: EFT Algorithms}\label{sec:appendix-eft}
+
+Find here concrete implementation details on the EFTs described
+in Theorem~\ref{thm:eft}. They do not use branches, nor access to the
+mantissa that can be time-consuming
+
+\begin{algorithm}[H]
+  \caption{\textit{EFT of the sum of two floating point numbers.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[S, \sigma\right] = \mathtt{TwoSum}\)}{$a, b$}
+      \State \(S = a \oplus b\)
+      \State \(z = S \ominus a\)
+      \State \(\sigma = (a \ominus (S \ominus z)) \oplus (b \ominus z)\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent In order to avoid branching to check which among
+\(\left|a\right|, \left|b\right|\) is largest, \texttt{TwoSum} uses 6 flops
+rather than 3.
+
+\begin{algorithm}[H]
+  \caption{\textit{Splitting of a floating point number into two parts.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[h, \ell\right] = \mathtt{Split}\)}{$a$}
+      \State \(z = a \otimes (2^r + 1)\)
+      \State \(h = z \ominus (z \ominus a)\)
+      \State \(\ell = a \ominus h\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent For IEEE-754 double precision floating point number, \(r = 26\)
+so \(2^r + 1\) will be known before \texttt{Split} is called. In all,
+\texttt{Split} uses 4 flops.
+
+\begin{algorithm}[H]
+  \caption{\textit{EFT of the product of two floating point numbers.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[P, \pi\right] = \mathtt{TwoProd}\)}{$a, b$}
+      \State \(P = a \otimes b\)
+      \State \(\left[a_h, a_{\ell}\right] = \mathtt{Split}(a)\)
+      \State \(\left[b_h, b_{\ell}\right] = \mathtt{Split}(b)\)
+      \State \(\pi = a_{\ell} \otimes b_{\ell} \ominus (((P \ominus
+          a_h \otimes b_h)
+          \ominus a_{\ell} \otimes b_h) \ominus a_h \otimes b_{\ell})\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent This implementation of \texttt{TwoProd} requires 17 flops.
+For processors that provide a fused-multipy-add operator (\texttt{FMA}),
+\texttt{TwoProd} can be rewritten to use only 2 flops:
+
+\begin{algorithm}[H]
+  \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
+
+  \begin{algorithmic}
+    \Function{\(\left[P, \pi\right] = \mathtt{TwoProdFMA}\)}{$a, b$}
+      \State \(P = a \otimes b\)
+      \State \(\pi = \mathtt{FMA}(a, b, -P)\)
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
 \end{document}

### 2018-04-30 nox.py

@@ -127,7 +127,9 @@ def make_images(session):
         "PYTHONPATH": get_path("src"),
         "SOURCE_DATE_EPOCH": "0",
     }
-    script = get_path("scripts", "error_against_cond.py")
-    session.run("python", script, env=env)
-    script = get_path("scripts", "smooth_drawing.py")
-    session.run("python", script, env=env)
+    names = (
+        "error_against_cond.py", "smooth_drawing.py", "horner_inferior.py"
+    )
+    for name in names:
+        script = get_path("scripts", name)
+        session.run("python", script, env=env)

### 2018-04-30 scripts/horner_inferior.py

@@ -21,45 +21,30 @@ import numpy as np
 import seaborn
 
 import de_casteljau
+import horner
 import file_utils
 
-
-# p(s) = (s - 1) (s - 3/4)^7
-BEZIER_COEFFS = (
-    2187.0 / 16384.0,
-    -5103.0 / 131072.0,
-    729.0 / 65536.0,
-    -405.0 / 131072.0,
-    27.0 / 32768.0,
-    -27.0 / 131072.0,
-    3.0 / 65536.0,
-    -1.0 / 131072.0,
-    0.0,
-)
-ROOT = 0.75
-DELTA_S = 1e-5
+# p(s) = (2s - 1)^3 = (-(1 - s) + s)^3
+POLY_COEFFS = (8.0, -12.0, 6.0, -1.0)
+BEZIER_COEFFS = (-1.0, 1.0, -1.0, 1.0)
+ROOT = 0.5
+DELTA_S = 5e-6
 NUM_POINTS = 400
 
 
 def main(filename=None):
     s_vals = np.linspace(ROOT - DELTA_S, ROOT + DELTA_S, NUM_POINTS)
 
-    evaluated1 = []
-    evaluated2 = []
-    evaluated3 = []
+    horner1 = []
+    de_casteljau1 = []
 
     for s in s_vals:
-        b, db, d2b = de_casteljau._compensated_k(s, BEZIER_COEFFS, 3)
-        evaluated1.append(b)
-        b2 = b + db
-        evaluated2.append(b2)
-        b3 = b2 + d2b
-        evaluated3.append(b3)
+        horner1.append(horner.basic(s, POLY_COEFFS))
+        de_casteljau1.append(de_casteljau.basic(s, BEZIER_COEFFS))
 
-    figure, (ax1, ax2, ax3) = plt.subplots(1, 3, sharex=True)
-    ax1.plot(s_vals, evaluated1)
-    ax2.plot(s_vals, evaluated2)
-    ax3.plot(s_vals, evaluated3)
+    figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
+    ax1.plot(s_vals, horner1)
+    ax2.plot(s_vals, de_casteljau1)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -72,21 +57,20 @@ def main(filename=None):
         ]
     )
 
[…]

### 2018-05-01 doc/paper.tex
Adding an example that shows compensated de Casteljau failing.
Also

- Removing all of the "placeholder" sections
- Adding an error table for an explicit computation (to match the
  plot that fails)
- Removing all `[H]` from floating elements
- Replacing `{algorithm}` with a custom `{breakablealgorithm}`
  that flows across pages like text
- Removing bibliography reference to the Newton application of
  compensated Horner's
- Making the underbraces have the same height via `vphantom`
- Adding actual flop count verification to `nox -s flop_counts`

@@
 \usepackage{mathtools}
+%% H/T: https://tex.stackexchange.com/a/33869/32270
+\makeatletter
+\newenvironment{breakablealgorithm}
+  {% \begin{breakablealgorithm}
+   \begin{center}
+     \refstepcounter{algorithm}% New algorithm
+     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
+     \renewcommand{\caption}[2][\relax]{% Make a new \caption
+       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
+       \ifx\relax##1\relax % #1 is \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
+       \else % #1 is not \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
+       \fi
+       \kern2pt\hrule\kern2pt
+     }
+  }{% \end{breakablealgorithm}
+     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
+   \end{center}
+  }
+\makeatother
 
@@
 optimally suited (\cite{Farouki1987}) for polynomial evaluation; it is
-is typically more accurate than the monomial basis, for example in
+typically more accurate than the monomial basis, for example in
 Figure~\ref{fig:horner-inferior} evaluation via Horner's method produces
-a jagged curve for points near a tripl root, but the de Casteljau algorithm
+a jagged curve for points near a triple root, but the de Casteljau algorithm
 produces a smooth curve. Nevertheless the de Casteljau
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}
   \caption{\textit{de Casteljau algorithm for polynomial evaluation.}}
@@
   \end{algorithmic}
-\end{algorithm}
+\end{breakablealgorithm}
 
@@
 \left[\widehat{r}, \rho\right] &= \mathtt{TwoSum}(1, -s) \\
-\left[P_1, \pi_1\right] &= \mathtt{TwoProd}\left(\widehat{r}, \widehat{b}_j^{(k + 1)}\right) \\
-\left[P_2, \pi_2\right] &= \mathtt{TwoProd}\left(s, \widehat{b}_{j + 1}^{(k + 1)}\right) \\
+\left[P_1, \pi_1\right] &= \mathtt{TwoProd}\left(
+    \widehat{r}, \widehat{b}_j^{(k + 1)}\right) \\
+\left[P_2, \pi_2\right] &= \mathtt{TwoProd}\left(
+    s, \widehat{b}_{j + 1}^{(k + 1)}\right) \\
 \left[\widehat{b}_j^{(k)}, \sigma_3\right] &= \mathtt{TwoSum}(P_1, P_2)
@@
 \begin{gather}
-\ell_{1, j}^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho\widehat{b}_j^{(k + 1)} \label{ell-j} \\
+\ell_{1, j}^{(k)} = \pi_1 + \pi_2 + \sigma_3 + \rho \cdot
+  \widehat{b}_j^{(k + 1)} \label{ell-j} \\
 (1 - s) \cdot \widehat{b}_j^{(k + 1)} +
@@
 
-\begin{figure}[H]
+\begin{figure}\label{fig:loc-err-accumulate}
 \centering
@@
 \end{equation}
-we can see that the local errors accumulate in
+we can see (Figure~\ref{fig:loc-err-accumulate}) that the local errors
+accumulate in
 \(\partial b^{(k)}\):
-\begin{equation}
+\begin{equation}\label{err-update}
   \partial b_j^{(k)} = (1 - s) \cdot \partial b_j^{(k + 1)} + s \cdot
@@
 \end{equation}
-Hence, when computed in exact arithmetic, we have
+When computed in exact arithmetic, we would have
 \begin{equation}
-  p(s) = \widehat{b}_0^{(0)} + \partial b_0^{(0)}.
+  p(s) = \widehat{b}_0^{(0)} + \partial b_0^{(0)}
 \end{equation}
+and by using \eqref{err-update}, we can continue to compute
+approximations of \(\partial b_j^{(k)}\). The idea behind the compensated
+de Casteljau algorithm is to compute both the local error and the updates
+of the global error with floating point operations:
 
-The idea behind the compensated de Casteljau algorithm
-is to compute both the local error and the updates of the global
-error with floating point operations:
-
-\begin{algorithm}[H]
+\begin{breakablealgorithm}\label{alg:comp-de-casteljau}
   \caption{\textit{Compensated de Casteljau algorithm for polynomial evaluation.}}
@@
 \end{theorem}
+
+\begin{figure}
+  \includegraphics[width=0.9375\textwidth]{../images/compensated_insufficient.pdf}
+  \centering
+  \caption{Compensated Casteljau method starts to lose accuracy
+    for \(p(s) = (2s - 1)^3 (s - 1)\) in the neighborhood of its
+    multiple root \(1/2\).}
+  \label{fig:compensated-insufficient}
+\end{figure}
+
+Unfortunately, Figure~\ref{fig:compensated-insufficient} shows how
+the \texttt{CompDeCasteljau} starts to break down in a region of
+high condition number (caused by a multiple root with multiplicity
+higher than two). For example, the point
+\(s = \frac{1}{2} + 1001\mach\)
+--- which is in the plotted region \(\left|s - \frac{1}{2}\right|
+\leq 1.5 \cdot 10^{-11}\) --- evaluates to exactly \(0\) when it should be
+\(\mathcal{O}\left(\mach^3\right)\). As shown in
+Table~\ref{tab:exact-computation}, the breakdown occurs because
+\(\widehat{b}_0^{(0)} = -\widehat{\partial b}_0^{(0)} = \mach / 16\).
+
+\begin{table}
+  \centering
+  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
+    \toprule
+    k & j & \widehat{b}_j^{(k)} & \widehat{\partial b}_j^{(k)} & \partial b_j^{(k)} - \widehat{\partial b}_j^{(k)} \\
+    \midrule
+    3 & 0 & 0.125 - 1.75 (1001 \mach) - 0.25 \mach & 0.25\mach & 0 \\
+    3 & 1 & -0.125 + 1.25(1001 \mach) + 0.25 \mach & -0.25\mach & 0 \\
+    3 & 2 & 0.125 - 0.75 (1001 \mach) & 0 & 0 \\
+    3 & 3 & -0.125 + 0.25 (1001 \mach) & 0 & 0 \\
+    \midrule
+    2 & 0 & -0.5 (1001 \mach) & 3 (1001 \mach)^2 & 0 \\
+    2 & 1 & 0.5(1001 \mach) + 0.125 \mach & -0.125\mach - 2 (1001 \mach)^2 & 0 \\
+    2 & 2 & -0.5 (1001 \mach) & (1001 \mach)^2 & 0 \\
+    \midrule
+    1 & 0 & 0.0625\mach + (1001 \mach)^2 + 239\mach^2 & -0.0625\mach + 0.5  (1001 \mach)^2 - 239 \mach^2 & -5 (1001\mach)^3 \\
+    1 & 1 & 0.0625\mach - (1001 \mach)^2 - 239\mach^2 & -0.0625\mach - 0.5  (1001 \mach)^2 + 239 \mach^2 & 3 (1001\mach)^3 \\
+    \midrule
+    0 & 0 & 0.0625 \mach & -0.0625 \mach & -4 (1001 \mach)^3 + 8 (1001 \mach)^4 \\
+    \bottomrule
+  \end{tabular}
+  \caption{Terms computed by \texttt{CompDeCasteljau} when evaluating
+    \(p(s) = (2s - 1)^3 (s - 1)\) at the point
+    \(s = \frac{1}{2} + 1001 \mach\)}
+  \label{tab:exact-computation}
+\end{table}
 
@@
 
-In \texttt{CompDeCasteljau}, after a single stage of error filtering we
-``give up'' and use \(\widehat{\partial b}^{(k)}\) instead of
-\(\partial b^{(k)}\) (without keeping around any information about the
+\begin{figure}\label{fig:error-filtration}
+\centering
+\begin{tikzpicture}[node distance = 0.2\textwidth]
+    \node [box] (bj) {\(b\)};
+    \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
+    \node [box, right of=bn0] (dbn0) {\(\widehat{\partial b}^{(n)}\)};
+    \node [box, right of=dbn0] (ddbn0) {\(\widehat{\partial^2 b}^{(n)}\)};
+    \node [box, right of=ddbn0] (dddbn0) {\(\widehat{\partial^3 b}^{(n)}\)};
+    %%
+    \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
+    \node [box, below of=dbn0] (dbn1) {\(\widehat{\partial b}^{(n - 1)}\)};
+    \node [box, below of=ddbn0] (ddbn1) {\(\widehat{\partial^2 b}^{(n - 1)}\)};
+    \node [box, below of=dddbn0] (dddbn1) {\(\widehat{\partial^3 b}^{(n - 1)}\)};
+    %%
+    \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
+    \node [box, below of=dbn1] (dbn2) {\(\widehat{\partial b}^{(n - 2)}\)};
+    \node [box, below of=ddbn1] (ddbn2) {\(\widehat{\partial^2 b}^{(n - 2)}\)};
+    \node [box, below of=dddbn1] (dddbn2) {\(\widehat{\partial^3 b}^{(n - 2)}\)};
+    %%
+    \path [line] (bn0) -- node[above] {\(\widehat{\ell}_1^{(n)}\)} (dbn0);
+    \path [line] (dbn0) -- node[above] {\(\widehat{\ell}_2^{(n)}\)} (ddbn0);
+    \path [line] (ddbn0) -- node[above] {\(\widehat{\ell}_3^{(n)}\)} (dddbn0);
+    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
+    \path [line] (dddbn0) -- node[above] {\(\widehat{\ell}_4^{(n)}\)} (right0);
+    %%
+    \path [line] (bn1) -- node[above] {\(\widehat{\ell}_1^{(n - 1)}\)} (dbn1);
+    \path [line] (dbn1) -- node[above] {\(\widehat{\ell}_2^{(n - 1)}\)} (ddbn1);
+    \path [line] (ddbn1) -- node[above] {\(\widehat{\ell}_3^{(n - 1)}\)} (dddbn1);
+    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
+    \path [line] (dddbn1) -- node[above] {\(\widehat{\ell}_4^{(n - 1)}\)} (right1);
+    %%
+    \path [line] (bn2) -- node[above] {\(\widehat{\ell}_1^{(n - 2)}\)} (dbn2);
+    \path [line] (dbn2) -- node[above] {\(\widehat{\ell}_2^{(n - 2)}\)} (ddbn2);
+    \path [line] (ddbn2) -- node[above] {\(\widehat{\ell}_3^{(n - 2)}\)} (dddbn2);
+    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
+    \path [line] (dddbn2) -- node[above] {\(\widehat{\ell}_4^{(n - 2)}\)} (right2);
+    %%
+    \path [line] (bj) -- (bn0);
+    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
+    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
+    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
+    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
+    %%
+    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
+    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
+    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
+    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
+    %%
+    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
+    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
+    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
+    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
+    %%
+    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
+    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
+    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
+    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
+\end{tikzpicture}
+\caption{Filtering errors}
+\end{figure}
+
+In \texttt{CompDeCasteljau} (Algorithm~\ref{alg:comp-de-casteljau}), after
+a single stage of error filtering we
+``give up'' and use \(\widehat{\partial b}\) instead of
+\(\partial b\) (without keeping around any information about the
 round-off error). In order to obtain results that are as accurate as if
 computed in \(K\) times the working precision, we must continue filtering
-errors down \((K - 1)\) times, so that at the final level we compute
-\(\widehat{\partial^{K - 1} b}^{(k)}\) and don't worry about round-off
-error.
+(see Figure~\ref{fig:error-filtration})
+errors down \((K - 1)\) times, and only at the final level do we accept
+the rounded
+\(\widehat{\partial^{K - 1} b}\) in place of the exact
+\(\partial^{K - 1} b\).
 
@@
 For example, in~\eqref{ell-j}:
+%% H/T: https://tex.stackexchange.com/a/154333/32270
 \begin{equation}
-\ell_{1, j}^{(k)} = \underbrace{\pi_1}_{P_1 = \widehat{r} \otimes
-    \widehat{b}_j^{(k + 1)}} +
-\underbrace{\pi_2}_{P_2 = s \otimes \widehat{b}_{j + 1}^{(k + 1)}} +
-\underbrace{\sigma_3}_{P_1 \oplus P_2} +
-\underbrace{\rho \cdot \widehat{b}_j^{(k + 1)}}_{(1 - s)
-  \widehat{b}_j^{(k + 1)}}
+\ell_{1, j}^{(k)} =
+    \underbrace{\vphantom{\rho \cdot \widehat{b}_j^{(k + 1)}} \pi_1}_{
+        \vphantom{(1 - s) \widehat{b}_j^{(k + 1)}}
+        P_1 = \widehat{r} \otimes \widehat{b}_j^{(k + 1)}} +
+    \underbrace{\vphantom{\rho \cdot \widehat{b}_j^{(k + 1)}} \pi_2}_{
+        \vphantom{(1 - s) \widehat{b}_j^{(k + 1)}}
+        P_2 = s \otimes \widehat{b}_{j + 1}^{(k + 1)}} +
+    \underbrace{\vphantom{\rho \cdot \widehat{b}_j^{(k + 1)}} \sigma_3}_{
+        \vphantom{(1 - s) \widehat{b}_j^{(k + 1)}}
+        P_1 \oplus P_2} +
+    \underbrace{\rho \cdot \widehat{b}_j^{(k + 1)}}_{
+        (1 - s) \widehat{b}_j^{(k + 1)}}
 \end{equation}
 After each stage, we'll always have
-\[\ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
-\widehat{\partial^{F - 1} b}_j^{(k + 1)}\]
+\begin{equation}
+\ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
+\widehat{\partial^{F - 1} b}_j^{(k + 1)}
+\end{equation}
 where the terms \(e_1, \ldots, e_M\) come from using \texttt{TwoSum} and
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}
   \caption{\textit{EFT for computing the local error.}}
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}\label{alg:k-comp-de-casteljau}
   \caption{\(K\)-\textit{compensated de Casteljau algorithm.}}
@@
   \end{algorithmic}
-\end{algorithm}
+\end{breakablealgorithm}
 
-\begin{figure}[H]
-\centering
-\begin{tikzpicture}[node distance = 0.2\textwidth]
-    \node [box] (bj) {\(b\)};
-    \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
-    \node [box, right of=bn0] (dbn0) {\(\widehat{\partial b}^{(n)}\)};
-    \node [box, right of=dbn0] (ddbn0) {\(\widehat{\partial^2 b}^{(n)}\)};
-    \node [box, right of=ddbn0] (dddbn0) {\(\widehat{\partial^3 b}^{(n)}\)};
-    %%
-    \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
-    \node [box, below of=dbn0] (dbn1) {\(\widehat{\partial b}^{(n - 1)}\)};
-    \node [box, below of=ddbn0] (ddbn1) {\(\widehat{\partial^2 b}^{(n - 1)}\)};
-    \node [box, below of=dddbn0] (dddbn1) {\(\widehat{\partial^3 b}^{(n - 1)}\)};
-    %%
-    \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
-    \node [box, below of=dbn1] (dbn2) {\(\widehat{\partial b}^{(n - 2)}\)};
-    \node [box, below of=ddbn1] (ddbn2) {\(\widehat{\partial^2 b}^{(n - 2)}\)};
-    \node [box, below of=dddbn1] (dddbn2) {\(\widehat{\partial^3 b}^{(n - 2)}\)};
-    %%
-    \path [line] (bn0) -- node[above] {\(\widehat{\ell}_1^{(n)}\)} (dbn0);
-    \path [line] (dbn0) -- node[above] {\(\widehat{\ell}_2^{(n)}\)} (ddbn0);
-    \path [line] (ddbn0) -- node[above] {\(\widehat{\ell}_3^{(n)}\)} (dddbn0);
-    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
-    \path [line] (dddbn0) -- node[above] {\(\widehat{\ell}_4^{(n)}\)} (right0);
-    %%
-    \path [line] (bn1) -- node[above] {\(\widehat{\ell}_1^{(n - 1)}\)} (dbn1);
-    \path [line] (dbn1) -- node[above] {\(\widehat{\ell}_2^{(n - 1)}\)} (ddbn1);
-    \path [line] (ddbn1) -- node[above] {\(\widehat{\ell}_3^{(n - 1)}\)} (dddbn1);
-    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
-    \path [line] (dddbn1) -- node[above] {\(\widehat{\ell}_4^{(n - 1)}\)} (right1);
-    %%
-    \path [line] (bn2) -- node[above] {\(\widehat{\ell}_1^{(n - 2)}\)} (dbn2);
-    \path [line] (dbn2) -- node[above] {\(\widehat{\ell}_2^{(n - 2)}\)} (ddbn2);
-    \path [line] (ddbn2) -- node[above] {\(\widehat{\ell}_3^{(n - 2)}\)} (dddbn2);
-    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
-    \path [line] (dddbn2) -- node[above] {\(\widehat{\ell}_4^{(n - 2)}\)} (right2);
-    %%
-    \path [line] (bj) -- (bn0);
-    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
-    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
-    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
-    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
-    %%
-    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
-    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
-    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
-    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
-    %%
-    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
-    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
-    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
-    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
-    %%
-    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
-    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
-    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
-    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
-\end{tikzpicture}
-\caption{Filtering errors}
-\end{figure}
+Noting that \(\ell_{F, j}\) contains \(5F - 2\) terms, one can
+show that Algorithm~\ref{alg:k-comp-de-casteljau} requires
+\begin{equation}
+(15K^2 + 11K - 34)T_n + K + 5 =
+\mathcal{O}\left(n^2 K^2\right)
+\end{equation}
+flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
+\(n\)th triangular number. As a comparison, the non-compensated form of
+de Casteljau requires \(3 T_n + 1\) flops. In total this will require
+\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
+FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
+instead, lowering the flop count by \(15(3K - 4)T_n\).
 
@@
 
-\section{de Casteljau's Method}
-
-Consider de Casteljau's method to evaluate a degree \(n\)
-polynomial in Bernstein-B\'{e}zier form with control points \(b_j\):
-\begin{align}
-    b_j^{(n)} &= b_j \\
-    b_j^{(k)} &= (1 - s) b_j^{(k + 1)} + s b_{j + 1}^{(k + 1)} \\
-    b(s) &= b_0^{(0)}.
-\end{align}
-
-\subsection{Condition Number}
-
-For a polynomial \(p(x)\) in the power basis, we have
-(\cite{langlois_et_al:DSP:2006:442}):
-\begin{equation}
-\cond{p(x)} = \frac{\widetilde{p}\left(\left|x\right|\right)}{
-  \left|p(x)\right|} = \frac{\sum_j \left|a_j\right| \left|x\right|^j}{
-  \left|p(x)\right|}.
-\end{equation}
-In particular, this means that if \(x \geq 0\) and each \(a_j \geq 0\)
-we must necessarily have \(\cond{p(x)} = 1\). To see an example in use,
-consider \(p(x) = (x - 1)^n\) and input values of the form \(x = 1 + \delta\)
-(with \(\left|\delta\right| \ll 1\)). Since \(a_j = \binom{n}{j} (-1)^{n - j}\)
-we have \(\widetilde{p}(x) = (x + 1)^n\) hence
-\begin{equation}
-\cond{p\left(1 + \delta\right)} = \frac{(2 + \delta)^n}{
-  \left|\delta\right|^n} = \left|1 + \frac{2}{\delta}\right|^n.
-\end{equation}
-As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
-
-For a polynomial \(p(s)\) in Bernstein form, we have (\cite{Jiang2010}):
-\begin{equation}
-\cond{p(s)} = \frac{\widetilde{p}\left(s\right)}{
-  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| \left|B_{j, n}(s)\right|}{
-  \left|p(s)\right|}.
-\end{equation}
-The Bernstein form is suited for \(s \in \left[0, 1\right]\), which means
-\(B_{j, n}(s) \geq 0\) typically. If \(s \in \left[0, 1\right]\) and each
-\(b_j \geq 0\) we must necessarily have \(\cond{p(s)} = 1\). To see an
-example in use, consider
-\begin{equation}
-p(s) = (1 - 2s)^n = \left[(1 - s) - s\right]^n = \sum_j \binom{n}{j}
-(1 - s)^{n - j} (-s)^j = \sum_j (-1)^j B_{j, n}(s)
-\end{equation}
-and input values of the form \(x = \frac{1}{2} + \delta\)
-(with \(\left|\delta\right| \ll \frac{1}{2}\)). Since \(b_j = (-1)^j\)
-we have \(\widetilde{p}(s) = \left[(1 - s) + s\right]^n = 1\)
-\begin{equation}
-\cond{p\left(\frac{1}{2} + \delta\right)} = \frac{1}{
-  \left|2\delta\right|^n}.
-\end{equation}
-As \(\delta \to 0\), this value approaches \(\infty\) (as expected).
-
-\subsection{Example of Compensated de Casteljau Failing}
-
-Consider
-\begin{equation}
-p(s) = (4s - 3)^3 (8s + 7) = (-189) B_{0, 4}(s) + (-54) B_{1, 4}(s) +
-57 B_{2, 4}(s) + (-32) B_{3, 4}(s) + 15 B_{4, 4}(s)
-\end{equation}
-at the point \(s = \frac{3}{4} + 800 \mach\). In exact arithmetic, we
-have \(p(s) = 13(3200\mach)^3 + \mathcal{O}\left(\mach^4\right)\).
-However, using the compensated de Casteljau algorithm results in
-\(\widehat{b}_0^{(0)} \oplus \widehat{\partial b}_0^{(0)} = 0\):
-
-\begin{center}
-  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
-    \toprule
-    k & j & \widehat{b}_j^{(k)} & \widehat{\partial b}_j^{(k)} & \widehat{\partial^2 b}_j^{(k)} & \widehat{\partial^3 b}_j^{(k)} \\
-    \midrule
-    3 & 0 & -87\frac{3}{4} + 108032\mach & -32\mach & 0 & 0 \\
-    3 & 1 & 29\frac{1}{4} + 88768\mach & 32\mach & 0 & 0 \\
-    3 & 2 & -9\frac{3}{4} - 71200\mach & 0 & 0 & 0 \\
-    3 & 3 & 3\frac{1}{4} + 37600\mach & 0 & 0 & 0 \\
-    \midrule
-    2 & 0 & 187200\mach & -15360000\mach^2 & 0 & 0 \\
-    2 & 1 & -62408\mach & 8\mach - 128000000\mach^2 & 0 & 0 \\
-    2 & 2 & 20800\mach & 87040000\mach^2 & 0 & 0 \\
-    \midrule
-    1 & 0 & -6\mach - 199688192\mach^2 & 6\mach - 99831808\mach^2 & -90112000000\mach^3 & 0 \\
-    1 & 1 & -2\mach + 66568192\mach^2  & 2\mach + 33271808\mach^2 & 172032000000\mach^3 & 0 \\
-    \midrule
-    0 & 0 & -3\mach + 7296\mach^2 & 3\mach - 7296\mach^2 & 13 (3200\mach)^3 + 3048(512\mach)^4 & 962(128\mach)^4 \\
-    \bottomrule
-  \end{tabular}
-\end{center}
-
-\subsection{Selection of Test Cases}
+\section{Quotes: Selection of Test Cases}
 
@@
 
-\subsection{Operation Counts}
-
-After implementing for \(K = 2, 3, \ldots, 12\) and instrumenting all
-relevant floating point operations, the \(K\)-fold Horner requires
-\begin{equation}
-(5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right) =
-\mathcal{O}\left((n + K)2^K\right)
-\end{equation}
-flops to evaluate a degree \(n\) polynomial (this only applies when
-\(n \geq K - 1\)). As a comparison, the
-non-compensated form of Horner requires \(2n\) flops. Of these,
-\(\left(2^{K - 1} - 1\right)n - 2^{K - 1}(K - 3) - 2\) are
-FMA (fused-multiply-add) instructions.
-
-After implementing for \(K = 2, 3, 4, 5\) and instrumenting all relevant
-floating point operations, the \(K\)-fold de Casteljau requires
-\begin{equation}
-(15K^2 - 34K + 26)T_n + K + 5 =
-\mathcal{O}\left(n^2 K^2\right)
-\end{equation}
-flops to evaluate a degree \(n\) polynomial. (Here \(T_n\) is the
-\(n\)th triangular number.) As a comparison, the non-compensated form of
-de Casteljau requires \(3 T_n + 1\) flops. Of these, \((3K - 4)T_n\) are
-FMA instructions. On hardware that doesn't support FMA,
-every FMA will be exchanged for 10 \(\ominus\)'s and 6 \(\otimes\)'s so the
-count will increase by \((10 + 6 - 1)(3K - 4)T_n\).
-
-\section{Bogus Section for Refs}
-
-Here they are, for now
-\begin{itemize}
-  \item Newton with compensated Horner (\cite{Graillat2008})
-\end{itemize}
-
 \bibliography{paper}
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}
   \caption{\textit{EFT of the sum of two floating point numbers.}}
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}
   \caption{\textit{Splitting of a floating point number into two parts.}}
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}
   \caption{\textit{EFT of the product of two floating point numbers.}}
@@
 
-\begin{algorithm}[H]
+\begin{breakablealgorithm}\label{alg:two-prod-fma}
   \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}

### 2018-05-01 nox.py

@@ -128,7 +128,10 @@ def make_images(session):
         "SOURCE_DATE_EPOCH": "0",
     }
     names = (
-        "error_against_cond.py", "smooth_drawing.py", "horner_inferior.py"
+        "error_against_cond.py",
+        "smooth_drawing.py",
+        "horner_inferior.py",
+        "compensated_insufficient.py",
     )
     for name in names:
         script = get_path("scripts", name)

### 2018-05-01 scripts/compensated_insufficient.py

@@ -1,0 +1,88 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Show an example where compensated de Casteljau is still not enough."""
+
+import fractions
+
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn
+
+import de_casteljau
+import file_utils
+
+
+F = fractions.Fraction
+# p(s) = (2s - 1)^3 (s - 1)
+BEZIER_COEFFS = (1.0, -0.75, 0.5, -0.25, 0.0)
+ROOT = 0.5
+DELTA_S = 1.5e-11
+NUM_POINTS = 400
+
+
+def main(filename=None):
+    s_vals = np.linspace(ROOT - DELTA_S, ROOT + DELTA_S, NUM_POINTS)
+
+    de_casteljau2 = []
+    exact = []
+
+    for s in s_vals:
+        de_casteljau2.append(de_casteljau.compensated(s, BEZIER_COEFFS))
+
+        exact_s = F(s)
+        exact_p = (2 * exact_s - 1) ** 3 * (exact_s - 1)
+        exact.append(float(exact_p))
+
+    figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
+
+    ax1.plot(s_vals, de_casteljau2)
+    ax2.plot(s_vals, exact)
+
+    # Since ``sharex=True``, ticks only need to be set once.
+    ax1.set_xticks(
+        [
+            ROOT - DELTA_S,
+            ROOT - 0.5 * DELTA_S,
+            ROOT,
+            ROOT + 0.5 * DELTA_S,
+            ROOT + DELTA_S,
+        ]
+    )
+
+    ax1.set_title(r"$\mathtt{CompDeCasteljau}$")
+    ax2.set_title("$p(s)$")
+
+    if filename is None:
+        plt.show()
+    else:
+        # NOTE: These are (intended to be) the same settings used in
+        #       ``horner_inferior.py``, so they should probably be
+        #       kept in sync.
+        figure.set_size_inches(9.87, 4.8)
[…]

### 2018-05-01 scripts/compute_counts.py

@@ -30,6 +30,7 @@ def count_add_eft():
     sum_, error = eft.add_eft(val1, val2)
     assert sum_.value == 2.0
     assert error.value == 0.5 ** 52
+    assert parent.count == 6
     print("     add_eft(): {}".format(parent.display))
 
 
@@ -39,6 +40,7 @@ def count__split():
     high, low = eft._split(val)
     assert high.value == 1.0
     assert low.value == 0.5 ** 27
+    assert parent.count == 4
     print("      _split(): {}".format(parent.display))
 
 
@@ -53,8 +55,10 @@ def count_multiply_eft():
         assert error.value == -0.5 ** 80
         if use_fma:
             description = "with FMA:    "
+            assert parent.count == 2
         else:
             description = "w / out FMA: "
+            assert parent.count == 17
         print("  {} {}".format(description, parent.display))
 
 
@@ -66,6 +70,7 @@ def count__vec_sum():
         eft._vec_sum(p)
 
         assert p[size_p - 1].value == float(size_p)
+        assert parent.count == 6 * (size_p - 1)
         print("  |p| = {}:      {}".format(size_p, parent.display))
 
 
@@ -79,6 +84,7 @@ def count_sum_k():
             total = eft.sum_k(p, k)
 
             assert total.value == float(size_p)
+            assert parent.count == (6 * k - 5) * (size_p - 1)
             print("    |p| = {}:    {}".format(size_p, parent.display))
 
 
@@ -90,7 +96,34 @@ def count_horner_basic():
         coeffs = (operation_count.Float(1.0, parent),) * (degree + 1)
         p = horner.basic(x, coeffs)
         assert p.value == 2.0 ** (degree + 1) - 1
+        assert parent.count == 2 * degree
         print("  degree {}:     {}".format(degree, parent.display))
+
+
+def horner_expected_total(K, n):
+    """Get the expected flop count for compensated Horner's method.
+
+    When using FMA, the count is
+
+    .. math::
+
+       (5 \cdot 2^K - 8)n + \left((K + 8) 2^K - 12K - 6\right).
+    """
+    return (5 * 2 ** K - 8) * n + ((K + 8) * 2 ** K - 12 * K - 6)
+
+
+def horner_expected_fma(K, n):
+    """Get the FMA count for compensated Horner's method.
+
+    When using FMA, the count is
+
+    .. math::
+
+       \left(2^{K - 1} - 1\right)n - 2^{K - 1}(K - 3) - 2
+
[…]

### 2018-05-01 scripts/error_against_cond.py

@@ -29,7 +29,7 @@ import file_utils
 
 F = fractions.Fraction
 U = F(1, 2 ** 53)
-# p(s) = (s - 1) (s - 3/4)^7 = SUM_j b_j B_{j, 8}(s)
+# p(s) = (s - 1) (s - 3/4)^7
 BEZIER_COEFFS = (
     2187.0 / 16384.0,
     -5103.0 / 131072.0,
@@ -41,6 +41,7 @@ BEZIER_COEFFS = (
     -1.0 / 131072.0,
     0.0,
 )
+ROOT = 0.75
 
 
 def main(filename=None):
@@ -56,7 +57,7 @@ def main(filename=None):
     forward_errs3 = []
     forward_errs4 = []
     for j in range(-5, -90 - 1, -1):
-        s = 0.75 - 1.3 ** j
+        s = ROOT - 1.3 ** j
         exact_s = F(s)
 
         # Compute the condition number.
@@ -248,7 +249,7 @@ def main_jlcs10(filename=None):
     forward_errs1 = []
     forward_errs2 = []
     for j in range(-5, -64 - 1, -1):
-        s = 0.75 - 1.3 ** j
+        s = ROOT - 1.3 ** j
         exact_s = F(s)
 
         # Compute the condition number.

### 2018-05-01 src/de_casteljau.py

@@ -92,8 +92,8 @@ def jlcs10_compensated(s, coeffs):
             S3, sigma3 = eft.add_eft(P1, P2)
             new_bk[0].append(S3)
             # Now update the error.
-            wA = pi1 + pi2 + sigma3 + rho * bk[0][j]
-            new_bk[1].append(wA + s * bk[1][j + 1] + r * bk[1][j])
+            l_hat = pi1 + pi2 + sigma3 + rho * bk[0][j]
+            new_bk[1].append(l_hat + s * bk[1][j + 1] + r * bk[1][j])
 
         # Update the "current" values.
         bk = new_bk

### 2018-05-01 doc/paper.tex
Making captions prettier.
Did so by making the label bold and limiting the
captions to 75% the width of a line.

Also fixing a slight type "Casteljau" --> "de Casteljau".

@@
 \makeatother
+%% H/T: https://tex.stackexchange.com/a/202047/32270
+%%      https://tex.stackexchange.com/a/32463/32270
+\usepackage[labelfont=bf]{caption}
 
@@
   \centering
+  \captionsetup{width=.75\linewidth}
   \caption{Comparing Horner's method to the de Casteljau method for
@@
   \centering
+  \captionsetup{width=.75\linewidth}
   \caption{Evaluation of \(p(s) = (s - 1)\left(s - 3/4\right)^7\)
@@
 \end{tikzpicture}
+\captionsetup{width=.75\linewidth}
 \caption{Local round-off errors}
@@
   \centering
-  \caption{Compensated Casteljau method starts to lose accuracy
+  \captionsetup{width=.75\linewidth}
+  \caption{The compensated de Casteljau method starts to lose accuracy
     for \(p(s) = (2s - 1)^3 (s - 1)\) in the neighborhood of its
@@
   \end{tabular}
-  \caption{Terms computed by \texttt{CompDeCasteljau} when evaluating
+  \captionsetup{width=.75\linewidth}
+  \caption{Terms computed by \texttt{CompDeCasteljau} when evaluating \\
     \(p(s) = (2s - 1)^3 (s - 1)\) at the point
@@
 \end{tikzpicture}
+\captionsetup{width=.75\linewidth}
 \caption{Filtering errors}
@@
   \centering
+  \captionsetup{width=.75\linewidth}
   \caption{Accuracy of evaluation of \(p(s) = (s - 1)\left(s - 3/4\right)^7\)

### 2018-05-01 doc/paper.tex
Removing dummy "Quotes" section.
Also

- Removing the Bezerra13 paper from bibliography (it certainly
  wasn't needed)
- Adding the other two papers about Bernstein evaluation to the
  Farouki reference (i.e. they are just supporting the fact that
  the de Casteljau algorithm is not so bad)
- Moving all `label`-s after `caption` when relevant (this was
  confusing `pdflatex`)
- Making sure the bibliography shows up in the TOC

@@
 \usepackage[labelfont=bf]{caption}
+%% H/T: https://tex.stackexchange.com/a/71153/32270
+\usepackage[nottoc,notlot,notlof]{tocbibind}
 
@@
 arithmetic operations grows quadratically. The Bernstein basis is
-optimally suited (\cite{Farouki1987}) for polynomial evaluation; it is
+optimally suited (\cite{Farouki1987}, \cite{Delgado2015}, \cite{Mainar2005})
+for polynomial evaluation; it is
 typically more accurate than the monomial basis, for example in
@@
 
-\begin{figure}\label{fig:loc-err-accumulate}
+\begin{figure}
 \centering
@@
 \caption{Local round-off errors}
+\label{fig:loc-err-accumulate}
 \end{figure}
@@
 
-\begin{breakablealgorithm}\label{alg:comp-de-casteljau}
+\begin{breakablealgorithm}
   \caption{\textit{Compensated de Casteljau algorithm for polynomial evaluation.}}
+  \label{alg:comp-de-casteljau}
 
@@
 
-\begin{figure}\label{fig:error-filtration}
+\begin{figure}
 \centering
@@
 \caption{Filtering errors}
+\label{fig:error-filtration}
 \end{figure}
@@
 
-\begin{breakablealgorithm}\label{alg:k-comp-de-casteljau}
+\begin{breakablealgorithm}
   \caption{\(K\)-\textit{compensated de Casteljau algorithm.}}
+  \label{alg:k-comp-de-casteljau}
 
@@
 
-\section{Quotes: Selection of Test Cases}
-
-From \cite{Delgado2015} (end of Section 3):
-
-\begin{quote}
-  We can observe that, in this case, the algorithm with a good
-  behavior everywhere is the de Casteljau algorithm
-\end{quote}
-
-\noindent In the same paper (when referring to \cite{Bezerra2013} at the
-beginning of Section 2):
-
-\begin{quote}
-  assuming that all control points are positive. This assumption avoided
-  ill-conditioned polynomials. In this section, we shall show that this is
-  a natural assumption in Computer Aided Geometric Design (from now on,
-  C.A.G.D.) and that it permits to assure high relative precision for the
-  evaluation through a large family of representations in C.A.G.D.
-\end{quote}
-
-\noindent From the same author, in \cite{Mainar2005} (towards the
-end of Section 5, at the bottom of page 109):
-
-\begin{quote}
-  Let us observe that in this case, the de Casteljau algorithm presents
-  better stability properties for the evaluation near the roots. In fact,
-  the de Casteljau algorithm has good behaviour even when using simple
-  precision, although the running error bound is not so accurate in points
-  close to the roots.
-\end{quote}
-
 \bibliography{paper}
@@
 
-\begin{breakablealgorithm}\label{alg:two-prod-fma}
+\begin{breakablealgorithm}
   \caption{\textit{EFT of the sum of two floating point numbers with a FMA.}}
+  \label{alg:two-prod-fma}
 

### 2018-05-02 doc/paper.tex
Adding the content of the "Numerical experiements" section.
Also adding the start of a theorem for the a priori error bound.

@@
 \renewcommand{\headrulewidth}{0pt}
+\renewcommand{\qed}{\(\blacksquare\)}
 \newcommand{\cond}[1]{\operatorname{cond}\left(#1\right)}
@@
   \caption{\textit{de Casteljau algorithm for polynomial evaluation.}}
+  \label{alg:de-casteljau}
 
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$b, s$}
+    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljauK}\)}{$b, s, K$}
       \State \(n = \texttt{length}(b) - 1\)
@@
 
+\begin{theorem}
+  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
+  \begin{equation}
+    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
+      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
+      \right] +
+      \left[\left(\alpha(n) \mach\right)^K + \mathcal{O}\left(
+      \mach^{K + 1}\right)\right] \cond{p, s}.
+  \end{equation}
+\end{theorem}
+
+\begin{proof}
+Note that
+\begin{equation}
+p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \partial b_0^{(0)}
+= \cdots
+= \widehat{b}_0^{(0)} + \widehat{\partial b}_0^{(0)} + \cdots
++ \widehat{\partial^{K - 2} b}_0^{(0)} + \partial^{K - 1} b_0^{(0)}
+\end{equation}
+and
+\begin{equation}
+\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
+\widehat{\partial b}_0^{(0)} \oplus \cdots
+\oplus \widehat{\partial^{K - 2} b}_0^{(0)} \oplus
+\widehat{\partial^{K - 1} b}_0^{(0)}.
+\end{equation}
+It will be crucial for us to bound
+\(\partial^K b_0^{(0)} = \partial^{K - 1} b_0^{(0)} -
+\widehat{\partial^{K - 1} b}_0^{(0)}\). In addition, note that
+\begin{align}
+&\widehat{b}_0^{(0)} \oplus
+\widehat{\partial b}_0^{(0)} \oplus \cdots
+\oplus \widehat{\partial^{K - 2} b}_0^{(0)} \oplus
+\widehat{\partial^{K - 1} b}_0^{(0)} \\
+&= \widehat{b}_0^{(0)} \left(1 + \theta_{K - 1}\right) +
+\widehat{\partial b}_0^{(0)} \left(1 + \theta_{K - 1}\right) + \cdots
++ \widehat{\partial^{K - 2} b}_0^{(0)} \left(1 + \theta_2\right) +
+\widehat{\partial^{K - 1} b}_0^{(0)} \left(1 + \theta_1\right).
+\end{align}
+We'll use these two quantities to prove a bound.
+\end{proof}
+
 \section{Numerical experiments}\label{sec:numerical}
 
-Figure~\ref{fig:smooth-drawing} and Figure~\ref{fig:compensated-k}
-show how things improve.
+All experiments were performed in IEEE-754 double precision.
+As in \cite{Jiang2010}, we consider the evaluation in the neighborhood
+of the multiple root of \(p(s) = (s - 1)\left(s - \frac{3}{4}\right)^7\),
+written in Bernstein form.
 
@@
 
+Figure~\ref{fig:smooth-drawing} shows the evaluation of \(p(s)\) at
+the 401 equally spaced\footnote{It's worth noting that \(0.1\) cannot
+be represented exactly in IEEE-754 double precision (or any binary
+arithmetic for that matter). Hence (most of) the points of the form
+\(a + b \cdot 10^{-c}\) can only be approximately represented.} points
+\(\left\{\frac{3}{4} + j \frac{10^{-7}}{2}\right\}_{j=-200}^{200}\)
+with \texttt{DeCasteljau} (Algorithm~\ref{alg:de-casteljau}),
+\texttt{CompDeCasteljau} (Algorithm~\ref{alg:comp-de-casteljau})
+and \texttt{CompDeCasteljau3} (Algorithm~\ref{alg:k-comp-de-casteljau}
+with \(K = 3\)). We see that \texttt{DeCasteljau} fails to get the
+magnitude correct, \texttt{CompDeCasteljau} has the right shape but
+lots of noise and \texttt{CompDeCasteljau3} is able to smoothly evaluate
+the function. This is in contrast to a similar figure in \cite{Jiang2010},
+where the plot was smooth for the 400 equally spaced points
+\(\left\{\frac{3}{4} + \frac{10^{-4}}{2} \frac{2j - 399}{399}
+\right\}_{j=0}^{399}\). The primary difference is that as the interval
+shrinks by a factor of \(\approx \frac{10^{-4}}{10^{-7}} = 10^3\), the
+condition number goes up by \(\approx 10^{21}\) and \texttt{CompDeCasteljau}
+is no longer accurate.
+
 \begin{figure}
@@
 \end{figure}
+
+Figure~\ref{fig:compensated-k} shows the relative forward errors compared
+against the condition number. To compute relative errors, each input and
+coefficient is converted to a fraction (i.e. infinite precision) and then
+compared to the corresponding computed values. Similar tools are used to
+\textbf{exactly} compute the condition number, though here we can rely
+on the fact that \(\widetilde{p}\left(s\right) = (s - 1)
+\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
+condition numbers are computed as fractions, they are rounded to the
+nearest IEEE-754 double precision value. As in \cite{Jiang2010}, we use
+values \(\left\{\frac{3}{4} - (1.3)^j\right\}_{j=-5}^{-90}\)\footnote{As with
+\(0.1\), it's worth noting that \((1.3)^j\) can't be represented exactly in
+IEEE-754 double precision. However, this geometric series still serves a
+useful purpose since it continues to raise \(\cond{p, s}\) as \(j\) decreases
+away from \(0\) and because it results in ``random'' changes in the bits of
+\(0.75\) that are impacted by subtracting \((1.3)^j\).}. The curves for
+\texttt{DeCasteljau} and \texttt{CompDeCasteljau} trace the same paths seen
+in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
+error that is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+\(1 / \mach\), at which point the relative error increases linearly with
+the condition number until it becomes \(\mathcal{O}(1)\) when
+\(\cond{p, s}\) reaches \(1 / \mach^2\).
+Similarly, the relative error in \texttt{CompDeCasteljau3}
+(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
+is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+\(1 / \mach^2\) at which point the relative error increases linearly
+to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
+and the relative error in \texttt{CompDeCasteljau4}
+(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
+is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+\(1 / \mach^3\) at which point the relative error increases linearly
+to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).
 

### 2018-05-02 scripts/error_against_cond.py

@@ -42,6 +42,7 @@ BEZIER_COEFFS = (
     0.0,
 )
 ROOT = 0.75
+POWER_VAL = 1.3
 
 
 def main(filename=None):
@@ -57,7 +58,7 @@ def main(filename=None):
     forward_errs3 = []
     forward_errs4 = []
     for j in range(-5, -90 - 1, -1):
-        s = ROOT - 1.3 ** j
+        s = ROOT - POWER_VAL ** j
         exact_s = F(s)
 
         # Compute the condition number.
@@ -249,7 +250,7 @@ def main_jlcs10(filename=None):
     forward_errs1 = []
     forward_errs2 = []
     for j in range(-5, -64 - 1, -1):
-        s = ROOT - 1.3 ** j
+        s = ROOT - POWER_VAL ** j
         exact_s = F(s)
 
         # Compute the condition number.

### 2018-05-02 scripts/smooth_drawing.py

@@ -38,7 +38,7 @@ BEZIER_COEFFS = (
 )
 ROOT = 0.75
 DELTA_S = 1e-5
-NUM_POINTS = 400
+NUM_POINTS = 401
 
 
 def main(filename=None):

### 2018-05-02 doc/paper.tex
Adding commands for the "partial b" terms.
Also

- Adding subsections to the K-compensated section
- Dropping some unnecessary parentheses
- Changing some terms in the proof sketch of the error
  bound

@@
 \newcommand{\mach}{\mathbf{u}}
+%% db == ``delta'' b
+\newcommand{\db}[1]{
+  \ifthenelse{\equal{#1}{1}}
+             {\partial b}
+             {\partial^{#1} b}
+}
+%% cdb == computed ``delta'' b
+\newcommand{\cdb}[1]{
+  \ifthenelse{\equal{#1}{1}}
+             {\widehat{\partial b}}
+             {\widehat{\partial^{#1} b}}
+}
 
@@
       n \mach / (1 - n \mach)\),
-  \item \((1 + \theta_k)(1 + \theta_j) \leq (1 + \theta_{k + j})\),
+  \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
   \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}\),
@@
 \begin{equation}
-  \partial b_j^{(k)} = b_j^{(k)} - \widehat{b}_j^{(k)}
+  \db{1}_j^{(k)} = b_j^{(k)} - \widehat{b}_j^{(k)}
 \end{equation}
@@
 accumulate in
-\(\partial b^{(k)}\):
+\(\db{1}^{(k)}\):
 \begin{equation}\label{err-update}
-  \partial b_j^{(k)} = (1 - s) \cdot \partial b_j^{(k + 1)} + s \cdot
-  \partial b_{j + 1}^{(k + 1)} + \ell_{1, j}^{(k)}.
+  \db{1}_j^{(k)} = (1 - s) \cdot \db{1}_j^{(k + 1)} + s \cdot
+  \db{1}_{j + 1}^{(k + 1)} + \ell_{1, j}^{(k)}.
 \end{equation}
@@
 \begin{equation}
-  p(s) = \widehat{b}_0^{(0)} + \partial b_0^{(0)}
+  p(s) = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
 \end{equation}
 and by using \eqref{err-update}, we can continue to compute
-approximations of \(\partial b_j^{(k)}\). The idea behind the compensated
+approximations of \(\db{1}_j^{(k)}\). The idea behind the compensated
 de Casteljau algorithm is to compute both the local error and the updates
@@
         \State \(\widehat{b}_j^{(n)} = b_j\)
-        \State \(\widehat{\partial b}_j^{(n)} = 0\)
+        \State \(\cdb{1}_j^{(n)} = 0\)
       \EndFor
@@
               \widehat{b}_j^{(k + 1)}\right)\)
-          \State \(\widehat{\partial b}_j^{(k)} =
+          \State \(\cdb{1}_j^{(k)} =
               \left(\widehat{r} \otimes
-              \widehat{\partial b}_j^{(k + 1)}\right) \oplus
-              \left(s \otimes \widehat{\partial b}_{j + 1}^{(k + 1)}
+              \cdb{1}_j^{(k + 1)}\right) \oplus
+              \left(s \otimes \cdb{1}_{j + 1}^{(k + 1)}
               \right) \oplus
@@
       \State \(\mathtt{res} = \widehat{b}_0^{(0)} \oplus
-          \widehat{\partial b}_0^{(0)}\)
+          \cdb{1}_0^{(0)}\)
     \EndFunction
@@
   \begin{equation}
-    \left|\partial b_0^{(0)} - \widehat{\partial b}_0^{(0)}\right| \leq
+    \left|\db{1}_0^{(0)} - \cdb{1}_0^{(0)}\right| \leq
     2 \gamma_{3n + 1} \gamma_{3(n - 1)}
-    \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
+    \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) =
+    2 \gamma_{3n + 1} \gamma_{3(n - 1)} \widetilde{p}(s).
   \end{equation}
@@
 
-\noindent Using the bound on the error in \(\partial b\), the algorithm can
+\noindent Using the bound on the error in \(\db{1}\), the algorithm can
 be shown to be as accurate as if the computations were done in twice
@@
 Table~\ref{tab:exact-computation}, the breakdown occurs because
-\(\widehat{b}_0^{(0)} = -\widehat{\partial b}_0^{(0)} = \mach / 16\).
+\(\widehat{b}_0^{(0)} = -\cdb{1}_0^{(0)} = \mach / 16\).
 
@@
     \toprule
-    k & j & \widehat{b}_j^{(k)} & \widehat{\partial b}_j^{(k)} & \partial b_j^{(k)} - \widehat{\partial b}_j^{(k)} \\
+    k & j & \widehat{b}_j^{(k)} & \cdb{1}_j^{(k)} & \db{1}_j^{(k)} - \cdb{1}_j^{(k)} \\
     \midrule
@@
 
+\subsection{Algorithm Specified}
+
 In order to raise from twice the working precision to \(K\) times the
 working precision, we continue using EFTs when computing
-\(\widehat{\partial b}^{(k)}\). By tracking the round-off from each
+\(\cdb{1}^{(k)}\). By tracking the round-off from each
 floating point evaluation via an EFT, we can form a cascade of global errors:
 \begin{align}
-  b_j^{(k)} &= \widehat{b}_j^{(k)} + \partial b_j^{(k)} \\
-  \partial b_j^{(k)} &= \widehat{\partial b}_j^{(k)} + \partial^2 b_j^{(k)} \\
-  \partial^2 b_j^{(k)} &= \widehat{\partial^2 b}_j^{(k)} +
-  \partial^3 b_j^{(k)} \\
+  b_j^{(k)} &= \widehat{b}_j^{(k)} + \db{1}_j^{(k)} \\
+  \db{1}_j^{(k)} &= \cdb{1}_j^{(k)} + \db{2}_j^{(k)} \\
+  \db{2}_j^{(k)} &= \cdb{2}_j^{(k)} +
+  \db{3}_j^{(k)} \\
   & \vdots \nonumber
@@
   \widehat{b}_j^{(k)} &&+ \ell_{1, j}^{(k)} \\
-  (1 - s) \cdot \widehat{\partial b}_j^{(k + 1)} &+
-  s \cdot \widehat{\partial b}_{j + 1}^{(k + 1)} &&+ \ell_{1, j}^{(k)} &&=
-  \widehat{\partial b}_j^{(k)} &&+ \ell_{2, j}^{(k)} \\
-  (1 - s) \cdot \widehat{\partial^2 b}_j^{(k + 1)} &+
-  s \cdot \widehat{\partial^2 b}_{j + 1}^{(k + 1)} &&+ \ell_{2, j}^{(k)} &&=
-  \widehat{\partial^2 b}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
+  (1 - s) \cdot \cdb{1}_j^{(k + 1)} &+
+  s \cdot \cdb{1}_{j + 1}^{(k + 1)} &&+ \ell_{1, j}^{(k)} &&=
+  \cdb{1}_j^{(k)} &&+ \ell_{2, j}^{(k)} \\
+  (1 - s) \cdot \cdb{2}_j^{(k + 1)} &+
+  s \cdot \cdb{2}_{j + 1}^{(k + 1)} &&+ \ell_{2, j}^{(k)} &&=
+  \cdb{2}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
   &  &&  &&\vdots && \nonumber
@@
     \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
-    \node [box, right of=bn0] (dbn0) {\(\widehat{\partial b}^{(n)}\)};
-    \node [box, right of=dbn0] (ddbn0) {\(\widehat{\partial^2 b}^{(n)}\)};
-    \node [box, right of=ddbn0] (dddbn0) {\(\widehat{\partial^3 b}^{(n)}\)};
+    \node [box, right of=bn0] (dbn0) {\(\cdb{1}^{(n)}\)};
+    \node [box, right of=dbn0] (ddbn0) {\(\cdb{2}^{(n)}\)};
+    \node [box, right of=ddbn0] (dddbn0) {\(\cdb{3}^{(n)}\)};
     %%
     \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
-    \node [box, below of=dbn0] (dbn1) {\(\widehat{\partial b}^{(n - 1)}\)};
-    \node [box, below of=ddbn0] (ddbn1) {\(\widehat{\partial^2 b}^{(n - 1)}\)};
-    \node [box, below of=dddbn0] (dddbn1) {\(\widehat{\partial^3 b}^{(n - 1)}\)};
+    \node [box, below of=dbn0] (dbn1) {\(\cdb{1}^{(n - 1)}\)};
+    \node [box, below of=ddbn0] (ddbn1) {\(\cdb{2}^{(n - 1)}\)};
+    \node [box, below of=dddbn0] (dddbn1) {\(\cdb{3}^{(n - 1)}\)};
     %%
     \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
-    \node [box, below of=dbn1] (dbn2) {\(\widehat{\partial b}^{(n - 2)}\)};
-    \node [box, below of=ddbn1] (ddbn2) {\(\widehat{\partial^2 b}^{(n - 2)}\)};
-    \node [box, below of=dddbn1] (dddbn2) {\(\widehat{\partial^3 b}^{(n - 2)}\)};
+    \node [box, below of=dbn1] (dbn2) {\(\cdb{1}^{(n - 2)}\)};
+    \node [box, below of=ddbn1] (ddbn2) {\(\cdb{2}^{(n - 2)}\)};
+    \node [box, below of=dddbn1] (dddbn2) {\(\cdb{3}^{(n - 2)}\)};
     %%
@@
 a single stage of error filtering we
-``give up'' and use \(\widehat{\partial b}\) instead of
-\(\partial b\) (without keeping around any information about the
+``give up'' and use \(\cdb{1}\) instead of
+\(\db{1}\) (without keeping around any information about the
 round-off error). In order to obtain results that are as accurate as if
@@
 the rounded
-\(\widehat{\partial^{K - 1} b}\) in place of the exact
-\(\partial^{K - 1} b\).
+\(\cdb{K - 1}\) in place of the exact
+\(\db{K - 1}\).
 
-When computing \(\widehat{\partial^F b}\) (i.e. the error after
+When computing \(\cdb{F}\) (i.e. the error after
 \(F\) stages of filtering)
@@
 \item an error
-for the ``missing'' \(\rho \cdot \widehat{\partial^F b}_j^{(k + 1)}\) in
-\((1 - s) \cdot \widehat{\partial^F b}_j^{(k + 1)}\)
+for the ``missing'' \(\rho \cdot \cdb{F}_j^{(k + 1)}\) in
+\((1 - s) \cdot \cdb{F}_j^{(k + 1)}\)
 \item an error from the product
-  \(\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\)
+  \(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\)
 \item an error from the product
-  \(s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\)
+  \(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\)
 \item two errors from the two \(\oplus\) when combining the three
   terms in
-  \(\left[\widehat{r} \otimes \widehat{\partial^F b}_j^{(k + 1)}\right] \oplus
-  \left[s \otimes \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right] \oplus
+  \(\left[\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right] \oplus
+  \left[s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right] \oplus
   \widehat{\ell}_{F, j}^{(k)}\)
@@
 \ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
-\widehat{\partial^{F - 1} b}_j^{(k + 1)}
+\cdb{F - 1}_j^{(k + 1)}
 \end{equation}
 where the terms \(e_1, \ldots, e_M\) come from using \texttt{TwoSum} and
-\texttt{TwoProd} when computing \(\widehat{\partial^{F - 1} b}_j^{(k)}\)
+\texttt{TwoProd} when computing \(\cdb{F - 1}_j^{(k)}\)
 and the \(\rho\) term comes from the roundoff
 in \(1 \ominus s\) when multiplying \((1 - s)\) by
-\(\widehat{\partial^{F - 1} b}_j^{(k + 1)}\). With this in mind, we
+\(\cdb{F - 1}_j^{(k + 1)}\). With this in mind, we
 can define an EFT (\texttt{LocalErrorEFT}) that computes
@@
         \For{\(F = 1, \ldots, K - 1\)}
-          \State \(\widehat{\partial^F b}_j^{(n)} = 0\)
+          \State \(\cdb{F}_j^{(n)} = 0\)
         \EndFor
@@
             \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
-                \widehat{r}, \widehat{\partial^F b}_j^{(k + 1)}\right)\)
+                \widehat{r}, \cdb{F}_j^{(k + 1)}\right)\)
             \State \(\left[P_2, \eta_{M + 2}\right] = \mathtt{TwoProd}\left(
-                s, \widehat{\partial^F b}_{j + 1}^{(k + 1)}\right)\)
+                s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
             \State \(\left[S_3, \eta_{M + 3}\right] =
                 \mathtt{TwoSum}(P_1, P_2)\)
-            \State \(\left[\widehat{\partial^F b}_j^{(k)}, \eta_{M + 4}\right]
+            \State \(\left[\cdb{F}_j^{(k)}, \eta_{M + 4}\right]
                 = \mathtt{TwoSum}\left(S_3, \widehat{\ell}\right)\)
@@
             \State \(e = \eta\)
-            \State \(\delta b = \widehat{\partial^F b}_j^{(k + 1)}\)
+            \State \(\delta b = \cdb{F}_j^{(k + 1)}\)
           \EndFor
@@
                 \mathtt{LocalError}(e, \rho, \delta b)\)
-          \State \(\widehat{\partial^{K - 1} b}_j^{(k)} =
+          \State \(\cdb{K - 1}_j^{(k)} =
               \left(\widehat{r} \otimes
-              \widehat{\partial^{K - 1} b}_j^{(k + 1)}\right) \oplus
-              \left(s \otimes \widehat{\partial^{K - 1} b}_{j + 1}^{(k + 1)}
+              \cdb{K - 1}_j^{(k + 1)}\right) \oplus
+              \left(s \otimes \cdb{K - 1}_{j + 1}^{(k + 1)}
               \right) \oplus
@@
         \State \(\mathtt{res} = \mathtt{res} \oplus
-            \widehat{\partial^F b}_0^{(0)}\)
+            \cdb{F}_0^{(0)}\)
       \EndFor
@@
 
+\subsection{Error bound for polynomial evaluation}
+
 \begin{theorem}
@@
 \begin{equation}
-p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \partial b_0^{(0)}
+p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
 = \cdots
-= \widehat{b}_0^{(0)} + \widehat{\partial b}_0^{(0)} + \cdots
-+ \widehat{\partial^{K - 2} b}_0^{(0)} + \partial^{K - 1} b_0^{(0)}
+= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
++ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
 \end{equation}
@@
 \mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
-\widehat{\partial b}_0^{(0)} \oplus \cdots
-\oplus \widehat{\partial^{K - 2} b}_0^{(0)} \oplus
-\widehat{\partial^{K - 1} b}_0^{(0)}.
+\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}.
 \end{equation}
-It will be crucial for us to bound
-\(\partial^K b_0^{(0)} = \partial^{K - 1} b_0^{(0)} -
-\widehat{\partial^{K - 1} b}_0^{(0)}\). In addition, note that
+It will be crucial for us to bound \(\db{K}_0^{(0)}\).
+In addition, note that
 \begin{align}
-&\widehat{b}_0^{(0)} \oplus
-\widehat{\partial b}_0^{(0)} \oplus \cdots
-\oplus \widehat{\partial^{K - 2} b}_0^{(0)} \oplus
-\widehat{\partial^{K - 1} b}_0^{(0)} \\
+&\widehat{b}_0^{(0)} \oplus \cdb{1}_0^{(0)} \oplus \cdots \oplus
+\cdb{K - 1}_0^{(0)} \\
 &= \widehat{b}_0^{(0)} \left(1 + \theta_{K - 1}\right) +
-\widehat{\partial b}_0^{(0)} \left(1 + \theta_{K - 1}\right) + \cdots
-+ \widehat{\partial^{K - 2} b}_0^{(0)} \left(1 + \theta_2\right) +
-\widehat{\partial^{K - 1} b}_0^{(0)} \left(1 + \theta_1\right).
+\cdb{1}_0^{(0)} \left(1 + \theta_{K - 1}\right) + \cdots
++ \cdb{K - 1}_0^{(0)} \left(1 + \theta_1\right).
 \end{align}

### 2018-05-02 doc/paper.tex
Changed error update order of ops to match JLCS10.
This is for good reason, they put (rhat x db_j) at the
end to avoid accruing an extra round-off error for rhat.

Also increased `NUM_POINTS` from 400 to 401 in all remaining
places.

Note that the order of operations change has changed all plots
that use compensated de Casteljau (but not in a qualitative way).

I also went back and double-checked that `Table 1` was still
correct after the order of operations change.

@@
 \(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
-\(\mach = 2^{-53}\).) We also assume that the computed result of
-\(\alpha \in \mathbf{R}\) in floating point arithmetic is denoted by
+\(\mach = 2^{-53}\).) We denote the computed result of
+\(\alpha \in \mathbf{R}\) in floating point arithmetic by
 \(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
@@
 For a pair floating point numbers \(a, b \in \mathbf{F}\), there exists
-a floating point number \(y\) satisfying \(a \circ b = x + y\) where
+\(y \in \mathbf{F}\) satisfying \(a \circ b = x + y\) where
 \(x = \fl{a \circ b}\) and \(\circ \in \left\{+, -, \times\right\}\).
@@
 \end{equation}
-When computed in exact arithmetic, we would have
+When computed in exact arithmetic
 \begin{equation}
@@
           \State \(\cdb{1}_j^{(k)} =
-              \left(\widehat{r} \otimes
-              \cdb{1}_j^{(k + 1)}\right) \oplus
+              \widehat{\ell}_{1, j}^{(k)} \oplus
               \left(s \otimes \cdb{1}_{j + 1}^{(k + 1)}
               \right) \oplus
-              \widehat{\ell}_{1, j}^{(k)}\)
+              \left(\widehat{r} \otimes
+              \cdb{1}_j^{(k + 1)}\right)\)
         \EndFor
@@
   \begin{equation}
-    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq \mach +
-    2 \gamma_{3n}^2 \cond{p, s}.
+    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
+      \left|p(s)\right|} \leq \mach + 2 \gamma_{3n}^2 \cond{p, s}.
   \end{equation}
@@
 Unfortunately, Figure~\ref{fig:compensated-insufficient} shows how
-the \texttt{CompDeCasteljau} starts to break down in a region of
+\texttt{CompDeCasteljau} starts to break down in a region of
 high condition number (caused by a multiple root with multiplicity
@@
 --- which is in the plotted region \(\left|s - \frac{1}{2}\right|
-\leq 1.5 \cdot 10^{-11}\) --- evaluates to exactly \(0\) when it should be
-\(\mathcal{O}\left(\mach^3\right)\). As shown in
+\leq \frac{3}{2} \cdot 10^{-11}\) --- evaluates to exactly \(0\) when
+it should be \(\mathcal{O}\left(\mach^3\right)\). As shown in
 Table~\ref{tab:exact-computation}, the breakdown occurs because
@@
   terms in
-  \(\left[\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right] \oplus
-  \left[s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right] \oplus
-  \widehat{\ell}_{F, j}^{(k)}\)
+  \(\widehat{\ell}_{F, j}^{(k)} \oplus
+  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
+  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right)\)
 \end{itemize}
@@
 
-\noindent With this EFT helper in place, we can perform \((K - 1)\)
+\noindent With this EFT in place, we can perform \((K - 1)\)
 error filtrations:
@@
             \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
+                s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
+            \State \(\left[S_2, \eta_{M + 2}\right] =
+                \mathtt{TwoSum}\left(\widehat{\ell}, P_1\right)\)
+            \State \(\left[P_3, \eta_{M + 3}\right] = \mathtt{TwoProd}\left(
                 \widehat{r}, \cdb{F}_j^{(k + 1)}\right)\)
-            \State \(\left[P_2, \eta_{M + 2}\right] = \mathtt{TwoProd}\left(
-                s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
-            \State \(\left[S_3, \eta_{M + 3}\right] =
-                \mathtt{TwoSum}(P_1, P_2)\)
             \State \(\left[\cdb{F}_j^{(k)}, \eta_{M + 4}\right]
-                = \mathtt{TwoSum}\left(S_3, \widehat{\ell}\right)\)
+                = \mathtt{TwoSum}\left(S_2, P_3\right)\)
             \\
@@
           \State \(\cdb{K - 1}_j^{(k)} =
-              \left(\widehat{r} \otimes
-              \cdb{K - 1}_j^{(k + 1)}\right) \oplus
+              \widehat{\ell} \oplus
               \left(s \otimes \cdb{K - 1}_{j + 1}^{(k + 1)}
               \right) \oplus
-              \widehat{\ell}\)
+              \left(\widehat{r} \otimes
+              \cdb{K - 1}_j^{(k + 1)}\right)\)
         \EndFor
@@
 We'll use these two quantities to prove a bound.
+
+First, note that \(\widehat{r} = (1 - s)(1 + \theta_1)\). Due to this,
+we put the \(\widehat{r}\) term at the end of each update step to reduce
+the amount of round-off:
+\begin{align}
+  \cdb{F}_j^{(k)} &=
+  \widehat{\ell}_{F, j}^{(k)} \oplus
+  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
+  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right) \\
+&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
+  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
+  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2) \\
+\Longrightarrow \cdb{F}^{(k)} &=
+  U_{k + 1}(s) \cdb{F}^{(k + 1)}(1 + \theta_3) +
+  \widehat{\ell}_{F}^{(k)} (1 + \theta_2).
+\end{align}
+This is quite similar to the relationship satisfied in
+exact arithmetic:
+\begin{equation}
+\db{F}^{(k)} = U_{k + 1}(s) \db{F}^{(k + 1)} + \ell_{F}^{(k)}.
+\end{equation}
+
 \end{proof}

### 2018-05-02 scripts/compensated_insufficient.py

@@ -27,7 +27,7 @@ F = fractions.Fraction
 BEZIER_COEFFS = (1.0, -0.75, 0.5, -0.25, 0.0)
 ROOT = 0.5
 DELTA_S = 1.5e-11
-NUM_POINTS = 400
+NUM_POINTS = 401
 
 
 def main(filename=None):

### 2018-05-02 scripts/error_against_cond.py

@@ -261,7 +261,7 @@ def main_jlcs10(filename=None):
         cond_nums.append(float(exact_cond))
 
         # Compute the forward error for uncompensated de Casteljau.
-        b, db = de_casteljau.jlcs10_compensated(s, BEZIER_COEFFS)
+        b, db = de_casteljau._compensated_k(s, BEZIER_COEFFS, 2)
         exact_b1 = F(b)
         exact_forward_err1 = abs((exact_b1 - exact_p) / exact_p)
         forward_errs1.append(float(exact_forward_err1))

### 2018-05-02 scripts/horner_inferior.py

@@ -29,7 +29,7 @@ POLY_COEFFS = (8.0, -12.0, 6.0, -1.0)
 BEZIER_COEFFS = (-1.0, 1.0, -1.0, 1.0)
 ROOT = 0.5
 DELTA_S = 5e-6
-NUM_POINTS = 400
+NUM_POINTS = 401
 
 
 def main(filename=None):

### 2018-05-02 src/de_casteljau.py

@@ -56,51 +56,6 @@ def basic(s, coeffs):
     return pk[0]
 
 
-def jlcs10_compensated(s, coeffs):
-    r"""Performs the compensated de Casteljau algorithm.
-
-    .. _JLCS10: https://doi.org/10.1016/j.camwa.2010.05.021
-
-    This is similar to :func:`compensated` below, but the order of
-    operations exactly matches the `JLCS10`_ paper. In particular,
-    :math:`\widehat{\partial b}_j^{(k)}` is computed as
-
-    .. math::
-
-        \widehat{ell}_{i, j}^{(k)} \oplus \left[s \otimes
-            \widehat{\partial b}_{j + 1}^{(k + 1)}\right] \oplus
-            \left[\widehat{r} \otimes \widehat{\partial b}_j^{(k + 1)}\right]
-
-    and the code in :func:`_compensated_k` uses the "typical" order
-
-    .. math::
-
-        \left[\widehat{r} \otimes \widehat{\partial b}_j^{(k + 1)}\right]
-            \oplus \left[s \otimes \widehat{\partial b}_{j + 1}^{(k + 1)}
-            \right] \oplus \widehat{ell}_{i, j}^{(k)}.
-    """
-    r, rho = eft.add_eft(1.0, -s)
-
-    degree = len(coeffs) - 1
-    bk = {0: list(coeffs), 1: (0.0,) * (degree + 1)}
-    for k in range(degree):
-        new_bk = {0: [], 1: []}
-
-        for j in range(degree - k):
-            P1, pi1 = eft.multiply_eft(r, bk[0][j])
-            P2, pi2 = eft.multiply_eft(s, bk[0][j + 1])
-            S3, sigma3 = eft.add_eft(P1, P2)
-            new_bk[0].append(S3)
-            # Now update the error.
-            l_hat = pi1 + pi2 + sigma3 + rho * bk[0][j]
-            new_bk[1].append(l_hat + s * bk[1][j + 1] + r * bk[1][j])
-
-        # Update the "current" values.
-        bk = new_bk
-
-    return bk[0][0], bk[1][0]
-
-
 def local_error(errors, rho, delta_b):
     r"""Compute :math:`\ell` from a list of errors.
 
@@ -138,6 +93,36 @@ def local_error_eft(errors, rho, delta_b):
 
 
 def _compensated_k(s, coeffs, K):
+    r"""Performs a K-compensated de Casteljau.
+
+    .. _JLCS10: https://doi.org/10.1016/j.camwa.2010.05.021
+
[…]

### 2018-05-02 doc/paper.tex
Adding two lemmas from JLCS10 about the error bound.
Also

- Making numbering of most things (everything except footnotes)
  be section specific
- Making vdots be horizontally aligned with equals sign
- Adding a footnote about the change of notation for the terms
  generated by the de Casteljau algorithm
- Removing a theorem in the "review" (Section 3) because it is
  now in subsection 4.2 (where I am attempting to prove / compute
  an a priori error bound)

@@
 \makeatother
+%% H/T: https://tex.stackexchange.com/a/28334/32270
+\usepackage{chngcntr}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+\counterwithin{equation}{section}
+\counterwithin{algorithm}{section}
 %% H/T: https://tex.stackexchange.com/a/202047/32270
@@
 \theoremstyle{definition}
-\newtheorem{theorem}{Theorem}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
 
@@
 
-Next, we recall the de Casteljau algorithm:
+Next, we recall\footnote{We have used slightly non-standard notation for the
+terms produced by the de Casteljau algorithm: we start the superscript at
+\(n\) and count down to \(0\) as is typically done when describing Horner's
+algorithm. For example, we use \(b_j^{(n - 2)}\) instead of
+\(b_j^{(2)}\).} the de Casteljau algorithm:
 
@@
 \begin{equation}
-\cond{p, s} = \frac{\widetilde{p}\left(s\right)}{
+\cond{p, s} = \frac{\widetilde{p}(s)}{
   \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
@@
 difference depends only on \(s\) and the Bernstein
-coefficients of \(p\):
-
-\begin{theorem}[\cite{Jiang2010}]
-  If no underflow occurs and \(n \geq 2\), then
-  \begin{equation}
-    \left|\db{1}_0^{(0)} - \cdb{1}_0^{(0)}\right| \leq
-    2 \gamma_{3n + 1} \gamma_{3(n - 1)}
-    \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) =
-    2 \gamma_{3n + 1} \gamma_{3(n - 1)} \widetilde{p}(s).
-  \end{equation}
-\end{theorem}
-
-\noindent Using the bound on the error in \(\db{1}\), the algorithm can
+coefficients of \(p\). Using a bound (Lemma~\ref{lemma:db-lemma}) on the
+round-off error when computing \(\db{1}^{(0)}\), the algorithm can
 be shown to be as accurate as if the computations were done in twice
@@
   \db{3}_j^{(k)} \\
-  & \vdots \nonumber
+  %% H/T: https://tex.stackexchange.com/a/7651/32270
+  &\mathrel{\makebox[\widthof{=}]{\vdots}} \nonumber
 \end{align}
@@
   \cdb{2}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
-  &  &&  &&\vdots && \nonumber
+  &  &&  &&\mathrel{\makebox[\widthof{=}]{\vdots}} && \nonumber
 \end{alignat}
@@
 
-Noting that \(\ell_{F, j}\) contains \(5F - 2\) terms, one can
-show that Algorithm~\ref{alg:k-comp-de-casteljau} requires
+\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
+show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
+requires
 \begin{equation}
@@
 \subsection{Error bound for polynomial evaluation}
+
+\begin{lemma}[\cite{Jiang2010}]
+For \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) a polynomial in Bernstein form
+and \(s \in \left[0, 1\right]\), the terms produced by the de Casteljau
+algorithm in exact arithmetic satisfy:
+\begin{equation}
+  \left|b_0^{(0)}\right| \leq \cdots \leq
+  \sum_{j = 0}^k \left|b_j^{(k)}\right| B_{j, k}(s) \leq \cdots \leq
+  \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) = \widetilde{p}(s).
+\end{equation}
+\end{lemma}
+
+\begin{lemma}[\cite{Jiang2010}]\label{lemma:db-lemma}
+The second order error \(\db{2}^{(0)}_0\) satisfies
+\begin{equation}
+  \left|\db{1}^{(0)}_0 - \cdb{1}^{(0)}_0\right| =
+  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 1} \gamma_{3(n - 1)}
+  \widetilde{p}(s).
+\end{equation}
+\end{lemma}
 
@@
 \textbf{exactly} compute the condition number, though here we can rely
-on the fact that \(\widetilde{p}\left(s\right) = (s - 1)
+on the fact that \(\widetilde{p}(s) = (s - 1)
 \left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and

### 2018-05-03 doc/paper.tex
Adding some more "insight" to the proof.
I used code to prove the claim to myself about how
the l-tilde terms relate to each other. We can actually
be a bit more specific. For example, write

   l1 = e1 + ... + e3 + rho b
   l2 = f1 + ... + f8 + rho db

then we can actually show that

   tilde(l2) <= gamma5 e1 + gamma5 e2 + gamma4 e3 +
                gamma4 (rho b) +
                gamma3[(1 - s) db{j} + s db{j + 1}].

To extend the example, write

   l3 = g1 + ... + g13 + rho d2b

then we can actually show that

   tilde(l3) <= gamma10 f1 + gamma10 f2 + gamma9 f3 + gamma8 f4 +

@@
 \usepackage{amsthm,amssymb,amsmath}
+%% https://tex.stackexchange.com/a/396829/32270
+\allowdisplaybreaks
 %% H/T: https://tex.stackexchange.com/a/202168/32270
@@
 errors came from sum and product, respectively. See the appendix
-in Section~\ref{sec:appendix-eft} for implementation details.
+in Section~\ref{sec:appendix-algo} for implementation details.
 \end{theorem}
@@
 \begin{equation}
-\cond{p, s} = \frac{\widetilde{p}(s)}{
-  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
-  \left|p(s)\right|},
-\end{equation}
-where \(B_{j, n} \geq 0\).
+\cond{p, s} = \frac{\widetilde{p}(s)}{\left|p(s)\right|},
+\end{equation}
+where \(B_{j, n}(s) \geq 0\) and
+\(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).
 
@@
 \begin{equation}
-\ell_{F, j}^{(k)} = e_1 + e_2 + \cdots + e_M + \rho \cdot
+\ell_{F, j}^{(k)} = e_1 + \cdots + e_{5F - 2} + \rho \cdot
 \cdb{F - 1}_j^{(k + 1)}
 \end{equation}
-where the terms \(e_1, \ldots, e_M\) come from using \texttt{TwoSum} and
-\texttt{TwoProd} when computing \(\cdb{F - 1}_j^{(k)}\)
+where the terms \(e_1, \ldots, e_{5F - 2}\) come from using \texttt{TwoSum}
+and \texttt{TwoProd} when computing \(\cdb{F - 1}_j^{(k)}\)
 and the \(\rho\) term comes from the roundoff
@@
   \caption{\textit{EFT for computing the local error.}}
+  \label{alg:local-error-eft}
 
@@
         \mathtt{LocalErrorEFT}\)}{$e, \rho, \delta b$}
-      \State \(M = \texttt{length}(e)\)
+      \State \(L = \texttt{length}(e)\)
       \\
@@
           \mathtt{TwoSum}(e_1, e_2)\)
-      \For{\(j = 3, \ldots, M\)}
+      \For{\(j = 3, \ldots, L\)}
         \State \(\left[\widehat{\ell}, \eta_{j - 1}\right] =
@@
       \\
-      \State \(\left[P, \eta_M\right] =
+      \State \(\left[P, \eta_L\right] =
           \mathtt{TwoProd}\left(\rho, \delta b\right)\)
-      \State \(\left[\widehat{\ell}, \eta_{M + 1}\right] =
+      \State \(\left[\widehat{\ell}, \eta_{L + 1}\right] =
           \mathtt{TwoSum}\left(\widehat{\ell}, P\right)\)
@@
 
-\noindent With this EFT in place, we can perform \((K - 1)\)
-error filtrations:
+\noindent With this EFT in place\footnote{And the related
+\texttt{LocalError} in Algorithm~\ref{alg:local-error}}, we can
+perform \((K - 1)\) error filtrations:
 
@@
                 \mathtt{LocalErrorEFT}(e, \rho, \delta b)\)
-            \State \(M = \texttt{length}(\eta)\)
-            \\
-            \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
+            \State \(L = \texttt{length}(\eta)\)
+            \\
+            \State \(\left[P_1, \eta_{L + 1}\right] = \mathtt{TwoProd}\left(
                 s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
-            \State \(\left[S_2, \eta_{M + 2}\right] =
+            \State \(\left[S_2, \eta_{L + 2}\right] =
                 \mathtt{TwoSum}\left(\widehat{\ell}, P_1\right)\)
-            \State \(\left[P_3, \eta_{M + 3}\right] = \mathtt{TwoProd}\left(
+            \State \(\left[P_3, \eta_{L + 3}\right] = \mathtt{TwoProd}\left(
                 \widehat{r}, \cdb{F}_j^{(k + 1)}\right)\)
-            \State \(\left[\cdb{F}_j^{(k)}, \eta_{M + 4}\right]
+            \State \(\left[\cdb{F}_j^{(k)}, \eta_{L + 4}\right]
                 = \mathtt{TwoSum}\left(S_2, P_3\right)\)
@@
 
-First, note that \(\widehat{r} = (1 - s)(1 + \theta_1)\). Due to this,
+To be able to express the algorithm in matrix form, we define
+the vectors
+\begin{gather}
+b^{(k)} = \left[\begin{array}{c c c} b_0^{(k)} & \cdots &
+b_k^{(k)}\end{array}\right]^T, \quad
+\widehat{b}^{(k)} = \left[\begin{array}{c c c} \widehat{b}_0^{(k)} & \cdots &
+\widehat{b}_k^{(k)}\end{array}\right]^T, \\
+\db{F}^{(k)} = \left[\begin{array}{c c c} \db{F}_0^{(k)} & \cdots &
+\db{F}_k^{(k)}\end{array}\right]^T, \quad
+\cdb{F}^{(k)} = \left[\begin{array}{c c c} \cdb{F}_0^{(k)} & \cdots &
+\cdb{F}_k^{(k)}\end{array}\right]^T, \\
+\ell_{F}^{(k)} = \left[\begin{array}{c c c} \ell_{F, 0}^{(k)} & \cdots &
+\ell_{F, k}^{(k)}\end{array}\right]^T, \quad
+\widehat{\ell}_{F}^{(k)} = \left[\begin{array}{c c c}
+\widehat{\ell}_{F, 0}^{(k)} & \cdots &
+\widehat{\ell}_{F, k}^{(k)}\end{array}\right]^T
+\end{gather}
+and the reduction matrices:
+\begin{equation}
+U_k = U_k(s) = \left[\begin{array}{c c c c c c}
+    1 - s  & s      & 0      & \cdots & \cdots & 0      \\
+    0      & 1 - s  & s      & \ddots &        & \vdots \\
+    \vdots & \ddots & \ddots & \ddots & \ddots & \vdots \\
+    \vdots &        & \ddots & \ddots & \ddots & 0 \\
+    0      & \cdots & \cdots & 0      & 1 - s  & s
+\end{array}\right] \in \mathbf{R}^{k \times (k + 1)}.
+\end{equation}
+With this, we can express the compensated de Casteljau algorithm as
+\begin{align}
+b^{(k)} &= U_{k + 1} b^{(k + 1)} \\
+\Longrightarrow b^{(0)} &= U_1 \cdots U_n b^{(n)} \\
+\db{F}^{(k)} &= U_{k + 1} \db{F}^{(k + 1)} + \ell_{F}^{(k)} \\
+\Longrightarrow \db{F}^{(0)} &= \sum_{k = 0}^{n - 1}
+U_1 \cdots U_k \ell_F^{(k)} = \sum_{k = 0}^{n - 1}
+\left[\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\right].
+\end{align}
+For the inexact equivalent of these things, first note that
+\(\widehat{r} = (1 - s)(1 + \theta_1)\). Due to this,
 we put the \(\widehat{r}\) term at the end of each update step to reduce
@@
 \Longrightarrow \cdb{F}^{(k)} &=
-  U_{k + 1}(s) \cdb{F}^{(k + 1)}(1 + \theta_3) +
-  \widehat{\ell}_{F}^{(k)} (1 + \theta_2).
+  U_{k + 1} \cdb{F}^{(k + 1)}(1 + \theta_3) +
+  \widehat{\ell}_{F}^{(k)} (1 + \theta_2) \\
+\Longrightarrow \cdb{F}^{(0)} &=
+  \sum_{k = 0}^{n - 1}
+  U_1 \cdots U_k \widehat{\ell}_F^{(k)} (1 + \theta_{3k - 1})
+  = \sum_{k = 0}^{n - 1}
+  \left[\sum_{j = 0}^k \widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k - 1})
+    B_{j, k}(s)\right].
 \end{align}
-This is quite similar to the relationship satisfied in
-exact arithmetic:
-\begin{equation}
-\db{F}^{(k)} = U_{k + 1}(s) \db{F}^{(k + 1)} + \ell_{F}^{(k)}.
-\end{equation}
-
+From
+\begin{align}
+\widehat{\ell}_{F, j}^{(k)} &= e_1 \oplus \cdots \oplus e_{5F - 2} \oplus
+\left(\rho \otimes \cdb{F - 1}_j^{(k + 1)}\right) \\
+&= e_1\left(1 + \theta_{5F - 2}\right) + \cdots +
+e_{5F - 2}\left(1 + \theta_2\right) +
+\rho \cdot \cdb{F - 1}_j^{(k + 1)} \left(1 + \theta_2\right) \\
+\Longrightarrow \widehat{\ell}_{F, j}^{(k)}\left(1 + \theta_{3k - 1}\right)
+&= \ell_{F, j}^{(k)} + \left[e_1 \theta_{3k + 5F - 3} + \cdots +
+e_{5F - 2} \theta_{3k + 1} +
+\rho \cdot \cdb{F - 1}_j^{(k + 1)} \theta_{3k + 1}\right] \\
+&\coloneqq \ell_{F, j}^{(k)} + d_{F, j}^{(k)}
+\end{align}
+we see that
+\begin{equation}
+\cdb{F}^{(0)} = \db{F}^{(0)} + \sum_{k = 0}^{n - 1}
+  \left[\sum_{j = 0}^k d_{F, j}^{(k)}
+B_{j, k}(s)\right]
+\end{equation}
+hence
+\begin{equation}
+\left|\db{F + 1}^{(0)}\right| =
+\left|\db{F}^{(0)} - \cdb{F}^{(0)}\right|
+\leq \sum_{j = 0}^k \left|d_{F, j}^{(k)}\right| B_{j, k}(s).
+\end{equation}
+To put a bound on this, define the following
+``tilde'' notation for a sum: if \(A = \sum_j \alpha_j\) then
+\(\widetilde{A} = \sum_j \left|\alpha_j\right|\) (similar notation is used
+for \(\widetilde{p}(s)\), which is used in the condition number).
+With this, we must have
+\begin{equation}
+\left|d_{F, j}^{(k)}\right| =
+\left|e_1 \theta_{3k + 5F - 3} + \cdots +
+e_{5F - 2} \theta_{3k + 1} +
+\rho \cdot \cdb{F - 1}_j^{(k + 1)} \theta_{3k + 1}\right|
+\leq \gamma_{3k + 5F - 3} \cdot \widetilde{\ell}_{F, j}^{(k)}.
+\end{equation}
+In addition, one can show that
+\begin{alignat}{10}
+\widetilde{\ell}_{1, j}^{(k)} &\leq
+  && &&
+  \gamma_3 \left(
+  (1 - s) \left|\widehat{b}_j\right| +
+  s \left|\widehat{b}_{j + 1}\right|\right) \\
+\widetilde{\ell}_{2, j}^{(k)} &\leq
+  \gamma_5 \cdot
+  \widetilde{\ell}_{1, j}^{(k)} &&+
+  &&
+  \gamma_3 \left(
+  (1 - s) \left|\cdb{1}_j\right| +
+  s \left|\cdb{1}_{j + 1}\right|\right) \\
+\widetilde{\ell}_{3, j}^{(k)} &\leq
+  \gamma_{10} \cdot
+  \widetilde{\ell}_{2, j}^{(k)} &&+
+  &&
+  \gamma_3 \left(
+  (1 - s) \left|\cdb{2}_j\right| +
+  s \left|\cdb{2}_{j + 1}\right|\right) \\
+\widetilde{\ell}_{F + 1, j}^{(k)} &\leq
+  \gamma_{5F} \cdot
+  \widetilde{\ell}_{F, j}^{(k)} &&+
+  &&
+  \gamma_3 \left(
+  (1 - s) \left|\cdb{F}_j\right| +
+  s \left|\cdb{F}_{j + 1}\right|\right).
+\end{alignat}
+To be continued...
 \end{proof}
@@
 
-\section{Appendix: EFT Algorithms}\label{sec:appendix-eft}
+\section{Appendix: Algorithms}\label{sec:appendix-algo}
 
@@
 
+\noindent Since the final error \(\cdb{K - 1}\) will not track the errors
+during computation, we have a non-EFT version of
+Algorithm~\ref{alg:local-error-eft}:
+
+\begin{breakablealgorithm}
+  \caption{\textit{Compute the local error (non-EFT).}}
+  \label{alg:local-error}
+
+  \begin{algorithmic}
+    \Function{\(\widehat{\ell} =
+        \mathtt{LocalError}\)}{$e, \rho, \delta b$}
+      \State \(L = \texttt{length}(e)\)
+      \\
+      \State \(\widehat{\ell} = e_1 \oplus e_2\)
+      \For{\(j = 3, \ldots, L\)}
+        \State \(\widehat{\ell} = \widehat{\ell} \oplus e_j\)
+      \EndFor
+      \\
+      \State \(\widehat{\ell} = \widehat{\ell} \oplus \left(\rho \otimes \delta b\right)\)
+    \EndFunction
+  \end{algorithmic}
+\end{breakablealgorithm}
+
 \end{document}

### 2018-05-07 doc/paper.tex
Large-scale changes.
Added a few lemmas with actual bounds and re-organized so
that the proof is mostly done, just needs a few blanks to
be filled in.

--- 2018-05-02 doc/paper.tex
@@
 
+\subsection{Floating Point and Forward Error Analysis}
+
 Throughout this paper we assume that the computation in floating point
@@
 \begin{equation}
-  a \star b = \fl{a \circ b} = (a \circ b)(1 + \eps_1) =
-  (a \circ b) / (1 + \eps_2)
+  a \star b = \fl{a \circ b} = (a \circ b)(1 + \delta_1) =
+  (a \circ b) / (1 + \delta_2)
 \end{equation}
 where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
-\in \left\{+, -, \times, \div\right\}\) and \(\left|\eps_1\right|,
-\left|\eps_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
+\in \left\{+, -, \times, \div\right\}\) and \(\left|\delta_1\right|,
+\left|\delta_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
 and \(\star\) represents the floating point computation, e.g.
@@
   \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
-  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}\),
-  \item \((1 + \mach)^j \leq 1 / (1 - j \mach)\).
+  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}
+    \Longleftrightarrow (1 + \gamma_k)(1 + \gamma_j) \leq 1 + \gamma_{k + j}\),
+  \item \((1 + \mach)^j \leq 1 / (1 - j \mach) \Longleftrightarrow
+  (1 + \mach)^j - 1 \leq \gamma_j\).
 \end{enumerate}
+
+\subsection{Error-Free Transformation}
 
@@
 errors came from sum and product, respectively. See the appendix
-in Section~\ref{sec:appendix-eft} for implementation details.
+(Section~\ref{sec:appendix-algo}) for implementation details.
 \end{theorem}
+
+\subsection{de Casteljau Algorithm}
 
@@
 \begin{equation}
-\cond{p, s} = \frac{\widetilde{p}(s)}{
-  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
-  \left|p(s)\right|},
-\end{equation}
-where \(B_{j, n} \geq 0\).
+\cond{p, s} = \frac{\widetilde{p}(s)}{\left|p(s)\right|},
+\end{equation}
+where \(B_{j, n}(s) \geq 0\) and
+\(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).
+
+To be able to express the algorithm in matrix form, we define
+the vectors
+\begin{equation}
+b^{(k)} = \left[\begin{array}{c c c} b_0^{(k)} & \cdots &
+b_k^{(k)}\end{array}\right]^T, \quad
+\widehat{b}^{(k)} = \left[\begin{array}{c c c} \widehat{b}_0^{(k)} & \cdots &
+    \widehat{b}_k^{(k)}\end{array}\right]^T
+\end{equation}
+and the reduction matrices:
+\begin{equation}
+U_k = U_k(s) = \left[\begin{array}{c c c c c c}
+    1 - s  & s      & 0      & \cdots & \cdots & 0      \\
+    0      & 1 - s  & s      & \ddots &        & \vdots \\
+    \vdots & \ddots & \ddots & \ddots & \ddots & \vdots \\
+    \vdots &        & \ddots & \ddots & \ddots & 0 \\
+    0      & \cdots & \cdots & 0      & 1 - s  & s
+\end{array}\right] \in \mathbf{R}^{k \times (k + 1)}.
+\end{equation}
+With this, we can express (\cite{Mainar1999}) the de Casteljau algorithm as
+\begin{equation}\label{matrix-de-casteljau}
+b^{(k)} = U_{k + 1} b^{(k + 1)}
+\Longrightarrow b^{(0)} = U_1 \cdots U_n b^{(n)}.
+\end{equation}
+
+In general, for a sequence \(v_0, \ldots, v_n\) we'll use the notation
+\(v = \left[\begin{array}{c c c} v_0 & \cdots &
+    v_n\end{array}\right]^T\) to indicate the vector containing all of
+the values.
 
@@
 \begin{table}
+  %% See: ``scripts/verify_table.py``.
   \centering
@@
       \\
-      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
-      \For{\(F = 1, \ldots, K - 1\)}
-        \State \(\mathtt{res} = \mathtt{res} \oplus
-            \cdb{F}_0^{(0)}\)
-      \EndFor
+      \State \(\mathtt{res} = \mathtt{SumK}\left(\left[
+        \widehat{b}_0^{(0)}, \ldots, \cdb{K - 1}_0^{(0)}\right], K\right)\)
     \EndFunction
@@
 \begin{equation}
-(15K^2 + 11K - 34)T_n + K + 5 =
+(15K^2 + 11K - 34)T_n + 6K^2 - 11K + 11 =
 \mathcal{O}\left(n^2 K^2\right)
@@
 FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
-instead, lowering the flop count by \(15(3K - 4)T_n\).
+instead, lowering the flop count by \(15(3K - 4)T_n\). Another way
+to lower the total flop count is to just use
+\(\widehat{b}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}\)
+instead of \texttt{SumK}. This will reduce the total by
+\(6(K - 1)^2\) flops, however the results produced are (empirically)
+identical. This makes sense: the whole point of \texttt{SumK}
+is to filter errors in a summation so that the final operation produces
+a sum of the form \(v_1 \oplus \cdots \oplus v_K\) where each
+term is smaller than the previous by a factor of \(\mach\). This
+property is already satisfied for the \(\cdb{F}_0^{(0)}\) so in
+practice the \(K\)-compensated summation is likely not needed.
 
@@
 
-\begin{lemma}[\cite{Jiang2010}]
-For \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) a polynomial in Bernstein form
-and \(s \in \left[0, 1\right]\), the terms produced by the de Casteljau
-algorithm in exact arithmetic satisfy:
-\begin{equation}
-  \left|b_0^{(0)}\right| \leq \cdots \leq
-  \sum_{j = 0}^k \left|b_j^{(k)}\right| B_{j, k}(s) \leq \cdots \leq
-  \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) = \widetilde{p}(s).
-\end{equation}
-\end{lemma}
+\begin{theorem}[\cite{Ogita2005}]\label{thm:sum-k}
+A summation can be computed (\texttt{SumK}, Algorithm~\ref{alg:sum-k})
+with results that are as accurate as if computed in \(K\) times the
+working precision. When computed this way, the result satisfies:
+\begin{equation}
+\left|\mathtt{SumK}(v, K) - \sum_{j = 1}^n v_j\right| \leq
+\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{j = 1}^n v_j\right| +
+\gamma_{2n - 2}^K \sum_{j = 1}^n \left|v_j\right|.
+\end{equation}
+\end{theorem}
 
@@
 
-\begin{theorem}
-  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
-  \begin{equation}
-    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
-      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
-      \right] +
-      \left[\left(\alpha(n) \mach\right)^K + \mathcal{O}\left(
-      \mach^{K + 1}\right)\right] \cond{p, s}.
-  \end{equation}
-\end{theorem}
+To enable a bound on the \(K\)-order error \(\db{K}^{(0)}_0\), it's necessary
+to understand the difference between the exact local errors \(\ell_{F, j}\)
+and the computed equivalents \(\widehat{\ell}_{F, j}\). To do this, we define
+\begin{equation}
+\widetilde{\ell}_{F, j} \coloneqq \left|e_1\right| +
+ + \cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
+\cdb{F - 1}_j^{(k + 1)}\right|.
+\end{equation}
+
+\begin{lemma}\label{lemma:ell-tilde}
+The local errors \(\widetilde{\ell}_{F, j}\) satisfy:
+\begin{align}
+\widetilde{\ell}_{1, j}^{(k)} &\leq
+  \gamma_3 \left(
+  (1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
+  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right) \\
+\widetilde{\ell}_{F + 1, j}^{(k)} &\leq
+  \gamma_3 \left(
+  (1 - s) \left|\cdb{F}_j^{(k + 1)}\right| +
+  s \left|\cdb{F}_{j + 1}^{(k + 1)}\right|\right) +
+  \gamma_{5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
+  \text{ for } F \geq 1.
+\end{align}
+\end{lemma}
+
+As we'll see soon (Lemma~\ref{lemma:k-order}), putting a bound on
+sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
+be necessary to get an overall bound on the relative error for
+\texttt{CompDeCasteljauK}. So we define
+\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) and the
+related
+\(D_{0, k} \coloneqq \sum_{j = 0}^k \left|\widehat{b}_j^{(k)}\right|
+B_{j, k}(s)\) and
+\(D_{F, k} \coloneqq \sum_{j = 0}^k \left|\cdb{F}_j^{(k)}\right| B_{j, k}(s)\).
+
+\begin{lemma}\label{lemma:L-and-D-bounds}
+For \(s \in \left[0, 1\right]\), the Bernstein-type sums defined above
+satisfy the following bounds:
+\begin{align}
+D_{0, n - k} &\leq \left(1 + \gamma_{3k}\right) \cdot \widetilde{p}(s) \\
+L_{1, n - k} &\leq \gamma_{3k} \cdot \widetilde{p}(s) \\
+D_{F + 1, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
+  \cdots \gamma_{5F} \cdot \left(1 + \gamma_{3k + 5F + 2}\right)
+  \cdot \widetilde{p}(s) \\
+L_{F + 2, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
+  \cdots \gamma_{5F} \cdot \gamma_{3k + 5F + 2} \cdot \widetilde{p}(s).
+\label{L-sum-bound}
+\end{align}
+In particular, this means that
+\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\) and
+\(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\).
+\end{lemma}
+
+See the appendix (Section~\ref{sec:appendix-proof-details}) for details.
+
+\begin{lemma}\label{lemma:k-order}
+The \(K\)-order error \(\db{K}^{(0)}_0\) satisfies
+\begin{equation}
+  \left|\db{K - 1}^{(0)}_0 - \cdb{K - 1}^{(0)}_0\right| =
+  \left|\db{K}^{(0)}_0\right| \leq \alpha(n) \mach^K
+  \widetilde{p}(s).
+\end{equation}
+\end{lemma}
 
 \begin{proof}
-Note that
-\begin{equation}
-p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
-= \cdots
-= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
-+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
-\end{equation}
-and
-\begin{equation}
-\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
-\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}.
-\end{equation}
-It will be crucial for us to bound \(\db{K}_0^{(0)}\).
-In addition, note that
-\begin{align}
-&\widehat{b}_0^{(0)} \oplus \cdb{1}_0^{(0)} \oplus \cdots \oplus
-\cdb{K - 1}_0^{(0)} \\
-&= \widehat{b}_0^{(0)} \left(1 + \theta_{K - 1}\right) +
-\cdb{1}_0^{(0)} \left(1 + \theta_{K - 1}\right) + \cdots
-+ \cdb{K - 1}_0^{(0)} \left(1 + \theta_1\right).
-\end{align}
-We'll use these two quantities to prove a bound.
-
-First, note that \(\widehat{r} = (1 - s)(1 + \theta_1)\). Due to this,
+As in \eqref{matrix-de-casteljau}, we can express the compensated
+de Casteljau algorithm as
+\begin{equation}
+\db{F}^{(k)} = U_{k + 1} \db{F}^{(k + 1)} + \ell_{F}^{(k)}
+\Longrightarrow \db{F}^{(0)} = \sum_{k = 0}^{n - 1}
+U_1 \cdots U_k \ell_F^{(k)} = \sum_{k = 0}^{n - 1}
+\left[\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\right].
+\end{equation}
+For the inexact equivalent of these things, first note that
+\(\widehat{r} = (1 - s)(1 + \delta)\). Due to this,
 we put the \(\widehat{r}\) term at the end of each update step to reduce
@@
 \Longrightarrow \cdb{F}^{(k)} &=
-  U_{k + 1}(s) \cdb{F}^{(k + 1)}(1 + \theta_3) +
-  \widehat{\ell}_{F}^{(k)} (1 + \theta_2).
+  U_{k + 1} \cdb{F}^{(k + 1)}(1 + \theta_3) +
+  \widehat{\ell}_{F}^{(k)} (1 + \theta_2) \\
+\Longrightarrow \cdb{F}^{(0)} &=
+  \sum_{k = 0}^{n - 1}
+  U_1 \cdots U_k \widehat{\ell}_F^{(k)} (1 + \theta_{3k + 2})
+  = \sum_{k = 0}^{n - 1}
+  \left[\sum_{j = 0}^k \widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})
+    B_{j, k}(s)\right].
 \end{align}
-This is quite similar to the relationship satisfied in
-exact arithmetic:
-\begin{equation}
-\db{F}^{(k)} = U_{k + 1}(s) \db{F}^{(k + 1)} + \ell_{F}^{(k)}.
-\end{equation}
-
+Since
+\begin{equation}
+\db{F + 1}_0^{(0)} = \db{F}_0^{(0)} - \cdb{F}_0^{(0)} = \sum_{k = 0}^{n - 1}
+\sum_{j = 0}^k \left(\ell_{F, j}^{(k)} -
+\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right) B_{j, k}(s)
+\end{equation}
+it's useful to put a bound on \(\ell_{F, j}^{(k)} -
+\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\). Via
+\begin{align}
+\widehat{\ell}_{F, j}^{(k)} &= e_1 \oplus \cdots \oplus e_{5F - 2} \oplus
+\left(\rho \otimes \cdb{F - 1}_j^{(k + 1)}\right) \\
+&= e_1\left(1 + \theta_{5F - 2}\right) + \cdots +
+e_{5F - 2}\left(1 + \theta_2\right) +
+\rho \cdot \cdb{F - 1}_j^{(k + 1)} \left(1 + \theta_2\right)
+\end{align}
+we see that
+\begin{equation}
+\left|\ell_{F, j}^{(k)} -
+\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right| \leq
+\gamma_{3k + 5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
+\Longrightarrow
+\left|\db{F + 1}_0^{(0)}\right| \leq \sum_{k = 0}^{n - 1}
+\gamma_{3k + 5F} \sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s).
+\end{equation}
+Since Lemma~\ref{lemma:L-and-D-bounds} (in particular \eqref{L-sum-bound})
+tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
+= L_{F, k} = \mathcal{O}\left(\mach^F\right)\) ...
+\end{proof}
+
+\begin{theorem}
+  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
+  \begin{equation}
+    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
+      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
+      \right] +
+      \left[\alpha(n) \mach^K + \mathcal{O}\left(
+      \mach^{K + 1}\right)\right] \cond{p, s}.
+  \end{equation}
+\end{theorem}
+
+\begin{proof}
+Since
+\begin{equation}
+\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
+\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)},
+\end{equation}
+applying Theorem~\ref{thm:sum-k} tells us that
+\begin{equation}\label{sum-k-applied}
+\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
+\cdb{F}_0^{(0)}\right| \leq
+\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
+\cdb{F}_0^{(0)}\right| +
+\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right|.
+\end{equation}
+Since
+\begin{equation}
+p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
+= \cdots
+= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
++ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
+\end{equation}
+we have
+\begin{gather}
+\left|\sum_{F = 0}^{K - 1} \cdb{F}_0^{(0)}\right|
+\leq \left|p(s)\right| + \left|\db{K}_0^{(0)}\right| \quad \text{and} \\
+\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \leq
+\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
+\cdb{F}_0^{(0)}\right| +
+\left|\db{K}_0^{(0)}\right| \label{triangle-ps}.
+\end{gather}
+Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
+\mathcal{O}\left(\mach^F\right) \widetilde{p}(s)\), hence
+\begin{align}
+\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
+\cdb{F}_0^{(0)}\right| &\leq
+\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
+\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
+\gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
+\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+&\leq
+\gamma_{2n - 2}^K \left[\left|p(s)\right| +
+  \mathcal{O}\left(\mach\right) \widetilde{p}(s)\right] +
+\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s).
+\end{align}
+Combining this with \eqref{sum-k-applied} and \eqref{triangle-ps}, we
+see
+\begin{align}
+\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| &\leq
+\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
+\left|\db{K}_0^{(0)}\right| +
+\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+&\leq
+\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
+\left[\alpha(n) \mach^K + \mathcal{O}\left(\mach^{K + 1}\right) \right]
+\widetilde{p}(s).
+\end{align}
+Dividing this by \(\left|p(s)\right|\), we have our result.
 \end{proof}
@@
 
+\noindent The following algorithms from \cite{Ogita2005} can be used as a
+compensated method for computing a sum of \(n\) numbers. The first is a vector
+transformation that is used as a helper:
+
+\begin{breakablealgorithm}
+  \caption{\textit{Error-free vector transformation for summation.}}
+  \label{alg:vec-sum}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{VecSum}\)}{$p$}
+      \State \(n = \texttt{length}(p)\)
+      \For{\(j = 2, \ldots, n\)}
+        \State \(\left[p_i, p_{i - 1}\right] = \mathtt{TwoSum}\left(
+            p_i, p_{i - 1}\right)\)
+      \EndFor
+    \EndFunction
+  \end{algorithmic}
+\end{breakablealgorithm}
+
+\noindent The second (\texttt{SumK}) computes a sum with results that are as
+accurate as if computed in \(K\) times the working precision. It requires
+\((6K - 5)(n - 1)\) floating point operations.
+
+\begin{breakablealgorithm}
+  \caption{\textit{Summation as in K-fold precision by \((K - 1)\)-fold error-free vector transformation.}}
+  \label{alg:sum-k}
+
+  \begin{algorithmic}
+    \Function{\(\mathtt{res} = \mathtt{SumK}\)}{$p, K$}
+      \For{\(j = 1, \ldots, K - 1\)}
+        \State \(p = \mathtt{VecSum}(p)\)
+      \EndFor
+      \State \(\mathtt{res} = p_1 \oplus p_2 \oplus \cdots \oplus p_n\)
+    \EndFunction
+  \end{algorithmic}
+\end{breakablealgorithm}
+
+\noindent Since the final error \(\cdb{K - 1}\) will not track the errors
+during computation, we have a non-EFT version of
+Algorithm~\ref{alg:local-error-eft}:
+
+\begin{breakablealgorithm}
+  \caption{\textit{Compute the local error (non-EFT).}}
+  \label{alg:local-error}
+
+  \begin{algorithmic}
+    \Function{\(\widehat{\ell} =
+        \mathtt{LocalError}\)}{$e, \rho, \delta b$}
+      \State \(L = \texttt{length}(e)\)
+      \\
+      \State \(\widehat{\ell} = e_1 \oplus e_2\)
+      \For{\(j = 3, \ldots, L\)}
+        \State \(\widehat{\ell} = \widehat{\ell} \oplus e_j\)
+      \EndFor
+      \\
+      \State \(\widehat{\ell} = \widehat{\ell} \oplus \left(\rho \otimes \delta b\right)\)
+    \EndFunction
+  \end{algorithmic}
+\end{breakablealgorithm}
+
+\section{Appendix: Proof Details}\label{sec:appendix-proof-details}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:ell-tilde}]
+Coming soon.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:L-and-D-bounds}]
+First, note that for \textbf{any} sequence \(v_0, \ldots, v_{k + 1}\) we
+must have
+\begin{equation}
+\sum_{j = 0}^k \left[(1 - s) v_j + s v_{j + 1}\right] B_{j, k}(s) =
+\sum_{j = 0}^{k + 1} v_j B_{j, k + 1}(s).
+\end{equation}
+Using this, along with Lemma~\ref{lemma:ell-tilde}, we have
+\begin{align}
+L_{1, k} &\leq \gamma_3 D_{0, k + 1} \label{ell-1-k} \\
+L_{F + 1, k} &\leq \gamma_3 D_{F, k + 1} + \gamma_{5F} L_{F, k}
+\label{ell-F-k}
+\end{align}
+In addition, for \(F \geq 1\) since
+\begin{align}
+\cdb{F}_j^{(k)} &= \widehat{\ell}_{F, j}^{(k)} \oplus \left(
+  s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus \left((1 \ominus s) \otimes
+  \cdb{F}_{j}^{(k + 1)}\right) \\
+&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
+  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
+  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2)
+\end{align}
+we have
+\begin{equation}\label{df-first}
+D_{F, k} \leq (1 + \gamma_3) D_{F, k + 1} + (1 + \gamma_2) \sum_{j = 0}^k
+\left|\widehat{\ell}_{F, j}^{(k)}\right| B_{j, k}(s).
+\end{equation}
+Since \(\ell_{F, j}^{(k)}\) has \(5F - 1\) terms (only the last of which
+involves a product), the terms in the computed value will be involved in
+at most \(5F - 2\) flops, hence
+\(\left|\widehat{\ell}_{F, j}^{(k)}\right| \leq
+\left(1 + \gamma_{5F - 2}\right) \widetilde{\ell}_{F, j}^{(k)}.\)
+Combined with \eqref{df-first} and the fact that there is no local error
+when \(F = 0\), this means
+\begin{align}
+D_{0, k} &\leq (1 + \gamma_3) D_{0, k + 1} \label{d-0-k} \\
+D_{F, k} &\leq (1 + \gamma_3) D_{F, k + 1} + (1 + \gamma_{5F}) L_{F, k}.
+\label{d-F-k}
+\end{align}
+The four inequalities \eqref{ell-1-k}, \eqref{ell-F-k}, \eqref{d-0-k}
+and \eqref{d-F-k} allow us to write all bounds in terms of
+\(D_{0, n} = \widetilde{p}(s)\) and \(D_{F, n} = 0\). From \eqref{d-0-k}
+we can conclude that \(D_{0, n - k} \leq \left(1 + \gamma_{3k}\right) \cdot
+\widetilde{p}(s)\) and from \eqref{ell-1-k} that \(L_{1, n - k} \leq
+\gamma_{3k} \cdot \widetilde{p}(s)\).
+
+To see that these inequalities make sense qualitatively, note the
+following. It's clear from \eqref{ell-F-k} that if
+\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot
+\widetilde{p}(s)\) and \(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot
+\widetilde{p}(s)\) then \(L_{F + 1, k} = \mathcal{O}\left(
+\mach^{F + 1}\right) \cdot \widetilde{p}(s)\) and \eqref{d-F-k} shows
+that \(D_{F, k} = \mathcal{O}\left(L_{F, k}\right)\). Actually verifying
+the quantitative bounds is quite tedious and does not provide any further
+insight into the behavior of these errors. For example, the bounds above
+are loose bounds and much tighter bounds exist that don't have a compact
+formula.
+\end{proof}
+
 \end{document}

### 2018-05-07 nox.py

@@ -113,6 +113,16 @@ def flop_counts(session):
 
 
 @nox.session
+def verify_table(session):
+    # No need to create a virtualenv.
+    session.virtualenv = False
+
+    env = {"PYTHONPATH": get_path("src")}
+    script = get_path("scripts", "verify_table.py")
+    session.run("python", script, env=env)
+
+
+@nox.session
 def make_images(session):
     session.interpreter = SINGLE_INTERP
     # Install all dependencies.

### 2018-05-07 scripts/compute_counts.py

@@ -236,10 +236,10 @@ def de_casteljau_expected_total(K, n):
 
     .. math::
 
-       (15K^2 - 34K + 26)T_n + K + 5.
+       (15K^2 - 34K + 26)T_n + 6K^2 - 11K + 11.
     """
     Tn = (n * (n + 1)) // 2
-    return (15 * K ** 2 - 34 * K + 26) * Tn + K + 5
+    return (15 * K ** 2 - 34 * K + 26) * Tn + 6 * K ** 2 - 11 * K + 11
 
 
 def de_casteljau_expected_fma(K, n):
@@ -258,7 +258,7 @@ def de_casteljau_expected_fma(K, n):
 
 
 def count_de_casteljau_compensated():
-    print("de_casteljau.compensated() (9n^2 + 9n + 7 = 18 T_n + 7):")
+    print("de_casteljau.compensated() (9n^2 + 9n + 7 = 18 T_n + 13):")
     for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -274,7 +274,9 @@ def count_de_casteljau_compensated():
 
 
 def count_de_casteljau_compensated3():
-    print("de_casteljau.compensated3() ((59n^2 + 59n + 16) / 2 = 59 T_n + 8):")
+    print(
+        "de_casteljau.compensated3() ((59n^2 + 59n + 16) / 2 = 59 T_n + 32):"
+    )
     for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -290,7 +292,7 @@ def count_de_casteljau_compensated3():
 
 
 def count_de_casteljau_compensated4():
-    print("de_casteljau.compensated4() (65n^2 + 65n + 9 = 130 T_n + 9):")
+    print("de_casteljau.compensated4() (65n^2 + 65n + 9 = 130 T_n + 63):")
     for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -308,7 +310,7 @@ def count_de_casteljau_compensated4():
 def count_de_casteljau_compensated5():
     msg = (
         "de_casteljau.compensated5() ((231n^2 + 231n + 20) / 2 "
-        "= 231 T_n + 10):"
+        "= 231 T_n + 106):"
     )
     print(msg)
     for degree in range(1, 5 + 1):

### 2018-05-07 scripts/verify_table.py

@@ -1,0 +1,198 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Verify table in manuscript.
+
+This performs compensated de Casteljau (``K = 2``) in both IEEE-754
+double precision and with exact arithmetic and verifies that the terms
+in the table are correct.
+"""
+
+import fractions
+
+import eft
+
+
+F = fractions.Fraction
+K = 1001
+ROOT = 0.5
+U = F(1, 2 ** 53)
+MACH = 0.5 ** 53
+S = ROOT + K * MACH
+R_HAT, RHO = eft.add_eft(1.0, -S)
+
+
+def exact_de_castlejau(b4_vals):
+    b = {}
+    b[4] = tuple(map(F, b4_vals))
+    s = F(S)
+
+    for k in (3, 2, 1, 0):
+        prev_b = b[k + 1]
+        b[k] = tuple(
+            (1 - s) * b1 + s * b2 for b1, b2 in zip(prev_b, prev_b[1:])
+        )
+
+    return b
+
+
+def stage1(b4_vals, exact_b):
+    b04, b14, b24, b34, b44 = b4_vals
+
+    P1, pi1 = eft.multiply_eft(R_HAT, b04)
+    P2, pi2 = eft.multiply_eft(S, b14)
+    b03, sigma3 = eft.add_eft(P1, P2)
+    e03 = pi1 + pi2 + sigma3 + RHO * b04
+
+    P1, pi1 = eft.multiply_eft(R_HAT, b14)
+    P2, pi2 = eft.multiply_eft(S, b24)
+    b13, sigma3 = eft.add_eft(P1, P2)
[…]

### 2018-05-07 src/de_casteljau.py

@@ -171,19 +171,19 @@ def _compensated_k(s, coeffs, K):
 
 def compensated(s, coeffs):
     b, db = _compensated_k(s, coeffs, 2)
-    return b + db
+    return eft.sum_k((b, db), 2)
 
 
 def compensated3(s, coeffs):
     b, db, d2b = _compensated_k(s, coeffs, 3)
-    return (b + db) + d2b
+    return eft.sum_k((b, db, d2b), 3)
 
 
 def compensated4(s, coeffs):
     b, db, d2b, d3b = _compensated_k(s, coeffs, 4)
-    return ((b + db) + d2b) + d3b
+    return eft.sum_k((b, db, d2b, d3b), 4)
 
 
 def compensated5(s, coeffs):
     b, db, d2b, d3b, d4b = _compensated_k(s, coeffs, 5)
-    return (((b + db) + d2b) + d3b) + d4b
+    return eft.sum_k((b, db, d2b, d3b, d4b), 5)

### 2018-05-07 doc/paper.tex
Updating the L{F, k} lemma to actually be something useful.
This was simply the result of "sleeping" on it.

@@
 \newcommand{\fl}[1]{\operatorname{fl}\left(#1\right)}
-\newcommand{\eps}{\varepsilon}
+\newcommand{\bigO}[1]{\mathcal{O}\left(#1\right)}
 \newcommand{\mach}{\mathbf{u}}
@@
   \frac{\left|p(s) - \mathtt{DeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq
-  \cond{p, s} \times \mathcal{O}(\mach).
+  \cond{p, s} \times \bigO{\mach}.
 \end{equation}
@@
     \left|p(s)\right|} \leq \mach + \cond{p, s} \times
-    \mathcal{O}\left(\mach^2\right).
+    \bigO{\mach^2}.
 \end{equation}
@@
     \left|p(s)\right|} \leq \mach + \cond{p, s} \times
-    \mathcal{O}\left(\mach^K\right).
+    \bigO{\mach^K}.
 \end{equation}
@@
 \leq \frac{3}{2} \cdot 10^{-11}\) --- evaluates to exactly \(0\) when
-it should be \(\mathcal{O}\left(\mach^3\right)\). As shown in
+it should be \(\bigO{\mach^3}\). As shown in
 Table~\ref{tab:exact-computation}, the breakdown occurs because
@@
 (15K^2 + 11K - 34)T_n + 6K^2 - 11K + 11 =
-\mathcal{O}\left(n^2 K^2\right)
+\bigO{n^2 K^2}
 \end{equation}
@@
   (1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
-  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right) \\
+  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right)
+  \label{ell-tilde-1} \\
 \widetilde{\ell}_{F + 1, j}^{(k)} &\leq
@@
 be necessary to get an overall bound on the relative error for
-\texttt{CompDeCasteljauK}. So we define
-\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) and the
-related
-\(D_{0, k} \coloneqq \sum_{j = 0}^k \left|\widehat{b}_j^{(k)}\right|
-B_{j, k}(s)\) and
-\(D_{F, k} \coloneqq \sum_{j = 0}^k \left|\cdb{F}_j^{(k)}\right| B_{j, k}(s)\).
+\texttt{CompDeCasteljauK}, so we define
+\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\).
 
 \begin{lemma}\label{lemma:L-and-D-bounds}
-For \(s \in \left[0, 1\right]\), the Bernstein-type sums defined above
-satisfy the following bounds:
-\begin{align}
-D_{0, n - k} &\leq \left(1 + \gamma_{3k}\right) \cdot \widetilde{p}(s) \\
-L_{1, n - k} &\leq \gamma_{3k} \cdot \widetilde{p}(s) \\
-D_{F + 1, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
-  \cdots \gamma_{5F} \cdot \left(1 + \gamma_{3k + 5F + 2}\right)
-  \cdot \widetilde{p}(s) \\
-L_{F + 2, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
-  \cdots \gamma_{5F} \cdot \gamma_{3k + 5F + 2} \cdot \widetilde{p}(s).
-\label{L-sum-bound}
-\end{align}
+For \(s \in \left[0, 1\right]\), the Bernstein-type sum defined above
+satisfies the following bound:
+\begin{equation}\label{L-sum-bound}
+L_{F, n - k} \leq \left[\left(3^F \binom{k}{F - 1} + \bigO{k^{F - 1}}\right)
+  \mach^F + \bigO{\mach^{F + 1}}\right] \cdot \widetilde{p}(s).
+\end{equation}
 In particular, this means that
-\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\) and
-\(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\).
+\(L_{F, k} = \bigO{(3 \mach)^F} \cdot \widetilde{p}(s)\).
 \end{lemma}
 
-See the appendix (Section~\ref{sec:appendix-proof-details}) for details.
+See the appendix (Section~\ref{sec:appendix-proof-details}) for details on
+proving Lemma~\ref{lemma:ell-tilde} and Lemma~\ref{lemma:L-and-D-bounds}.
 
@@
 tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
-= L_{F, k} = \mathcal{O}\left(\mach^F\right)\) ...
+= L_{F, k} = \bigO{\mach^F}\) ...
 \end{proof}
@@
     \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
-      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
+      \left|p(s)\right|} \leq \left[\mach + \bigO{\mach^2}
       \right] +
-      \left[\alpha(n) \mach^K + \mathcal{O}\left(
-      \mach^{K + 1}\right)\right] \cond{p, s}.
+      \left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}}\right] \cond{p, s}.
   \end{equation}
@@
 Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
-\mathcal{O}\left(\mach^F\right) \widetilde{p}(s)\), hence
+\bigO{\mach^F} \widetilde{p}(s)\), hence
 \begin{align}
@@
 \cdb{F}_0^{(0)}\right| &\leq
-\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
-\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
+\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
 \gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
 \gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
-\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
 &\leq
 \gamma_{2n - 2}^K \left[\left|p(s)\right| +
-  \mathcal{O}\left(\mach\right) \widetilde{p}(s)\right] +
-\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s).
+  \bigO{\mach} \widetilde{p}(s)\right] +
+\bigO{\mach^{K + 1}} \widetilde{p}(s).
 \end{align}
@@
 \left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| &\leq
-\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
+\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
 \left|\db{K}_0^{(0)}\right| +
-\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
+\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
 &\leq
-\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
-\left[\alpha(n) \mach^K + \mathcal{O}\left(\mach^{K + 1}\right) \right]
+\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
+\left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}} \right]
 \widetilde{p}(s).
@@
 in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
-error that is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+error that is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
 \(1 / \mach\), at which point the relative error increases linearly with
-the condition number until it becomes \(\mathcal{O}(1)\) when
+the condition number until it becomes \(\bigO{1}\) when
 \(\cond{p, s}\) reaches \(1 / \mach^2\).
@@
 (Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
-is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
 \(1 / \mach^2\) at which point the relative error increases linearly
-to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
+to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
 and the relative error in \texttt{CompDeCasteljau4}
 (Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
-is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
+is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
 \(1 / \mach^3\) at which point the relative error increases linearly
-to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).
+to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).
 
@@
 \end{equation}
-Using this, along with Lemma~\ref{lemma:ell-tilde}, we have
+For example of this in use, via \eqref{ell-tilde-1}, we have
+\begin{equation}
+  L_{1, k} \leq \gamma_3 \sum_{j = 0}^{k + 1} \left|
+  \widehat{b}_j^{(k + 1)}\right| B_{j, k + 1}(s).
+\end{equation}
+In order to work with sums of this form, we define Bernstein-type
+sums related to \(L_{F, k}\):
+\begin{align}
+D_{0, k} &\coloneqq \sum_{j = 0}^k \left|\widehat{b}_j^{(k)}\right|
+B_{j, k}(s) \\
+D_{F, k} &\coloneqq \sum_{j = 0}^k \left|\cdb{F}_j^{(k)}\right| B_{j, k}(s).
+\end{align}
+Hence Lemma~\ref{lemma:ell-tilde} gives
 \begin{align}
@@
 \widetilde{p}(s)\) and from \eqref{ell-1-k} that \(L_{1, n - k} \leq
-\gamma_{3k} \cdot \widetilde{p}(s)\).
-
-To see that these inequalities make sense qualitatively, note the
-following. It's clear from \eqref{ell-F-k} that if
-\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot
-\widetilde{p}(s)\) and \(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot
-\widetilde{p}(s)\) then \(L_{F + 1, k} = \mathcal{O}\left(
-\mach^{F + 1}\right) \cdot \widetilde{p}(s)\) and \eqref{d-F-k} shows
-that \(D_{F, k} = \mathcal{O}\left(L_{F, k}\right)\). Actually verifying
-the quantitative bounds is quite tedious and does not provide any further
-insight into the behavior of these errors. For example, the bounds above
-are loose bounds and much tighter bounds exist that don't have a compact
-formula.
+\gamma_3 \left(1 + \gamma_{3(k - 1)}\right) \cdot \widetilde{p}(s)\).
+
+To show the bounds for higher values of \(F\), we'll assume we have
+bounds of the form
+\(D_{F, n - k} \leq \left(q_F(k) \mach^F + \bigO{\mach^{F + 1}}\right) \cdot
+\widetilde{p}(s)\) and
+\(L_{F, n - k} \leq \left(r_F(k) \mach^F + \bigO{\mach^{F + 1}}\right) \cdot
+\widetilde{p}(s)\) for two families of polynomials \(q_F(k), r_F(k)\). We
+have \(q_0(k) = 1\) and \(r_1(k) = 3\) as our base cases and can build from
+there. To satisfy \eqref{d-F-k}, we'd like
+\(q_F(k) = q_F(k - 1) + r_F(k)\)
+and for \eqref{ell-F-k}
+\(r_{F + 1}(k) = 3 q_F(k - 1) + 5 F r_F(k)\).
+Since the forward difference \(\Delta q_F(k) = r_F(k + 1)\) is known,
+we can inductively solve for \(q_F\) in terms of \(q_F(0)\). But
+\(D_{F, n} = 0\) gives \(q_F(0) = 0\).
+
+For example, since we have \(r_1(k) = 3 \binom{k}{0}\) we'll have
+\(q_1(k) = 3 \binom{k}{1}\). Once this is known
+\begin{equation}
+r_2(k) = 3 q_1(k - 1) + 5 r_1(k) = 3 \cdot 3 \binom{k - 1}{1} +
+5 \cdot 3 \binom{k}{0} = 9 \binom{k}{1} + 6 \binom{k}{0}.
+\end{equation}
+If we write these polynomials in the ``falling factorial'' basis of
+forward differences, then we can show that
+\begin{equation}
+r_F(k) = 3^F \binom{k}{F} + \cdots
+\end{equation}
+which will complete the proof. To see this, first note that for a polynomial
+in this basis
+\(f(k) = A \binom{k}{d} + B \binom{k}{d - 1} + C \binom{k}{d - 2} +
+D \binom{k}{d - 3} + \cdots\) we have
+\begin{align}
+f(k + 1) &= A \binom{k}{d} + (A + B) \binom{k}{d - 1} +
+  (B + C) \binom{k}{d - 2} + (C + D) \binom{k}{d - 3} + \cdots \\
+f(k - 1) &= A \binom{k}{d} + (B - A) \binom{k}{d - 1} +
+  (C - B + A) \binom{k}{d - 2} + (D - C + B - A) \binom{k}{d - 3} + \cdots
+\end{align}
+Using these, we can show that if
+\(r_F(k) = \sum_{j = 0}^{F - 1} c_j \binom{k}{j}\) then
+\begin{align}
+q_F(k) &= c_{F - 1} \binom{k}{F} + \sum_{j = 1}^{F - 1}
+  (c_j + c_{j - 1}) \binom{k}{j} \\
+r_{F + 1}(k) &= 3 \left[-c_0 \binom{k}{0} +
+  \sum_{j = 1}^F c_{j - 1} \binom{k}{j}\right] +
+(5F) \sum_{j = 0}^{F - 1} c_j \binom{k}{j} =
+3 c_{F - 1} \binom{k}{F} + \cdots
+\end{align}
+Under the inductive hypothesis \(c_{F - 1} = 3^F\) so that
+the lead term in \(r_{F + 1}(k)\) is \(3 c_{F - 1} \binom{k}{F}
+= 3^{F + 1} \binom{k}{F}\).
 \end{proof}

### 2018-05-07 doc/paper.tex
Finished the a prior bounds proof(s).
Still needed:

- Proof of Lemma 4.2
- Add comment about what the a priori bounds are for small
  values of K
- Use the a priori bounds in my plot (and maybe in the JLCS10
  plot)

@@
 sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
-be necessary to get an overall bound on the relative error for
+be useful to get an overall bound on the relative error for
 \texttt{CompDeCasteljauK}, so we define
@@
 For \(s \in \left[0, 1\right]\), the Bernstein-type sum defined above
-satisfies the following bound:
-\begin{equation}\label{L-sum-bound}
-L_{F, n - k} \leq \left[\left(3^F \binom{k}{F - 1} + \bigO{k^{F - 1}}\right)
-  \mach^F + \bigO{\mach^{F + 1}}\right] \cdot \widetilde{p}(s).
-\end{equation}
+satisfies the following bounds:
+\begin{align}
+L_{F, n - k} &\leq \left[\left(3^F \binom{k}{F - 1} + \bigO{k^{F - 1}}\right)
+  \mach^F + \bigO{\mach^{F + 1}}\right] \cdot \widetilde{p}(s) \\
+\sum_{k = 0}^{n - 1} \gamma_{3k + 5F} L_{F, k} &\leq
+  \left[\left(3^{F + 1} \binom{n}{F + 1} + \bigO{n^F}\right)
+  \mach^{F + 1} + \bigO{\mach^{F + 2}}\right] \cdot \widetilde{p}(s).
+  \label{L-sum-bound}
+\end{align}
 In particular, this means that
-\(L_{F, k} = \bigO{(3 \mach)^F} \cdot \widetilde{p}(s)\).
+\(\sum_{k = 0}^{n - 1} \gamma_{3k + 5F} L_{F, k} =
+\bigO{(3 n \mach)^{F + 1}} \cdot \widetilde{p}(s)\).
 \end{lemma}
@@
   \left|\db{K - 1}^{(0)}_0 - \cdb{K - 1}^{(0)}_0\right| =
-  \left|\db{K}^{(0)}_0\right| \leq \alpha(n) \mach^K
-  \widetilde{p}(s).
+  \left|\db{K}^{(0)}_0\right| \leq
+  \left[\left(3^{K} \binom{n}{K} + \bigO{n^{K - 1}}\right)
+  \mach^{K} + \bigO{\mach^{K + 1}}\right] \cdot \widetilde{p}(s).
 \end{equation}
@@
 \end{equation}
-Since Lemma~\ref{lemma:L-and-D-bounds} (in particular \eqref{L-sum-bound})
-tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
-= L_{F, k} = \bigO{\mach^F}\) ...
+Applying \eqref{L-sum-bound} directly gives
+\begin{equation}
+\left|\db{F + 1}_0^{(0)}\right| \leq
+  \left[\left(3^{F + 1} \binom{n}{F + 1} + \bigO{n^F}\right)
+  \mach^{F + 1} + \bigO{\mach^{F + 2}}\right] \cdot \widetilde{p}(s).
+\end{equation}
+Letting \(K = F + 1\) we have our result.
 \end{proof}
@@
   If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
-  \begin{equation}
+  \begin{multline}
     \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
       \left|p(s)\right|} \leq \left[\mach + \bigO{\mach^2}
-      \right] +
-      \left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}}\right] \cond{p, s}.
-  \end{equation}
+      \right] + \\
+      \left[\left(3^{K} \binom{n}{K} + \bigO{n^{K - 1}}\right) \mach^K +
+      \bigO{\mach^{K + 1}}\right] \cond{p, s}.
+  \end{multline}
 \end{theorem}
@@
 \begin{align}
-\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| &\leq
+& \left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \\
+\leq &
 \left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
@@
 \bigO{\mach^{K + 1}} \widetilde{p}(s) \\
-&\leq
+\leq &
 \left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
-\left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}} \right]
+\left[\left(3^{K} \binom{n}{K} + \bigO{n^{K - 1}}\right) \mach^K +
+\bigO{\mach^{K + 1}} \right]
 \widetilde{p}(s).
@@
 = 3^{F + 1} \binom{k}{F}\).
+
+For the second inequality, we'll show that
+\begin{equation}
+\sum_{k = 0}^{n - 1} \gamma_{3k + 5F} L_{F, k} \leq
+  \left[q_{F + 1}(n) \mach^{F + 1} +
+  \bigO{\mach^{F + 2}}\right] \cdot \widetilde{p}(s)
+\end{equation}
+and then we'll have our result since we showed above that
+\(q_{F + 1}(n) = 3^{F + 1} \binom{n}{F + 1} + \bigO{n^F}\). Since
+\(\gamma_{3k + 5F} L_{F, k} \leq (3k + 5F) L_{F, k} \mach +
+\bigO{\mach^{F + 1}} \widetilde{p}(s)\) it's enough to consider
+\begin{equation}
+\sum_{k = 0}^{n - 1} (3k + 5F) r_F(n - k) =
+\sum_{k = 1}^n (3(n - k) + 5F) r_F(k).
+\end{equation}
+Since \(q_F(k) = q_F(k - 1) + r_F(k)\) and \(q_F(0) = 0\) we have
+\(q_{F}(n) = \sum_{k = 1}^n r_{F}(k)\) thus
+\begin{equation}
+q_{F + 1}(n) = \sum_{k = 1}^n r_{F + 1}(k)
+= \sum_{k = 1}^n 3 q_F(k - 1) + 5 F r_F(k)
+= \sum_{k = 1}^n 3 \left[\sum_{j = 1}^{k - 1} r_F(j)\right] + 5 F r_F(k).
+\end{equation}
+Swapping the order of summation and grouping like terms, we have our
+result.
 \end{proof}

### 2018-05-07 doc/paper.tex
Updating plots with a priori bounds.
Also

- adding example coefficient functions for bounds
  when K = 1, 2, 3, 4
- adding a sketch of a proof for Lemma 4.2.

@@
 
+For the first few values of \(K\) the coefficient of
+\(\cond{p, s}\) in the bound is
+\begin{center}
+  \begin{tabular}{>{$}c<{$} c >{$}c<{$}}
+    K & Method & \text{Multiplier} \\
+    \midrule
+    1 & \texttt{DeCasteljau} & 3 \binom{n}{1} \mach =
+      3n \mach \approx \gamma_{3n} \\[0.125cm]
+    2 & \texttt{CompDeCasteljau} & \left[9 \binom{n}{2} +
+      15 \binom{n}{1}\right]\mach^2 = \frac{3n(3n + 7)}{2} \mach^2
+      \approx \frac{1}{4} \cdot 2 \gamma_{3n}^2 \\[0.125cm]
+    3 & \texttt{CompDeCasteljau3} & \left[27 \binom{n}{3} +
+      135 \binom{n}{2} + 150 \binom{n}{1}\right] \mach^3 =
+      \frac{3n(3n^2 + 36n + 61)}{2} \mach^3 \\[0.125cm]
+    4 & \texttt{CompDeCasteljau4} & \left[81 \binom{n}{4} + 810 \binom{n}{3} +
+      2475 \binom{n}{2} + 2250 \binom{n}{1}\right] \mach^4 \\[0.125cm]
+  \end{tabular}
+\end{center}
+See the \hyperref[proof:L-and-D-bounds]{proof} of
+Lemma~\ref{lemma:L-and-D-bounds} for more details on where these
+polynomials come from.
+
 \section{Numerical experiments}\label{sec:numerical}
@@
       \\
-      \State \(\widehat{\ell} = \widehat{\ell} \oplus \left(\rho \otimes \delta b\right)\)
+      \State \(\widehat{\ell} = \widehat{\ell} \oplus \left(
+          \rho \otimes \delta b\right)\)
     \EndFunction
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:ell-tilde}]
-Coming soon.
+We'll start with the \(F = 1\) case. Recall where the terms originate:
+\begin{align}
+\left[P_1, e_1\right] &= \mathtt{TwoProd}\left(\widehat{r},
+  \widehat{b}_j^{(k + 1)}\right) \\
+\left[P_2, e_2\right] &= \mathtt{TwoProd}\left(s,
+  \widehat{b}_{j + 1}^{(k + 1)}\right) \\
+\left[\widehat{b}_j^{(k)}, e_3\right] &= \mathtt{TwoSum}\left(P_1, P_2\right).
+\end{align}
+Hence Theorem~\ref{thm:eft} tells us that
+\begin{align}
+\left|P_1\right| &\leq (1 + \mach)\left|\widehat{r} \cdot
+  \widehat{b}_j^{(k + 1)}\right| \leq (1 + \mach)^2 (1 - s)
+  \left|\widehat{b}_j^{(k + 1)}\right| \\
+\left|e_1\right| &\leq \mach \left|\widehat{r} \cdot
+  \widehat{b}_j^{(k + 1)}\right| \leq \mach(1 + \mach)(1 - s) \left|
+  \widehat{b}_j^{(k + 1)}\right| \\
+\left|P_2\right| &\leq (1 + \mach) s \left|\widehat{b}_{j + 1}^{(k + 1)}\right| \\
+\left|e_2\right| &\leq \mach s \left|\widehat{b}_{j + 1}^{(k + 1)}\right| \\
+\left|e_3\right| &\leq \mach \left|P_1\right| + \mach\left|P_2\right| \\
+\left|\rho \cdot \widehat{b}_j^{(k + 1)}\right| &\leq
+(1 + \mach)(1 - s) \left|\widehat{b}_j^{(k + 1)}\right|.
+\end{align}
+In general, we can swap \(\mach\left|P_j\right|\) for
+\((1 + \mach)\left|e_j\right|\) based on how closely related the bound
+on the result and the bound on the error are. Thus
+\begin{align}
+\widetilde{\ell}_{1, j}^{(k)} &= \left|e_1\right| + \left|e_2\right| +
+  \left|e_3\right| + \left|\rho \cdot \widehat{b}_j^{(k + 1)}\right| \\
+&\leq (2 + \mach)\left(\left|e_1\right| + \left|e_2\right|\right) +
+  (1 + \mach)(1 - s) \left|\widehat{b}_j^{(k + 1)}\right| \\
+&\leq \left[(1 + \mach)^3 - 1\right] (1 - s) \left|
+  \widehat{b}_j^{(k + 1)}\right| + \left[(1 + \mach)^2 - 1\right] s \left|
+  \widehat{b}_{j + 1}^{(k + 1)}\right| \\
+&\leq \gamma_3 \left((1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
+  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right).
+\end{align}
+For \(\widetilde{\ell}_{F + 1}\), we want to relate the ``current'' errors
+\(e_1, \ldots, e_{5F + 3}\) to the ``previous'' errors \(e_1',
+\ldots, e_{5F - 2}'\). In the same fashion as above, we track where
+the current errors come from:
+\begin{align}
+\left[S_1, e_1\right] &= \mathtt{TwoSum}\left(e_1', e_2'\right) \\
+\left[S_2, e_2\right] &= \mathtt{TwoSum}\left(S_1, e_3'\right) \\
+&\mathrel{\makebox[\widthof{=}]{\vdots}} \nonumber \\
+\left[S_{5F - 3}, e_{5F - 3}\right] &=
+  \mathtt{TwoSum}\left(S_{5F - 4}, e_{5F - 2}'\right) \\
+\left[P_{5F - 2}, e_{5F - 2}\right] &= \mathtt{TwoProd}\left(\rho,
+  \cdb{F - 1}_j^{(k + 1)}\right) \\
+\left[\widehat{\ell}_{F, j}^{(k)}, e_{5F - 1}\right] &=
+  \mathtt{TwoSum}\left(S_{5F - 3}, P_{5F - 2}\right) \\
+\left[P_{5F}, e_{5F}\right] &= \mathtt{TwoProd}\left(s,
+  \cdb{F}_{j + 1}^{(k + 1)}\right) \\
+\left[S_{5F + 1}, e_{5F + 1}\right] &=
+  \mathtt{TwoSum}\left(\widehat{\ell}_{F, j}^{(k)}, P_{5F}\right) \\
+\left[P_{5F + 2}, e_{5F + 2}\right] &= \mathtt{TwoProd}\left(\rho,
+  \cdb{F}_j^{(k + 1)}\right) \\
+\left[\cdb{F}_j^{(k)}, e_{5F + 3}\right] &= \mathtt{TwoSum}\left(
+  S_{5F + 1}, P_{5F + 2}\right).
+\end{align}
+Arguing as we did above, we start with
+\(\left|e_1\right| \leq \mach \left|e_1'\right| + \mach \left|e_2'\right|\)
+and build each bound recursively based on the previous, e.g.
+\(\left|e_2\right| \leq \mach \left|S_1\right| + \mach \left|e_3'\right| \leq
+(1 + \mach) \mach \left|e_1'\right| + (1 + \mach) \mach \left|e_2'\right| +
+\mach \left|e_3'\right|\). Proceeding in this fashion, we find
+\begin{align}
+\widetilde{\ell}_{F + 1, j}^{(k)} &= \left|e_1\right| + \cdots +
+  \left|e_{5F + 3}\right| + \left|\rho \cdot \cdb{F}_j^{(k + 1)}\right| \\
+&\leq \gamma_{5F} \left|e_1'\right| + \gamma_{5F} \left|e_2'\right| +
+  \gamma_{5F - 1} \left|e_3'\right| + \cdots +
+  \gamma_4 \left|e_{5F - 2}'\right| +
+  \gamma_4 \left|\rho \cdot \cdb{F - 1}_j^{(k + 1)}\right| \\
+&\qquad + \gamma_3 (1 - s) \left|
+  \cdb{F}_j^{(k + 1)}\right| + \gamma_3 s \left|
+  \cdb{F}_{j + 1}^{(k + 1)}\right| \\
+&\leq \gamma_3 \left(
+  (1 - s) \left|\cdb{F}_j^{(k + 1)}\right| +
+  s \left|\cdb{F}_{j + 1}^{(k + 1)}\right|\right) +
+  \gamma_{5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
+\end{align}
+as desired.
 \end{proof}
 
-\begin{proof}[Proof of Lemma~\ref{lemma:L-and-D-bounds}]
+\begin{proof}[Proof of Lemma~\ref{lemma:L-and-D-bounds}]\label{proof:L-and-D-bounds}
 First, note that for \textbf{any} sequence \(v_0, \ldots, v_{k + 1}\) we

### 2018-05-07 nox.py

@@ -37,7 +37,7 @@ MATPLOTLIB_DEPS = (
 DEPS = {
     "matplotlib": MATPLOTLIB_DEPS,
     "seaborn": MATPLOTLIB_DEPS
-    + ("pandas == 0.22.0", "scipy == 1.0.1", "seaborn == 0.8.1"),
+    + ("pandas == 0.22.0", "scipy == 1.1.0", "seaborn == 0.8.1"),
 }
 
 

### 2018-05-07 scripts/error_against_cond.py

@@ -46,11 +46,17 @@ POWER_VAL = 1.3
 
 
 def main(filename=None):
-    # n = 8
-    gamma2n = 16 * U / (1 - 16 * U)
+    n = 8
+    gamma2n = (2 * n * U) / (1 - 2 * n * U)
     bound_coeff1 = float(gamma2n)
-    gamma3n = 24 * U / (1 - 24 * U)
-    bound_coeff2 = float(2 * gamma3n ** 2)
+    bound_coeff2 = 3 * n * (3 * n + 7) * U ** 2 / 2
+    bound_coeff2 = float(bound_coeff2)
+    bound_coeff3 = 3 * n * (3 * n ** 2 + 36 * n + 61) * U ** 3 / 2
+    bound_coeff3 = float(bound_coeff3)
+    bound_coeff4 = 9 * n * (
+        3 * n ** 3 + 102 * n ** 2 + 773 * n + 1122
+    ) * U ** 4 / 8
+    bound_coeff4 = float(bound_coeff4)
 
     cond_nums = []
     forward_errs1 = []
@@ -151,6 +157,20 @@ def main(filename=None):
         alpha=alpha,
         zorder=1,
     )
+    ax.loglog(
+        [min_x, max_x],
+        [bound_coeff3 * min_x, bound_coeff3 * max_x],
+        color="black",
+        alpha=alpha,
+        zorder=1,
+    )
+    ax.loglog(
+        [min_x, max_x],
+        [bound_coeff4 * min_x, bound_coeff4 * max_x],
+        color="black",
+        alpha=alpha,
+        zorder=1,
+    )
     # Add the ``x = 1/u^k`` vertical lines.
     delta_y = max_y - min_y
     for exponent in (1, 2, 3, 4):
@@ -240,11 +260,11 @@ def main_jlcs10(filename=None):
     ``-64``. In particular, it shows that the compensated de Casteljau
     algorithm produces exactly zero.
     """
-    # n = 8
-    gamma2n = 16 * U / (1 - 16 * U)
+    n = 8
+    gamma2n = (2 * n * U) / (1 - 2 * n * U)
     bound_coeff1 = float(gamma2n)
-    gamma3n = 24 * U / (1 - 24 * U)
-    bound_coeff2 = float(2 * gamma3n ** 2)
+    bound_coeff2 = 3 * n * (3 * n + 7) * U ** 2 / 2
+    bound_coeff2 = float(bound_coeff2)
 
     cond_nums = []
     forward_errs1 = []

### 2018-05-08 doc/paper.tex
Did a "final" pass and edit.

@@
 \begin{equation}
-  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s)\right|}{
+  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s, K)\right|}{
     \left|p(s)\right|} \leq \mach + \cond{p, s} \times
@@
 
-Throughout this paper we assume that the computation in floating point
-arithmetic obeys the model
+We assume all floating point operations obey
 \begin{equation}
@@
 \left|\delta_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
-and \(\star\) represents the floating point computation, e.g.
+and \(\star\) is a floating point operation, e.g.
 \(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
@@
 \(\alpha \in \mathbf{R}\) in floating point arithmetic by
-\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
+\(\widehat{\alpha}\) or \(\fl{\alpha}\) and use \(\mathbf{F}\) as the set of
 all floating point numbers (see \cite{Higham2002} for more details).
@@
 
-Now let us introduce some results concerning error-free transformation (EFT).
-For a pair floating point numbers \(a, b \in \mathbf{F}\), there exists
-\(y \in \mathbf{F}\) satisfying \(a \circ b = x + y\) where
-\(x = \fl{a \circ b}\) and \(\circ \in \left\{+, -, \times\right\}\).
-The transformation \((a, b) \longrightarrow (x, y)\) is regarded as an
-error-free transformation. The error-free transformation algorithms of the
-sum and product of two floating point numbers used later in this paper are
+An error-free transformation is a computational method where both
+the computed result and the round-off error are returned. It
+is considered ``free'' of error if the round-off can be represented
+exactly as an element of \(\mathbf{F}\).
+The error-free transformations used in this paper are
 the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
-algorithm by Dekker \cite{Dekker1971}, respectively. The following theorem
-exhibits the important properties of \texttt{TwoSum} and \texttt{TwoProd}
-algorithms (see \cite{Ogita2005}).
+algorithm by Dekker \cite{Dekker1971}, respectively.
 
@@
 \texttt{TwoSum} and \texttt{TwoProd} satisfy
-\begin{itemize}
-  \item \(\left[S, \sigma\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
-      \(S + \sigma = a + b\), \(\sigma \leq \mach \left|S\right|\),
-      \(\sigma \leq \mach \left|a + b\right|\)
-  \item \(\left[P, \pi\right] = \mathtt{TwoProd}(a, b)\),
-      \(P = \fl{a \times b}\), \(P + \pi = a \times b\),
-      \(\pi \leq \mach \left|P\right|\),
-      \(\pi \leq \mach \left|a \times b\right|\).
-\end{itemize}
+\begin{alignat}{4}
+\left[S, \sigma\right] &= \mathtt{TwoSum}(a, b), & \, S &= \fl{a + b},
+  S + \sigma &= a + b, \sigma &\leq \mach \left|S\right|,
+  & \, \sigma &\leq \mach \left|a + b\right| \\
+\left[P, \pi\right] &= \mathtt{TwoProd}(a, b),
+  & \, P &= \fl{a \times b}, P + \pi &= a \times b,
+  \pi &\leq \mach \left|P\right|,
+  & \, \pi &\leq \mach \left|a \times b\right|.
+\end{alignat}
 The letters \(\sigma\) and \(\pi\) are used to indicate that the
@@
 \end{equation}
-where \(B_{j, n}(s) \geq 0\) and
+where \(B_{j, n}(s) = \binom{n}{j} (1 - s)^{n - j} s^j \geq 0\) and
 \(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).
@@
 \texttt{LocalError} in Algorithm~\ref{alg:local-error}}, we can
-perform \((K - 1)\) error filtrations:
+perform \((K - 1)\) error filtrations. Once we've computed the \(K\) stages
+of global errors, they can be combined with
+\texttt{SumK} (Algorithm~\ref{alg:sum-k}) to produce a sum that is as
+accurate as if computed in \(K\) times the working precision.
 
@@
 \widetilde{\ell}_{F, j} \coloneqq \left|e_1\right| +
- + \cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
+\cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
 \cdb{F - 1}_j^{(k + 1)}\right|.
@@
 \begin{equation}
-\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
-\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)},
+\mathtt{CompDeCasteljau}(p, s, K) = \mathtt{SumK}\left(\left[
+  \widehat{b}_0^{(0)}, \ldots, \cdb{K - 1}_0^{(0)}\right], K\right),
 \end{equation}
@@
 against the condition number. To compute relative errors, each input and
-coefficient is converted to a fraction (i.e. infinite precision) and then
+coefficient is converted to a fraction (i.e. infinite precision) and
+\(p(s)\) is computed exactly as a fraction, then
 compared to the corresponding computed values. Similar tools are used to
@@
 in Theorem~\ref{thm:eft}. They do not use branches, nor access to the
-mantissa that can be time-consuming
+mantissa that can be time-consuming.
 
@@
 \noindent The following algorithms from \cite{Ogita2005} can be used as a
-compensated method for computing a sum of \(n\) numbers. The first is a vector
+compensated method for computing a sum of numbers. The first is a vector
 transformation that is used as a helper:
@@
       \For{\(j = 2, \ldots, n\)}
-        \State \(\left[p_i, p_{i - 1}\right] = \mathtt{TwoSum}\left(
-            p_i, p_{i - 1}\right)\)
+        \State \(\left[p_j, p_{j - 1}\right] = \mathtt{TwoSum}\left(
+            p_j, p_{j - 1}\right)\)
       \EndFor
@@
 \(e_1, \ldots, e_{5F + 3}\) to the ``previous'' errors \(e_1',
-\ldots, e_{5F - 2}'\). In the same fashion as above, we track where
-the current errors come from:
+\ldots, e_{5F - 2}'\) that show up in \(\widetilde{\ell}_F\). In the same
+fashion as above, we track where the current errors come from:
 \begin{align}
@@
 \end{equation}
-which will complete the proof. To see this, first note that for a polynomial
-in this basis
+which will complete the proof of the first inequality. To see this, first
+note that for a polynomial in this basis
 \(f(k) = A \binom{k}{d} + B \binom{k}{d - 1} + C \binom{k}{d - 2} +
@@
   \sum_{j = 1}^F c_{j - 1} \binom{k}{j}\right] +
-(5F) \sum_{j = 0}^{F - 1} c_j \binom{k}{j} =
+5F \left[\sum_{j = 0}^{F - 1} c_j \binom{k}{j}\right] =
 3 c_{F - 1} \binom{k}{F} + \cdots
@@
 \(\gamma_{3k + 5F} L_{F, k} \leq (3k + 5F) L_{F, k} \mach +
-\bigO{\mach^{F + 1}} \widetilde{p}(s)\) it's enough to consider
+\bigO{\mach^{F + 2}} \widetilde{p}(s)\) it's enough to consider
 \begin{equation}

### 2018-05-08 scripts/compensated_insufficient.py
Removing color from figures.

@@ -44,9 +44,9 @@ def main(filename=None):
         exact.append(float(exact_p))
 
     figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
-
-    ax1.plot(s_vals, de_casteljau2)
-    ax2.plot(s_vals, exact)
+    alpha = 0.75
+    ax1.plot(s_vals, de_casteljau2, color="black", alpha=alpha)
+    ax2.plot(s_vals, exact, color="black", alpha=alpha)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -84,5 +84,5 @@ def main(filename=None):
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white")
+    seaborn.set(style="white", palette="Greys")
     main(filename="compensated_insufficient.pdf")

### 2018-05-08 scripts/error_against_cond.py

@@ -411,6 +411,6 @@ def main_jlcs10(filename=None):
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white")
+    seaborn.set(style="white", palette="Greys")
     main(filename="de_casteljau_rel_error.pdf")
     main_jlcs10(filename="jlcs10_plot.pdf")

### 2018-05-08 scripts/horner_inferior.py

@@ -43,8 +43,9 @@ def main(filename=None):
         de_casteljau1.append(de_casteljau.basic(s, BEZIER_COEFFS))
 
     figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
-    ax1.plot(s_vals, horner1)
-    ax2.plot(s_vals, de_casteljau1)
+    alpha = 0.75
+    ax1.plot(s_vals, horner1, color="black", alpha=alpha)
+    ax2.plot(s_vals, de_casteljau1, color="black", alpha=alpha)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -79,5 +80,5 @@ def main(filename=None):
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white")
+    seaborn.set(style="white", palette="Greys")
     main(filename="horner_inferior.pdf")

### 2018-05-08 scripts/smooth_drawing.py

@@ -57,9 +57,10 @@ def main(filename=None):
         evaluated3.append(b3)
 
     figure, (ax1, ax2, ax3) = plt.subplots(1, 3, sharex=True)
-    ax1.plot(s_vals, evaluated1)
-    ax2.plot(s_vals, evaluated2)
-    ax3.plot(s_vals, evaluated3)
+    alpha = 0.75
+    ax1.plot(s_vals, evaluated1, color="black", alpha=alpha)
+    ax2.plot(s_vals, evaluated2, color="black", alpha=alpha)
+    ax3.plot(s_vals, evaluated3, color="black", alpha=alpha)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -95,5 +96,5 @@ def main(filename=None):
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white")
+    seaborn.set(style="white", palette="Greys")
     main(filename="de_casteljau_smooth_drawing.pdf")

### 2018-05-08 scripts/compensated_insufficient.py
Changing colormap from Greys to gray.
The problem with "Greys" is that it started light and
got dark while "gray" does the other way around.

I also moved some shared behavior into a `plot_utils` module
to avoid the copy-pasta with the setting of the palette.

@@ -16,10 +16,9 @@ import fractions
 
 import matplotlib.pyplot as plt
 import numpy as np
-import seaborn
 
 import de_casteljau
-import file_utils
+import plot_utils
 
 
 F = fractions.Fraction
@@ -44,9 +43,8 @@ def main(filename=None):
         exact.append(float(exact_p))
 
     figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
-    alpha = 0.75
-    ax1.plot(s_vals, de_casteljau2, color="black", alpha=alpha)
-    ax2.plot(s_vals, exact, color="black", alpha=alpha)
+    ax1.plot(s_vals, de_casteljau2)
+    ax2.plot(s_vals, exact)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -77,12 +75,12 @@ def main(filename=None):
             wspace=0.13,
             hspace=0.20,
         )
-        path = file_utils.get_path(filename)
+        path = plot_utils.get_path(filename)
         figure.savefig(path, bbox_inches="tight")
         print("Saved {}".format(filename))
         plt.close(figure)
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white", palette="Greys")
+    plot_utils.set_styles()
     main(filename="compensated_insufficient.pdf")

### 2018-05-08 scripts/error_against_cond.py

@@ -21,10 +21,9 @@ import fractions
 
 import matplotlib.pyplot as plt
 import numpy as np
-import seaborn
 
 import de_casteljau
-import file_utils
+import plot_utils
 
 
 F = fractions.Fraction
@@ -43,6 +42,7 @@ BEZIER_COEFFS = (
 )
 ROOT = 0.75
 POWER_VAL = 1.3
+ALPHA = 0.25
 
 
 def main(filename=None):
@@ -107,7 +107,6 @@ def main(filename=None):
 
     figure = plt.figure()
     ax = figure.gca()
-    alpha = 0.25
     ax.loglog(
         cond_nums,
         forward_errs1,
@@ -147,28 +146,28 @@ def main(filename=None):
         [min_x, max_x],
         [bound_coeff1 * min_x, bound_coeff1 * max_x],
         color="black",
-        alpha=alpha,
+        alpha=ALPHA,
         zorder=1,
     )
     ax.loglog(
         [min_x, max_x],
         [bound_coeff2 * min_x, bound_coeff2 * max_x],
         color="black",
-        alpha=alpha,
+        alpha=ALPHA,
         zorder=1,
     )
     ax.loglog(
         [min_x, max_x],
         [bound_coeff3 * min_x, bound_coeff3 * max_x],
         color="black",
-        alpha=alpha,
+        alpha=ALPHA,
         zorder=1,
     )
     ax.loglog(
         [min_x, max_x],
         [bound_coeff4 * min_x, bound_coeff4 * max_x],
         color="black",
-        alpha=alpha,
+        alpha=ALPHA,
         zorder=1,
     )
     # Add the ``x = 1/u^k`` vertical lines.
@@ -180,7 +179,7 @@ def main(filename=None):
             [min_y - 0.05 * delta_y, max_y + 0.05 * delta_y],
             color="black",
             linestyle="dashed",
-            alpha=alpha,
+            alpha=ALPHA,
             zorder=1,
         )
     # Add the ``y = u`` horizontal lines.
@@ -189,7 +188,7 @@ def main(filename=None):
         [float(U), float(U)],
         color="black",
         linestyle="dashed",
-        alpha=alpha,
+        alpha=ALPHA,
         zorder=1,
     )
 
@@ -241,7 +240,7 @@ def main(filename=None):
     if filename is None:
         plt.show()
     else:
-        path = file_utils.get_path(filename)
+        path = plot_utils.get_path(filename)
[…]

### 2018-05-08 scripts/horner_inferior.py

@@ -18,11 +18,10 @@ and show that more accuracy produces smoother plots.
 
 import matplotlib.pyplot as plt
 import numpy as np
-import seaborn
 
 import de_casteljau
 import horner
-import file_utils
+import plot_utils
 
 # p(s) = (2s - 1)^3 = (-(1 - s) + s)^3
 POLY_COEFFS = (8.0, -12.0, 6.0, -1.0)
@@ -43,9 +42,8 @@ def main(filename=None):
         de_casteljau1.append(de_casteljau.basic(s, BEZIER_COEFFS))
 
     figure, (ax1, ax2) = plt.subplots(1, 2, sharex=True, sharey=True)
-    alpha = 0.75
-    ax1.plot(s_vals, horner1, color="black", alpha=alpha)
-    ax2.plot(s_vals, de_casteljau1, color="black", alpha=alpha)
+    ax1.plot(s_vals, horner1)
+    ax2.plot(s_vals, de_casteljau1)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -73,12 +71,12 @@ def main(filename=None):
             wspace=0.13,
             hspace=0.20,
         )
-        path = file_utils.get_path(filename)
+        path = plot_utils.get_path(filename)
         figure.savefig(path, bbox_inches="tight")
         print("Saved {}".format(filename))
         plt.close(figure)
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white", palette="Greys")
+    plot_utils.set_styles()
     main(filename="horner_inferior.pdf")

### 2018-05-08 scripts/smooth_drawing.py

@@ -18,10 +18,9 @@ and show that more accuracy produces smoother plots.
 
 import matplotlib.pyplot as plt
 import numpy as np
-import seaborn
 
 import de_casteljau
-import file_utils
+import plot_utils
 
 
 # p(s) = (s - 1) (s - 3/4)^7
@@ -57,10 +56,9 @@ def main(filename=None):
         evaluated3.append(b3)
 
     figure, (ax1, ax2, ax3) = plt.subplots(1, 3, sharex=True)
-    alpha = 0.75
-    ax1.plot(s_vals, evaluated1, color="black", alpha=alpha)
-    ax2.plot(s_vals, evaluated2, color="black", alpha=alpha)
-    ax3.plot(s_vals, evaluated3, color="black", alpha=alpha)
+    ax1.plot(s_vals, evaluated1)
+    ax2.plot(s_vals, evaluated2)
+    ax3.plot(s_vals, evaluated3)
 
     # Since ``sharex=True``, ticks only need to be set once.
     ax1.set_xticks(
@@ -89,12 +87,12 @@ def main(filename=None):
             wspace=0.15,
             hspace=0.19,
         )
-        path = file_utils.get_path(filename)
+        path = plot_utils.get_path(filename)
         figure.savefig(path, bbox_inches="tight")
         print("Saved {}".format(filename))
         plt.close(figure)
 
 
 if __name__ == "__main__":
-    seaborn.set(style="white", palette="Greys")
+    plot_utils.set_styles()
     main(filename="de_casteljau_smooth_drawing.pdf")

### 2018-05-08 src/plot_utils.py

@@ -10,9 +10,17 @@
 # See the License for the specific language governing permissions and
 # limitations under the License.
 
-"""Helper(s) for dealing with files."""
+"""Shared utilities and settings for plotting."""
+
 
 import os
+
+import seaborn
+
+
+def set_styles():
+    """Set the styles used for plotting."""
+    seaborn.set(style="white", palette="gray")
 
 
 def get_path(filename):

### 2018-05-08 doc/paper.tex
Another "review" pass.

@@
 
-The paper is organized as follows. Section~\ref{sec:notation} introduces some
-basic notation and results about floating point arithmetic, error-free
-transformations and the de Casteljau algorithm. In
-Section~\ref{sec:compensated-2},
+The paper is organized as follows. Section~\ref{sec:notation} establishes
+notation for error analysis with floating point operations, reviews
+results about error-free transformations and reviews the
+de Casteljau algorithm. In Section~\ref{sec:compensated-2},
 the compensated algorithm for polynomial evaluation from \cite{Jiang2010} is
 reviewed and notation is established for the expansion. In
-Section~\ref{sec:compensated-k}, the \(K\) compensated algorithm is provided and
-a forward error analysis is performed. Finally, in Section~\ref{sec:numerical} we
-give numerical tests to illustrate the practical efficiency of our algorithms.
+Section~\ref{sec:compensated-k}, the \(K\) compensated algorithm is provided
+and a forward error analysis is performed. Finally, in
+Section~\ref{sec:numerical} we perform two numerical experiments to
+give practical examples of the theoretical error bounds.
 
@@
 is considered ``free'' of error if the round-off can be represented
-exactly as an element of \(\mathbf{F}\).
+exactly as an element or elements of \(\mathbf{F}\).
 The error-free transformations used in this paper are
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{DeCasteljau}\)}{$b, s$}
+    \Function{\(\mathtt{result} = \mathtt{DeCasteljau}\)}{$b, s$}
       \State \(n = \texttt{length}(b) - 1\)
@@
           \State \(\widehat{b}_j^{(k)} = \left(
-              r \otimes \widehat{b}_j^{(k + 1)}\right) \oplus
+              \widehat{r} \otimes \widehat{b}_j^{(k + 1)}\right) \oplus
               \left(s \otimes \widehat{b}_{j + 1}^{(k + 1)}\right)\)
@@
       \\
-      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
+      \State \(\mathtt{result} = \widehat{b}_0^{(0)}\)
     \EndFunction
@@
 
-In general, for a sequence \(v_0, \ldots, v_n\) we'll use the notation
+In general, for a sequence \(v_0, \ldots, v_n\) we'll refer to \(v\)
+as the vector containing all of the values:
 \(v = \left[\begin{array}{c c c} v_0 & \cdots &
-    v_n\end{array}\right]^T\) to indicate the vector containing all of
-the values.
+    v_n\end{array}\right]^T.\)
 
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljau}\)}{$b, s$}
+    \Function{\(\mathtt{result} = \mathtt{CompDeCasteljau}\)}{$b, s$}
       \State \(n = \texttt{length}(b) - 1\)
@@
       \\
-      \State \(\mathtt{res} = \widehat{b}_0^{(0)} \oplus
+      \State \(\mathtt{result} = \widehat{b}_0^{(0)} \oplus
           \cdb{1}_0^{(0)}\)
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljauK}\)}{$b, s, K$}
+    \Function{\(\mathtt{result} = \mathtt{CompDeCasteljauK}\)}{$b, s, K$}
       \State \(n = \texttt{length}(b) - 1\)
@@
       \\
-      \State \(\mathtt{res} = \mathtt{SumK}\left(\left[
+      \State \(\mathtt{result} = \mathtt{SumK}\left(\left[
         \widehat{b}_0^{(0)}, \ldots, \cdb{K - 1}_0^{(0)}\right], K\right)\)
@@
 \(\widehat{b}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}\)
-instead of \texttt{SumK}. This will reduce the total by
-\(6(K - 1)^2\) flops, however the results produced are (empirically)
-identical. This makes sense: the whole point of \texttt{SumK}
+instead of \texttt{SumK}; this will reduce the total by
+\(6(K - 1)^2\) flops. When using a standard sum, the results produced
+are (empirically) identical to those with \texttt{SumK}. This makes
+sense: the whole point of \texttt{SumK}
 is to filter errors in a summation so that the final operation produces
@@
 \begin{lemma}\label{lemma:ell-tilde}
-The local errors \(\widetilde{\ell}_{F, j}\) satisfy:
+The local error bounds \(\widetilde{\ell}_{F, j}\) satisfy:
 \begin{align}
@@
 \begin{lemma}\label{lemma:L-and-D-bounds}
-For \(s \in \left[0, 1\right]\), the Bernstein-type sum defined above
+For \(s \in \left[0, 1\right]\), the Bernstein-type error sum defined above
 satisfies the following bounds:
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{VecSum}\)}{$p$}
+    \Function{\(\mathtt{VecSum}\)}{$p$}
       \State \(n = \texttt{length}(p)\)
@@
   \begin{algorithmic}
-    \Function{\(\mathtt{res} = \mathtt{SumK}\)}{$p, K$}
+    \Function{\(\mathtt{result} = \mathtt{SumK}\)}{$p, K$}
       \For{\(j = 1, \ldots, K - 1\)}
@@
       \EndFor
-      \State \(\mathtt{res} = p_1 \oplus p_2 \oplus \cdots \oplus p_n\)
+      \State \(\mathtt{result} = p_1 \oplus p_2 \oplus \cdots \oplus p_n\)
     \EndFunction

### 2018-05-08 scripts/compute_counts.py
A few Pylint fixes (though there are many more unfixed).x

@@ -101,7 +101,7 @@ def count_horner_basic():
 
 
 def horner_expected_total(K, n):
-    """Get the expected flop count for compensated Horner's method.
+    r"""Get the expected flop count for compensated Horner's method.
 
     When using FMA, the count is
 
@@ -113,7 +113,7 @@ def horner_expected_total(K, n):
 
 
 def horner_expected_fma(K, n):
-    """Get the FMA count for compensated Horner's method.
+    r"""Get the FMA count for compensated Horner's method.
 
     When using FMA, the count is
 

### 2018-05-08 src/de_casteljau.py

@@ -10,7 +10,7 @@
 # See the License for the specific language governing permissions and
 # limitations under the License.
 
-"""Performs de Casteljau's method.
+r"""Performs de Casteljau's method.
 
 de Casteljau's method evaluates a function in Bernstein-Bezier form
 
@@ -42,6 +42,7 @@ import eft
 
 
 def basic(s, coeffs):
+    """Performs the "standard" de Casteljau algorithm."""
     r = 1.0 - s
 
     degree = len(coeffs) - 1

### 2018-05-08 src/horner.py

@@ -10,7 +10,7 @@
 # See the License for the specific language governing permissions and
 # limitations under the License.
 
-"""Performs Horner's method.
+r"""Performs Horner's method.
 
 Horner's method computes
 

### 2018-05-09 nox.py
Adding note about needing `bibtex`.
Also making sure to use Python 3.6 in `nox - s build_tex`.
It actually runs Python code when modifying the PDF ID.

@@ -82,6 +82,8 @@ def build_tex_file(session, base, new_id, extensions=(), with_bibtex=False):
 
 @nox.session
 def build_tex(session):
+    session.interpreter = SINGLE_INTERP
+
     if py.path.local.sysfind("pdflatex") is None:
         session.skip("`pdflatex` must be installed")
 

### 2018-05-09 doc/paper.tex
Fixing "a prior" --> "a priori" typo.
Also adding a basic implementation of the VS method (though
slightly modified to avoid any division, except for the division
in binomial coefficients).

@@
 \texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
-that satisfies the following a prior bound for any arbitrary integer \(K\):
+that satisfies the following a priori bound for any arbitrary integer \(K\):
 \begin{equation}

### 2018-05-09 scripts/compute_counts.py

@@ -18,6 +18,7 @@ import de_casteljau
 import eft
 import horner
 import operation_count
+import vs_method
 
 
 SEPARATOR = "-" * 80
@@ -86,6 +87,21 @@ def count_sum_k():
             assert total.value == float(size_p)
             assert parent.count == (6 * k - 5) * (size_p - 1)
             print("    |p| = {}:    {}".format(size_p, parent.display))
+
+
+def count_vs_method_basic():
+    print("vs_method.basic() (7n - 2):")
+    for degree in range(1, 5 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = vs_method.basic(x, coeffs)
+        assert p.value == 0.5 ** degree
+        assert parent.count == 7 * degree - 2
+        print("  degree {}:     {}".format(degree, parent.display))
 
 
 def count_horner_basic():
@@ -337,6 +353,8 @@ def main():
     count__vec_sum()
     print(SEPARATOR)
     count_sum_k()
+    print(SEPARATOR)
+    count_vs_method_basic()
     print(SEPARATOR)
     count_horner_basic()
     print(SEPARATOR)

### 2018-05-09 src/operation_count.py

@@ -70,6 +70,12 @@ class Float(object):
             return other.value
         elif isinstance(other, float):
             return other
+        elif isinstance(other, int):
+            other_float = float(other)
+            if other == other_float:
+                return other_float
+            else:
+                return None
         else:
             return None
 
@@ -116,6 +122,22 @@ class Float(object):
         self.computation.mul_count += 1
         return Float(self.value * value, self.computation)
 
+    def __rmul__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.mul_count += 1
+        return Float(value * self.value, self.computation)
+
+    def __truediv__(self, other):
+        value = self._get_value(other)
+        if value is None:
+            return NotImplemented
+
+        self.computation.mul_count += 1
+        return Float(self.value / value, self.computation)
+
     def fma(self, val1, val2, val3):
         float1 = self._get_value(val1)
         float2 = self._get_value(val2)
@@ -129,3 +151,25 @@ class Float(object):
         frac3 = fractions.Fraction(float3)
         result = float(frac1 * frac2 + frac3)
         return Float(result, self.computation)
+
+
+def as_type(value, typed_value):
+    """Converts a basic float to a type.
+
+    Args:
+        value (float): The value to convert to a type.
+        typed_value (object): The type to match (and the value
+            may contain come metadata).
+
+    Returns:
+        object: A wrapper of ``value`` with the same type as ``typed_value``.
+
+    Raises:
+        TypeError: If ``typed_value`` is not a ``float`` or ``Float``.
+    """
+    if isinstance(typed_value, float):
+        return value
+    elif isinstance(typed_value, Float):
+        return Float(value, typed_value.computation)
+    else:
+        raise TypeError("Unexpected type", typed_value)

### 2018-05-09 src/vs_method.py

@@ -1,0 +1,42 @@
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+r"""Evaluates a polynomial in Bernstein form via the `VS method`_.
+
+.. _VS method: https://doi.org/10.1016/0167-8396(86)90018-X
+.. _survey: http://dx.doi.org/10.1016/j.amc.2015.08.086
+
+This method has been described in the 1D case in a `survey`_
+paper that followed the original formulation.
+"""
+
+
+import operation_count
+
+
+def basic(s, coeffs):
+    degree = len(coeffs) - 1
+    r = 1.0 - s
+
+    result = r * coeffs[0]
+
+    binom_val = operation_count.as_type(1.0, s)
+    s_pow = operation_count.as_type(1.0, s)
+    for j in range(1, degree):
+        s_pow = s * s_pow
+        binom_val = binom_val * (degree - j + 1)
+        binom_val = binom_val / j
+        result = result + binom_val * s_pow * coeffs[j]
+        result = result * r
+
+    result = result + s * s_pow * coeffs[degree]
+    return result

### 2018-05-10 src/plot_utils.py
Switching back to default palette in ``seaborn``.
I.e. adding color back to plots.

@@ -20,7 +20,7 @@ import seaborn
 
 def set_styles():
     """Set the styles used for plotting."""
-    seaborn.set(style="white", palette="gray")
+    seaborn.set(style="white")
 
 
 def get_path(filename):

### 2018-05-15 doc/paper.tex
Updating manuscript for submission.
Did so by using the `elsarticle` class (and noting which
journal it is targeting).

@@
-\documentclass[letterpaper,10pt]{article}
-
-\usepackage[margin=1in]{geometry}
+\documentclass[1p]{elsarticle}
+\geometry{margin=1in}
 
@@
 \usepackage[labelfont=bf]{caption}
-%% H/T: https://tex.stackexchange.com/a/71153/32270
-\usepackage[nottoc,notlot,notlof]{tocbibind}
+%% H/T: https://tex.stackexchange.com/a/163250/32270
+\usepackage{adjustbox}
+%% H/T: https://tex.stackexchange.com/a/26348/32270
+\renewcommand*{\appendixname}{}
 
@@
   colorlinks=true,
-  urlcolor=blue,
-  linkcolor=MidnightBlue,
-  citecolor=ForestGreen,
   pdfinfo={
@@
 
-\usepackage{embedfile}
-\embedfile{\jobname.tex}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{\(K\)-compensated de Casteljau}
-\rhead{Danny Hermes}
-
-\renewcommand{\headrulewidth}{0pt}
 \renewcommand{\qed}{\(\blacksquare\)}
@@
 
+\journal{Applied Mathematics and Computation}
+
 \begin{document}
+%% H/T: https://tex.stackexchange.com/a/263503/32270
+\hypersetup{
+  linkcolor=MidnightBlue,
+  citecolor=ForestGreen,
+}
+
+\begin{frontmatter}
+
+\title{\(K\)-compensated de Casteljau}
+\author[djh]{Danny Hermes}\ead{dhermes@berkeley.edu}
+\address[djh]{UC Berkeley, 970 Evans Hall \#3840, Berkeley, CA 94720-3840 USA}
 
@@
 
-\tableofcontents
+\begin{keyword}
+Polynomial evaluation \sep Compensated algorithm \sep
+Floating-point arithmetic \sep Bernstein polynomial \sep
+Error-free transformation \sep Round-off error
+\end{keyword}
+
+\end{frontmatter}
 
@@
 reviewed and notation is established for the expansion. In
-Section~\ref{sec:compensated-k}, the \(K\) compensated algorithm is provided
+Section~\ref{sec:compensated-k}, the \(K\)-compensated algorithm is provided
 and a forward error analysis is performed. Finally, in
@@
   \centering
+  \begin{adjustbox}{max width=\textwidth}
   \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
@@
   \end{tabular}
+  \end{adjustbox}
   \captionsetup{width=.75\linewidth}
@@
 
-\section{\texorpdfstring{\(K\)}{K} Compensated de Casteljau}\label{sec:compensated-k}
+\section{\texorpdfstring{\(K\)}{K}-Compensated de Casteljau}\label{sec:compensated-k}
 
@@
 
-\section{Appendix: Algorithms}\label{sec:appendix-algo}
+\appendix
+
+\section{Algorithms}\label{sec:appendix-algo}
 
@@
 
-\section{Appendix: Proof Details}\label{sec:appendix-proof-details}
+\section{Proof Details}\label{sec:appendix-proof-details}
 

### 2018-05-15 nox.py

@@ -57,22 +57,16 @@ class Remove(object):
             os.remove(path)
 
 
-def build_tex_file(session, base, new_id, extensions=(), with_bibtex=False):
+def build_tex_file(session, base, new_id, extensions=()):
     # NOTE: This assumes that ``session.chdir(get_path('doc'))``
     #       has been called.
     modify_id = get_path("scripts", "modify_pdf_id.py")
 
-    if with_bibtex:
-        session.run("pdflatex", base)
-        session.run("bibtex", base)
-        session.run("pdflatex", base)
-        session.run("bibtex", base)
-        session.run("pdflatex", base)
-    else:
-        # Run ``pdflatex`` thrice.
-        session.run("pdflatex", base)
-        session.run("pdflatex", base)
-        session.run("pdflatex", base)
+    session.run("pdflatex", base)
+    session.run("bibtex", base)
+    session.run("pdflatex", base)
+    session.run("bibtex", base)
+    session.run("pdflatex", base)
 
     path = get_path("doc", base)
     remove = Remove(path, extensions)
@@ -99,8 +93,7 @@ def build_tex(session):
         session,
         "paper",
         "F092359D979FDC08931DA1922F3E123E",
-        extensions=("aux", "log", "out", "toc"),
-        with_bibtex=True,
+        extensions=("aux", "bbl", "blg", "log", "out", "spl"),
     )
 
 

### 2018-05-15 scripts/error_against_cond.py

@@ -255,8 +255,8 @@ def main_jlcs10(filename=None):
     code could be shared, but it isn't worth the effort).
 
     This seeks to recreate the plot from the original paper, but to show
-    what happens as the exponent on ``1.3`` decreases from ``-44`` down to
-    ``-64``. In particular, it shows that the compensated de Casteljau
+    what happens as the exponent on ``POWER_VAL`` decreases from ``-44`` down
+    to ``-64``. In particular, it shows that the compensated de Casteljau
     algorithm produces exactly zero.
     """
     n = 8

### 2018-05-16 doc/paper.tex
A few small tweaks.
- Used a longer (more descriptive) title
- Changed from 1p to 3p (and ditched the hack to change margins
  to 1in)
- Used "round-off" everywhere (instead of "roundoff" sometimes)
- Built the doc one extra time with `pdflatex` (was getting a
  warning about changed references on the final run)
- Upper-casing a month name in the bibliography
- Switching bibliography style to "elsarticle-harv" and using
  "authoryear" for citations (also making MidnightBlue the URL
  color for references, since it isn't as loud as "blue")
- Adding "References" as a section header for the bibli

@@
-\documentclass[1p]{elsarticle}
-\geometry{margin=1in}
+\documentclass[3p, authoryear, square]{elsarticle}
+\bibliographystyle{elsarticle-harv}
 
@@
 \hypersetup{
+  urlcolor=MidnightBlue,
   linkcolor=MidnightBlue,
@@
 
-\title{\(K\)-compensated de Casteljau}
+\title{Compensated de Casteljau algorithm in \(K\) times the working precision}
 \author[djh]{Danny Hermes}\ead{dhermes@berkeley.edu}
@@
 arithmetic operations grows quadratically. The Bernstein basis is
-optimally suited (\cite{Farouki1987}, \cite{Delgado2015}, \cite{Mainar2005})
+optimally suited (\cite{Farouki1987, Delgado2015, Mainar2005})
 for polynomial evaluation; it is
@@
 To address ill-conditioned problems, error-free transformations (EFT) can
-be applied in \textit{compensated algorithms} to account for roundoff.
+be applied in \textit{compensated algorithms} to account for round-off.
 Error-free transformations were studied in great detail in \cite{Ogita2005}
@@
 causes the number of operations to grow exponentially in \(K\). Here, we
-instead have a fixed number of error groups, each corresponding to roundoff
+instead have a fixed number of error groups, each corresponding to round-off
 from the group above it. For example, when
@@
 The letters \(\sigma\) and \(\pi\) are used to indicate that the
-errors came from sum and product, respectively. See the appendix
-(Section~\ref{sec:appendix-algo}) for implementation details.
+errors came from sum and product, respectively. See
+Appendix~\ref{sec:appendix-algo} for implementation details.
 \end{theorem}
@@
 and \texttt{TwoProd} when computing \(\cdb{F - 1}_j^{(k)}\)
-and the \(\rho\) term comes from the roundoff
+and the \(\rho\) term comes from the round-off
 in \(1 \ominus s\) when multiplying \((1 - s)\) by
@@
 
-See the appendix (Section~\ref{sec:appendix-proof-details}) for details on
+See Appendix~\ref{sec:appendix-proof-details} for details on
 proving Lemma~\ref{lemma:ell-tilde} and Lemma~\ref{lemma:L-and-D-bounds}.
@@
 
+%% H/T: https://tex.stackexchange.com/a/137379/32270
+\section*{\refname}
 \bibliography{paper}
-\bibliographystyle{alpha}
 

### 2018-05-16 nox.py

@@ -67,6 +67,7 @@ def build_tex_file(session, base, new_id, extensions=()):
     session.run("pdflatex", base)
     session.run("bibtex", base)
     session.run("pdflatex", base)
+    session.run("pdflatex", base)
 
     path = get_path("doc", base)
     remove = Remove(path, extensions)

### 2018-05-16 doc/cover_letter.tex
Adding cover letter and doing final pass of manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,a4paper]{letter}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  pdfinfo={
    CreationDate={D:20180516090844},
    ModDate={D:20180516090844},
  },
}

% Create a new command for the horizontal rule in the document
% which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%------------------
% DOCUMENT MARGINS
%------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

%--------------------
% SENDER INFORMATION
%--------------------

\def\Who{Danny Hermes}
\def\School{UC Berkeley}
\def\Where{Department of Mathematics}
\def\Address{970 Evans Hall \#3840}
\def\CityZip{Berkeley CA 94720-3840}
\def\Email{dhermes@berkeley.edu}

%-----------------------------------
% HEADER AND FROM ADDRESS STRUCTURE
%-----------------------------------

\address{
\includegraphics[width=1in]{uc_berkeley_seal.pdf} % Include the logo of your institution
\hspace{5.1in} % Position of the institution logo, increase to move left, decrease to move right
\vskip -1.07in~\\ % Position of the text in relation to the institution logo, increase to move down, decrease to move up
\Large\hspace{1.5in}UNIVERSITY \hfill ~\\[0.05in] % First line of institution name, adjust hspace if your logo is wide
\hspace{1.5in}OF CALIFORNIA \hfill \normalsize % Second line of institution name, adjust hspace if your logo is wide
\makebox[0ex][r]{\bf \Who}\hspace{0.08in} % Print your name and title with a little whitespace to the right
~\\[-0.11in] % Reduce the whitespace above the horizontal rule
\hspace{1.5in}\vhrulefill{1pt} \\ % Horizontal rule, adjust hspace if your logo is wide and \vhrulefill for the thickness of the rule
\hspace{\fill}\parbox[t]{2.85in}{ % Create a box for your details underneath the horizontal rule on the right
\footnotesize % Use a smaller font size for the details
\textit{\School}  \\
\textit{\Where}   \\
\textit{\Address} \\
\textit{\CityZip} \\
\textit{\Email}   \\
}
\hspace{-1.4in} % Horizontal position of this block, increase to move left, decrease to move right
\vspace{-1in} % Move the letter content up for a more compact look
}

%----------------------
% TO ADDRESS STRUCTURE
%----------------------

\def\opening#1{\thispagestyle{empty}
{\centering\fromaddress \vspace{0.6in} \\ % Print the header and from address here, add whitespace to move date down
\hspace*{\longindentation}\today\hspace*{\fill}\par} % Print today's date, remove \today to not display it
{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{0.4in} % White space after the to address
\noindent #1 % Print the opening line
}

\signature{
  \includegraphics[width=2.5in]{daniel_hermes_signature.jpg}
}

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
\hspace*{\longindentation} % Move the signature right
\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.05in % Whitespace between the signature text and your name
\fromsig}} % Print your name and title

\begin{document}

\begin{letter}
{Editors \\
Applied Mathematics and Computation}

\opening{Dear Editors,}

I am pleased to submit the original research article
``Compensated de Casteljau algorithm in \(K\) times the working precision''
for consideration for publication in Applied Mathematics and Computation.
This manuscript builds on prior work of Jiang et al in evaluating a polynomial
in Bernstein form with greater accuracy. It follows Graillat et al in
producing a method for evaluation that produces results as accurate as if
performed in \(K\) times the working precision, then rounded back into the
working precision.

In this manuscript, error-free transformations for the sum and product
are used to track the \textbf{exact} round-off error throughout the
de Casteljau algorithm. At each stage of computation, round-off error is
passed on to first order errors, then to second order errors, and so on.
After the computation has been ``filtered'' \((K - 1)\) times via this
process, the resulting output is as accurate as the de Casteljau algorithm
performed in \(K\) times the working precision.

I believe that this manuscript is appropriate for publication by
Applied Mathematics and Computation because it describes a computational
method for extending the accuracy of a well-studied algorithm to
arbitrarily high precision. The accompanying error analysis confirms
the numerical results.

I confirm that this manuscript has not been published elsewhere and is not
under consideration by another journal. All authors have approved the
manuscript and agree with its submission to Applied Mathematics and
Computation.

Thank you for your consideration!

\closing{Sincerely,}

\end{letter}
\end{document}

### 2018-05-16 doc/paper.tex
Adding cover letter and doing final pass of manuscript.

@@
   pdfinfo={
-    CreationDate={D:20180425160342},
-    ModDate={D:20180425160342},
+    CreationDate={D:20180516090844},
+    ModDate={D:20180516090844},
   },
@@
 
-To enable a bound on the \(K\)-order error \(\db{K}^{(0)}_0\), it's necessary
+To enable a bound on the \(K\) order error \(\db{K}^{(0)}_0\), it's necessary
 to understand the difference between the exact local errors \(\ell_{F, j}\)
@@
 \begin{lemma}\label{lemma:k-order}
-The \(K\)-order error \(\db{K}^{(0)}_0\) satisfies
+The \(K\) order error \(\db{K}^{(0)}_0\) satisfies
 \begin{equation}
@@
 
-All experiments were performed in IEEE-754 double precision.
-As in \cite{Jiang2010}, we consider the evaluation in the neighborhood
-of the multiple root of \(p(s) = (s - 1)\left(s - \frac{3}{4}\right)^7\),
-written in Bernstein form.
-
 \begin{figure}
@@
 
+All experiments were performed in IEEE-754 double precision.
+As in \cite{Jiang2010}, we consider the evaluation in the neighborhood
+of the multiple root of \(p(s) = (s - 1)\left(s - 3/4\right)^7\),
+written in Bernstein form.
 Figure~\ref{fig:smooth-drawing} shows the evaluation of \(p(s)\) at
@@
 on the fact that \(\widetilde{p}(s) = (s - 1)
-\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
+\left(s/2 - 3/4\right)^7\). Once the relative errors and
 condition numbers are computed as fractions, they are rounded to the

### 2018-05-16 nox.py

@@ -30,14 +30,14 @@ MATPLOTLIB_DEPS = (
     "matplotlib == 2.2.2",
     "numpy == 1.14.3",
     "pyparsing == 2.2.0",
-    "python-dateutil == 2.7.2",
+    "python-dateutil == 2.7.3",
     "pytz == 2018.4",
     "six == 1.11.0",
 )
 DEPS = {
     "matplotlib": MATPLOTLIB_DEPS,
     "seaborn": MATPLOTLIB_DEPS
-    + ("pandas == 0.22.0", "scipy == 1.1.0", "seaborn == 0.8.1"),
+    + ("pandas == 0.23.0", "scipy == 1.1.0", "seaborn == 0.8.1"),
 }
 
 
@@ -57,17 +57,22 @@ class Remove(object):
             os.remove(path)
 
 
-def build_tex_file(session, base, new_id, extensions=()):
+def build_tex_file(session, base, new_id, extensions=(), with_bibtex=False):
     # NOTE: This assumes that ``session.chdir(get_path('doc'))``
     #       has been called.
     modify_id = get_path("scripts", "modify_pdf_id.py")
 
-    session.run("pdflatex", base)
-    session.run("bibtex", base)
-    session.run("pdflatex", base)
-    session.run("bibtex", base)
-    session.run("pdflatex", base)
-    session.run("pdflatex", base)
+    if with_bibtex:
+        session.run("pdflatex", base)
+        session.run("bibtex", base)
+        session.run("pdflatex", base)
+        session.run("bibtex", base)
+        session.run("pdflatex", base)
+        session.run("pdflatex", base)
+    else:
+        session.run("pdflatex", base)
+        session.run("pdflatex", base)
+        session.run("pdflatex", base)
 
     path = get_path("doc", base)
     remove = Remove(path, extensions)
@@ -95,6 +100,14 @@ def build_tex(session):
         "paper",
         "F092359D979FDC08931DA1922F3E123E",
         extensions=("aux", "bbl", "blg", "log", "out", "spl"),
+        with_bibtex=True,
+    )
+
+    build_tex_file(
+        session,
+        "cover_letter",
+        "BACA8D659970198BDF7D11B67FEA6299",
+        extensions=("aux", "log", "out"),
     )
 
 

### 2018-05-16 scripts/error_against_cond.py

@@ -53,9 +53,9 @@ def main(filename=None):
     bound_coeff2 = float(bound_coeff2)
     bound_coeff3 = 3 * n * (3 * n ** 2 + 36 * n + 61) * U ** 3 / 2
     bound_coeff3 = float(bound_coeff3)
-    bound_coeff4 = 9 * n * (
-        3 * n ** 3 + 102 * n ** 2 + 773 * n + 1122
-    ) * U ** 4 / 8
+    bound_coeff4 = (
+        9 * n * (3 * n ** 3 + 102 * n ** 2 + 773 * n + 1122) * U ** 4 / 8
+    )
     bound_coeff4 = float(bound_coeff4)
 
     cond_nums = []

### 2018-05-16 doc/paper.tex
Making image symlinks in doc directory.
Referring to these images in `paper.tex`. This will
make it easier on the journal to build the source.

@@
 \begin{figure}
-  \includegraphics[width=0.9375\textwidth]{../images/horner_inferior.pdf}
+  \includegraphics[width=0.9375\textwidth]{horner_inferior.pdf}
   \centering
@@
 \begin{figure}
-  \includegraphics[width=0.8125\textwidth]{../images/jlcs10_plot.pdf}
+  \includegraphics[width=0.8125\textwidth]{jlcs10_plot.pdf}
   \centering
@@
 \begin{figure}
-  \includegraphics[width=0.9375\textwidth]{../images/compensated_insufficient.pdf}
+  \includegraphics[width=0.9375\textwidth]{compensated_insufficient.pdf}
   \centering
@@
 \begin{figure}
-  \includegraphics[width=0.9375\textwidth]{../images/de_casteljau_smooth_drawing.pdf}
+  \includegraphics[width=0.9375\textwidth]{de_casteljau_smooth_drawing.pdf}
   \centering
@@
 \begin{figure}
-  \includegraphics[width=0.8125\textwidth]{../images/de_casteljau_rel_error.pdf}
+  \includegraphics[width=0.8125\textwidth]{de_casteljau_rel_error.pdf}
   \centering

### 2018-05-16 doc/paper.tex
Isolating Tikz images as standalone PDFs.
This is both a way to "streamline" the building of the main paper
and an effort to "pass" the build on EES.

@@
 
-%% H/T: https://tex.stackexchange.com/a/99051/32270
-\PassOptionsToPackage{usenames,dvipsnames}{xcolor}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-\tikzstyle{box} = [
-  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
-  text centered, draw=black]
-\tikzstyle{empty-box} = [
-  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
-  text width=0.05\textwidth]
-\tikzstyle{empty-box-vert} = [
-  rectangle, minimum width=0.1\textwidth, minimum height=0.1\textwidth,
-  text centered, text height=0.001\textwidth]
-\tikzstyle{line} = [draw, -latex']
-
 \renewcommand{\qed}{\(\blacksquare\)}
@@
 \begin{figure}
+  \includegraphics[width=0.375\textwidth]{tikz_local_err.pdf}
   \centering
-\begin{tikzpicture}[node distance = 0.2\textwidth]
-  \node (b0) {\(\widehat{b}^{(n)}\)}
-    child {node (b1) {\(\widehat{b}^{(n - 1)}\)}
-      child {node (b2) {\(\widehat{b}^{(n - 2)}\)}
-        child {node (bdots) {\(\vdots\)}
-          child {node (bn) {\(\widehat{b}^{(0)}\)}}
-          child {node (ln) {\(\ell^{(0)}\)}}
-        }
-        child {node (ldots) {\(\vdots\)}}
-      }
-      child {node (l2) {\(\ell^{(n - 2)}\)}}
-    }
-    child {node (l1) {\(\ell^{(n - 1)}\)}};
-  %%
-  \path (b1) -- (l1) node [midway] {\(+\)};
-  \path (b2) -- (l2) node [midway] {\(+\)};
-  \path (bn) -- (ln) node [midway] {\(+\)};
-  %% eb == exact b
-  \node [right of=b0] (eb0) {\(b^{(n)}\)}
-    child {node (eb1) {\(b^{(n - 1)}\)}
-      child {node (eb2) {\(b^{(n - 2)}\)}
-        child {node (ebdots) {\(\vdots\)}
-          child {node (ebn) {\(b^{(0)}\)}}
-        }
-      }
-    };
-\end{tikzpicture}
   \captionsetup{width=.75\linewidth}
@@
 \begin{figure}
+  \includegraphics[width=0.9375\textwidth]{tikz_filtration.pdf}
   \centering
-\begin{tikzpicture}[node distance = 0.2\textwidth]
-    \node [box] (bj) {\(b\)};
-    \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
-    \node [box, right of=bn0] (dbn0) {\(\cdb{1}^{(n)}\)};
-    \node [box, right of=dbn0] (ddbn0) {\(\cdb{2}^{(n)}\)};
-    \node [box, right of=ddbn0] (dddbn0) {\(\cdb{3}^{(n)}\)};
-    %%
-    \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
-    \node [box, below of=dbn0] (dbn1) {\(\cdb{1}^{(n - 1)}\)};
-    \node [box, below of=ddbn0] (ddbn1) {\(\cdb{2}^{(n - 1)}\)};
-    \node [box, below of=dddbn0] (dddbn1) {\(\cdb{3}^{(n - 1)}\)};
-    %%
-    \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
-    \node [box, below of=dbn1] (dbn2) {\(\cdb{1}^{(n - 2)}\)};
-    \node [box, below of=ddbn1] (ddbn2) {\(\cdb{2}^{(n - 2)}\)};
-    \node [box, below of=dddbn1] (dddbn2) {\(\cdb{3}^{(n - 2)}\)};
-    %%
-    \path [line] (bn0) -- node[above] {\(\widehat{\ell}_1^{(n)}\)} (dbn0);
-    \path [line] (dbn0) -- node[above] {\(\widehat{\ell}_2^{(n)}\)} (ddbn0);
-    \path [line] (ddbn0) -- node[above] {\(\widehat{\ell}_3^{(n)}\)} (dddbn0);
-    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
-    \path [line] (dddbn0) -- node[above] {\(\widehat{\ell}_4^{(n)}\)} (right0);
-    %%
-    \path [line] (bn1) -- node[above] {\(\widehat{\ell}_1^{(n - 1)}\)} (dbn1);
-    \path [line] (dbn1) -- node[above] {\(\widehat{\ell}_2^{(n - 1)}\)} (ddbn1);
-    \path [line] (ddbn1) -- node[above] {\(\widehat{\ell}_3^{(n - 1)}\)} (dddbn1);
-    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
-    \path [line] (dddbn1) -- node[above] {\(\widehat{\ell}_4^{(n - 1)}\)} (right1);
-    %%
-    \path [line] (bn2) -- node[above] {\(\widehat{\ell}_1^{(n - 2)}\)} (dbn2);
-    \path [line] (dbn2) -- node[above] {\(\widehat{\ell}_2^{(n - 2)}\)} (ddbn2);
-    \path [line] (ddbn2) -- node[above] {\(\widehat{\ell}_3^{(n - 2)}\)} (dddbn2);
-    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
-    \path [line] (dddbn2) -- node[above] {\(\widehat{\ell}_4^{(n - 2)}\)} (right2);
-    %%
-    \path [line] (bj) -- (bn0);
-    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
-    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
-    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
-    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
-    %%
-    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
-    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
-    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
-    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
-    %%
-    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
-    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
-    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
-    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
-    %%
-    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
-    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
-    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
-    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
-\end{tikzpicture}
   \captionsetup{width=.75\linewidth}

### 2018-05-16 doc/tikz_filtration.tex
Isolating Tikz images as standalone PDFs.
This is both a way to "streamline" the building of the main paper
and an effort to "pass" the build on EES.

\documentclass[tikz]{standalone}

\usepackage{hyperref}
\hypersetup{
  pdfinfo={
    CreationDate={D:20180516090844},
    ModDate={D:20180516090844},
  },
}

\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{box} = [
  rectangle, minimum width=0.13\textwidth, minimum height=0.13\textwidth,
  text centered, draw=black]
\tikzstyle{empty-box} = [
  rectangle, minimum width=0.13\textwidth, minimum height=0.13\textwidth,
  text width=0.05\textwidth]
\tikzstyle{empty-box-vert} = [
  rectangle, minimum width=0.13\textwidth, minimum height=0.13\textwidth,
  text centered, text height=0.001\textwidth]
\tikzstyle{line} = [draw, -latex']

%% cdb == computed ``delta'' b
\usepackage{ifthen}
\newcommand{\cdb}[1]{
  \ifthenelse{\equal{#1}{1}}
             {\widehat{\partial b}}
             {\widehat{\partial^{#1} b}}
}

\begin{document}
\begin{tikzpicture}[node distance = 0.26\textwidth]
    \node [box] (bj) {\(b\)};
    \node [box, below of=bj] (bn0) {\(\widehat{b}^{(n)}\)};
    \node [box, right of=bn0] (dbn0) {\(\cdb{1}^{(n)}\)};
    \node [box, right of=dbn0] (ddbn0) {\(\cdb{2}^{(n)}\)};
    \node [box, right of=ddbn0] (dddbn0) {\(\cdb{3}^{(n)}\)};
    %%
    \node [box, below of=bn0] (bn1) {\(\widehat{b}^{(n - 1)}\)};
    \node [box, below of=dbn0] (dbn1) {\(\cdb{1}^{(n - 1)}\)};
    \node [box, below of=ddbn0] (ddbn1) {\(\cdb{2}^{(n - 1)}\)};
    \node [box, below of=dddbn0] (dddbn1) {\(\cdb{3}^{(n - 1)}\)};
    %%
    \node [box, below of=bn1] (bn2) {\(\widehat{b}^{(n - 2)}\)};
    \node [box, below of=dbn1] (dbn2) {\(\cdb{1}^{(n - 2)}\)};
    \node [box, below of=ddbn1] (ddbn2) {\(\cdb{2}^{(n - 2)}\)};
    \node [box, below of=dddbn1] (dddbn2) {\(\cdb{3}^{(n - 2)}\)};
    %%
    \path [line] (bn0) -- node[above] {\(\widehat{\ell}_1^{(n)}\)} (dbn0);
    \path [line] (dbn0) -- node[above] {\(\widehat{\ell}_2^{(n)}\)} (ddbn0);
    \path [line] (ddbn0) -- node[above] {\(\widehat{\ell}_3^{(n)}\)} (dddbn0);
    \node [empty-box, right of=dddbn0] (right0) {\(\cdots\)};
    \path [line] (dddbn0) -- node[above] {\(\widehat{\ell}_4^{(n)}\)} (right0);
    %%
    \path [line] (bn1) -- node[above] {\(\widehat{\ell}_1^{(n - 1)}\)} (dbn1);
    \path [line] (dbn1) -- node[above] {\(\widehat{\ell}_2^{(n - 1)}\)} (ddbn1);
    \path [line] (ddbn1) -- node[above] {\(\widehat{\ell}_3^{(n - 1)}\)} (dddbn1);
    \node [empty-box, right of=dddbn1] (right1) {\(\cdots\)};
    \path [line] (dddbn1) -- node[above] {\(\widehat{\ell}_4^{(n - 1)}\)} (right1);
    %%
    \path [line] (bn2) -- node[above] {\(\widehat{\ell}_1^{(n - 2)}\)} (dbn2);
    \path [line] (dbn2) -- node[above] {\(\widehat{\ell}_2^{(n - 2)}\)} (ddbn2);
    \path [line] (ddbn2) -- node[above] {\(\widehat{\ell}_3^{(n - 2)}\)} (dddbn2);
    \node [empty-box, right of=dddbn2] (right2) {\(\cdots\)};
    \path [line] (dddbn2) -- node[above] {\(\widehat{\ell}_4^{(n - 2)}\)} (right2);
    %%
    \path [line] (bj) -- (bn0);
    \path [line] (bn0) -- node[fill=white] {\(U_n(s)\)} (bn1);
    \path [line] (bn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (bn2);
    \node [empty-box-vert, below of=bn2] (bottom0) {\(\vdots\)};
    \path [line] (bn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom0);
    %%
    \path [line] (dbn0) -- node[fill=white] {\(U_n(s)\)} (dbn1);
    \path [line] (dbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dbn2);
    \node [empty-box-vert, below of=dbn2] (bottom1) {\(\vdots\)};
    \path [line] (dbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom1);
    %%
    \path [line] (ddbn0) -- node[fill=white] {\(U_n(s)\)} (ddbn1);
    \path [line] (ddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (ddbn2);
    \node [empty-box-vert, below of=ddbn2] (bottom2) {\(\vdots\)};
    \path [line] (ddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom2);
    %%
    \path [line] (dddbn0) -- node[fill=white] {\(U_n(s)\)} (dddbn1);
    \path [line] (dddbn1) -- node[fill=white] {\(U_{n - 1}(s)\)} (dddbn2);
    \node [empty-box-vert, below of=dddbn2] (bottom3) {\(\vdots\)};
    \path [line] (dddbn2) -- node[fill=white] {\(U_{n - 2}(s)\)} (bottom3);
\end{tikzpicture}
\end{document}

### 2018-05-16 nox.py

@@ -110,6 +110,20 @@ def build_tex(session):
         extensions=("aux", "log", "out"),
     )
 
+    build_tex_file(
+        session,
+        "tikz_local_err",
+        "EF7ADBEFE6118EFEE506836A7AFF7C9E",
+        extensions=("aux", "log", "out"),
+    )
+
+    build_tex_file(
+        session,
+        "tikz_filtration",
+        "52F169AADDA4C4C85C6D5038361816C9",
+        extensions=("aux", "log", "out"),
+    )
+
 
 @nox.session
 def flop_counts(session):

### 2018-05-17 scripts/modify_pdf_id.py
Fixing a bug in `modify_pdf_id.py`.
It was producing PDFs that were "invalid" according to certain
readers, e.g. Preview on OS X.

@@ -22,6 +22,23 @@ import argparse
 
 
 SPLIT_TEXT = b"\n/ID [<"
+ID_LINE = u"{0}> <{0}>]"
+
+
+def verify_id_snippet(id_snippet):
+    """Verifies an ID snippet being replaced.
+
+    ``id_snippet`` is expected to be of the form
+    ``'{X}> <{X}>]'`` where ``X`` is 32 characters.
+    """
+    if len(id_snippet) != 69:
+        raise ValueError(id_snippet, "Invalid length")
+    actual_id = id_snippet[:32].decode("ascii")
+    expected_line = ID_LINE.format(actual_id).encode("ascii")
+    if id_snippet != expected_line:
+        raise ValueError(
+            id_snippet, "Unexpected ID line. Expected", expected_line
+        )
 
 
 def do_replace(path, new_id):
@@ -32,12 +49,12 @@ def do_replace(path, new_id):
     pre, post = contents.split(SPLIT_TEXT)
     # ID line expected to be of the form:
     #     /ID [<...> <...>]
-    _, post = post.split(b"\n", 1)
+    id_snippet, post = post.split(b"\n", 1)
+    verify_id_snippet(id_snippet)
     # which would just leave `<...> <...>]` after the split.
-    new_id_line = u"<{}> <{}>]\n".format(new_id, new_id)
-    new_id_line = new_id_line.encode("ascii")
+    new_id_line = ID_LINE.format(new_id).encode("ascii")
 
-    new_contents = pre + new_id_line + post
+    new_contents = pre + SPLIT_TEXT + new_id_line + b"\n" + post
     with open(path, "wb") as file_obj:
         file_obj.write(new_contents)
 
@@ -45,9 +62,7 @@ def do_replace(path, new_id):
 
 
 def main():
-    description = (
-        "Modify the `/ID` property in a PDF generated by `pdflatex`."
-    )
+    description = "Modify the `/ID` property in a PDF generated by `pdflatex`."
     parser = argparse.ArgumentParser(description=description)
     parser.add_argument(
         "--base", required=True, help="Base path for PDF file to be modified."

### 2018-05-17 scripts/modify_pdf_id.py
Adding an ID check to `modify_pdf_id.py`.

@@ -33,7 +33,11 @@ def verify_id_snippet(id_snippet):
     """
     if len(id_snippet) != 69:
         raise ValueError(id_snippet, "Invalid length")
+
     actual_id = id_snippet[:32].decode("ascii")
+    if not (set(actual_id) <= set("0123456789ABCDEF")):
+        raise ValueError(actual_id, "ID is not composed of hex characters.")
+
     expected_line = ID_LINE.format(actual_id).encode("ascii")
     if id_snippet != expected_line:
         raise ValueError(

### 2018-05-18 doc/cover_letter.tex
Hard-coding May 16 in the two documents that used `\today`.

@@
 {\centering\fromaddress \vspace{0.6in} \\ % Print the header and from address here, add whitespace to move date down
-\hspace*{\longindentation}\today\hspace*{\fill}\par} % Print today's date, remove \today to not display it
+\hspace*{\longindentation}May 16, 2018\hspace*{\fill}\par}  % WAS: \today
 {\raggedright \toname \\ \toaddress \par} % Print the to name and address

### 2018-05-24 doc/paper.tex
Hard-coding May 16 in the two documents that used `\today`.
Fixing error in `Split` algorithm in manuscript.
Updating manuscript.
- Adding more specific details when citing a theorem or result
  from another paper, e.g. (\cite{Foo}, Theorem Bar).
- Updating the `tikz_filtration.pdf` image to have one less "row"
  and one less "column" (it was unnecessarily big)

@@
 \renewcommand*{\appendixname}{}
+%% H/T: https://tex.stackexchange.com/a/23242/32270
+\renewcommand*{\today}{May 16, 2018}
 
@@
       n \mach / (1 - n \mach)\),
-  \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
+  \item \((1 + \theta_k)(1 + \theta_j) = 1 + \theta_{k + j}\),
   \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}
@@
 The error-free transformations used in this paper are
-the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
-algorithm by Dekker \cite{Dekker1971}, respectively.
-
-\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
+the \texttt{TwoSum} algorithm by Knuth (\cite{Knuth1997}) and
+\texttt{TwoProd} algorithm by Dekker (\cite{Dekker1971}, Section 5),
+respectively.
+
+\begin{theorem}[\cite{Ogita2005}, Theorem 3.4]\label{thm:eft}
 For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
@@
 
-\begin{theorem}[\cite{Mainar1999}]
+\begin{theorem}[\cite{Mainar1999}, Corollary 3.2]
 If \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) and \(\mathtt{DeCasteljau}(p, s)\)
-is the value computed by the de Casteljau algorithm then
-\begin{equation}
-\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{2n}
+is the value computed by the de Casteljau algorithm then\footnote{In the
+original paper the factor on \(\widetilde{p}(s)\) is \(\gamma_{2n}\),
+but the authors did not consider round-off when computing
+\(1 \ominus s\).}
+\begin{equation}
+\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{3n}
 \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
@@
 
-\begin{theorem}[\cite{Jiang2010}]
+\begin{theorem}[\cite{Jiang2010}, Theorem 5]
   If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
@@
 \begin{figure}
-  \includegraphics[width=0.9375\textwidth]{tikz_filtration.pdf}
+  \includegraphics[width=0.875\textwidth]{tikz_filtration.pdf}
   \centering
@@
 
-\begin{theorem}[\cite{Ogita2005}]\label{thm:sum-k}
+\begin{theorem}[\cite{Ogita2005}, Proposition 4.10]\label{thm:sum-k}
 A summation can be computed (\texttt{SumK}, Algorithm~\ref{alg:sum-k})
@@
 
-\begin{lemma}[\cite{Jiang2010}]\label{lemma:db-lemma}
-The second order error \(\db{2}^{(0)}_0\) satisfies
+\begin{lemma}[\cite{Jiang2010}, Theorem 4]\label{lemma:db-lemma}
+The second order error \(\db{2}^{(0)}_0\) satisfies\footnote{The authors
+  missed one round-off error so used \(\gamma_{3n + 1}\) where
+  \(\gamma_{3n + 2}\) would have followed from their arguments.}
 \begin{equation}
   \left|\db{1}^{(0)}_0 - \cdb{1}^{(0)}_0\right| =
-  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 1} \gamma_{3(n - 1)}
+  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 2} \gamma_{3(n - 1)}
   \widetilde{p}(s).
@@
 
-\noindent For IEEE-754 double precision floating point number, \(r = 26\)
+\noindent For IEEE-754 double precision floating point number, \(r = 27\)
 so \(2^r + 1\) will be known before \texttt{Split} is called. In all,

### 2018-05-24 scripts/compute_counts.py
Adding a compensated VS-method.
Also modifying the basic implementation for the VS method.

@@ -90,7 +90,7 @@ def count_sum_k():
 
 
 def count_vs_method_basic():
-    print("vs_method.basic() (7n - 2):")
+    print("vs_method.basic() (5n + 1, w/o binomial):")
     for degree in range(1, 5 + 1):
         parent = operation_count.Computation()
         x = operation_count.Float(0.25, parent)
@@ -100,7 +100,22 @@ def count_vs_method_basic():
         )
         p = vs_method.basic(x, coeffs)
         assert p.value == 0.5 ** degree
-        assert parent.count == 7 * degree - 2
+        assert parent.count == 5 * degree + 1
+        print("  degree {}:     {}".format(degree, parent.display))
+
+
+def count_vs_method_compensated():
+    print("vs_method.compensated() (26n + 7, w/o binomial):")
+    for degree in range(1, 5 + 1):
+        parent = operation_count.Computation()
+        x = operation_count.Float(0.25, parent)
+        coeffs = tuple(
+            operation_count.Float((-1.0) ** k, parent)
+            for k in range(degree + 1)
+        )
+        p = vs_method.compensated(x, coeffs)
+        assert p.value == 0.5 ** degree
+        assert parent.count == 26 * degree + 7
         print("  degree {}:     {}".format(degree, parent.display))
 
 
@@ -355,6 +370,8 @@ def main():
     count_sum_k()
     print(SEPARATOR)
     count_vs_method_basic()
+    print(SEPARATOR)
+    count_vs_method_compensated()
     print(SEPARATOR)
     count_horner_basic()
     print(SEPARATOR)

### 2018-05-24 src/eft.py

@@ -54,10 +54,8 @@ def multiply_eft(val1, val2, use_fma=True):
     else:
         high1, low1 = _split(val1)
         high2, low2 = _split(val2)
-        error = (
-            low1
-            * low2
-            - (((product - high1 * high2) - low1 * high2) - high1 * low2)
+        error = low1 * low2 - (
+            ((product - high1 * high2) - low1 * high2) - high1 * low2
         )
 
     return product, error

### 2018-05-24 src/operation_count.py

@@ -151,25 +151,3 @@ class Float(object):
         frac3 = fractions.Fraction(float3)
         result = float(frac1 * frac2 + frac3)
         return Float(result, self.computation)
-
-
-def as_type(value, typed_value):
-    """Converts a basic float to a type.
-
-    Args:
-        value (float): The value to convert to a type.
-        typed_value (object): The type to match (and the value
-            may contain come metadata).
-
-    Returns:
-        object: A wrapper of ``value`` with the same type as ``typed_value``.
-
-    Raises:
-        TypeError: If ``typed_value`` is not a ``float`` or ``Float``.
-    """
-    if isinstance(typed_value, float):
-        return value
-    elif isinstance(typed_value, Float):
-        return Float(value, typed_value.computation)
-    else:
-        raise TypeError("Unexpected type", typed_value)

### 2018-05-24 src/vs_method.py

@@ -19,24 +19,59 @@ This method has been described in the 1D case in a `survey`_
 paper that followed the original formulation.
 """
 
+import fractions
+import math
 
-import operation_count
+import eft
+
+
+def binomial(n, k):
+    numerator = math.factorial(n)
+    denominator = math.factorial(k) * math.factorial(n - k)
+    result = fractions.Fraction(numerator, denominator)
+    if float(result) != result:
+        raise ValueError(n, k)
+    return float(result)
 
 
 def basic(s, coeffs):
-    degree = len(coeffs) - 1
+    n = len(coeffs) - 1
     r = 1.0 - s
 
-    result = r * coeffs[0]
+    result = coeffs[0]
 
-    binom_val = operation_count.as_type(1.0, s)
-    s_pow = operation_count.as_type(1.0, s)
-    for j in range(1, degree):
+    s_pow = 1.0
+    for j in range(1, n + 1):
         s_pow = s * s_pow
-        binom_val = binom_val * (degree - j + 1)
-        binom_val = binom_val / j
-        result = result + binom_val * s_pow * coeffs[j]
-        result = result * r
+        binom_val = binomial(n, j)
+        result = r * result + binom_val * s_pow * coeffs[j]
 
-    result = result + s * s_pow * coeffs[degree]
     return result
+
+
+def compensated(s, coeffs):
+    n = len(coeffs) - 1
+    r, rho = eft.add_eft(1.0, -s)
+
+    pk = coeffs[0]
+    dpk = 0.0
+
+    s_pow = 1.0
+    ds = 0.0
+    for j in range(1, n + 1):
+        # Update ``s^k``, using
+        #   s^k = s_pow + ds ==> s^{k + 1} = s(s_pow) + s(ds)
+        s_pow, pi_s = eft.multiply_eft(s, s_pow)
+        ds = s * ds + pi_s
+        # Now, update ``pk`` and ``dpk``.
+        P1, pi1 = eft.multiply_eft(r, pk)
+        local_err = pi1 + rho * pk
+        P2, pi2 = eft.multiply_eft(coeffs[j], binomial(n, j))
+        local_err += P2 * ds + pi2 * s_pow
+        P3, pi3 = eft.multiply_eft(P2, s_pow)
+        local_err += pi3
+        pk, sigma4 = eft.add_eft(P1, P3)
+        local_err += sigma4
+        dpk = r * dpk + local_err
+
+    return pk + dpk

### 2018-08-28 noxfile.py
Adding "optimal" flags for C/C++ libraries.
Note that ``-ffast-math`` is a very bad flag to use here since
it will treat arithmetic as transitive and modify ``two_sum`` to
always have ``error == 0``.

Also using include guards on C++ header files.

@@ -183,6 +183,8 @@ def verify_cpp(session):
     session.run(
         "g++",
         "-std=c++11",
+        "-O3",
+        "-march=native",
         "-o",
         "main",
         os.path.join("scripts", "tests.cpp"),
@@ -217,6 +219,8 @@ def verify_c(session):
     session.run(
         "gcc",
         "-std=c99",
+        "-O3",
+        "-march=native",
         "-o",
         "main",
         os.path.join("scripts", "tests.c"),

### 2018-08-28 src/de_casteljau.hpp

@@ -10,6 +10,9 @@
 // See the License for the specific language governing permissions and
 // limitations under the License.
 
+#ifndef DE_CASTELJAU_HPP
+#define DE_CASTELJAU_HPP
+
 #include "eft.hpp"
 #include <array>
 #include <tuple>
@@ -90,3 +93,5 @@ std::array<double, K> compensated(
     return b_hat;
 }
 }
+
+#endif /* DE_CASTELJAU_HPP */

### 2018-08-28 src/eft.hpp

@@ -10,6 +10,9 @@
 // See the License for the specific language governing permissions and
 // limitations under the License.
 
+#ifndef EFT_HPP
+#define EFT_HPP
+
 #include <utility>
 #include <vector>
 
@@ -18,3 +21,5 @@ std::pair<double, double> two_prod(double val1, double val2);
 std::pair<double, double> two_sum(double val1, double val2);
 double sum_k(const std::vector<double>& vec, size_t K);
 }
+
+#endif /* EFT_HPP */

### 2018-08-29 doc/paper.tex
Adding "Acknowledgements" section to paper.
This includes a reference to the GitHub repository so that people
may access the source code. In addition:

- Updated `README.md` to give specific pointers to where the
  algorithm is implemented in C, C++ and Python
- Updated the Knuth reference to the 1st edition to make it clear
  that the algorithm was stated in 1969, not 1997
- Added Babuska reference into the "Acknowledgements" section since
  a similar compensated algorithm was described there
- Added a missing tilde over `\ell` in the definition of `L_{F, k}`

@@
 The error-free transformations used in this paper are
-the \texttt{TwoSum} algorithm by Knuth (\cite{Knuth1997}) and
+the \texttt{TwoSum} algorithm by Knuth (\cite{Knuth1969}) and
 \texttt{TwoProd} algorithm by Dekker (\cite{Dekker1971}, Section 5),
@@
 As we'll see soon (Lemma~\ref{lemma:k-order}), putting a bound on
-sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
+sums of the form \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)\) will
 be useful to get an overall bound on the relative error for
 \texttt{CompDeCasteljauK}, so we define
-\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\).
+\(L_{F, k} \coloneqq \sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)\).
 
@@
 
+\section{Acknowledgements}
+
+The author would like to thank \cite{Ogita2005}, \cite{Graillat2009} and
+\cite{Jiang2010} for their papers that motivated this work. In particular,
+the work of \cite{Ogita2005} reignited the path to compensated algorithms set
+forth in \cite{Babuska1968}, \cite{Knuth1969} and \cite{Dekker1971}.
+
+The family of algorithms described in this paper have been implemented in
+C, C++ and Python by the author (\cite{KCompensatedGitHub}).
+
 %% H/T: https://tex.stackexchange.com/a/137379/32270

### 2019-03-12 scripts/error_against_cond.py
Reducing y-limit to 1.0 in `jlcs10_plot.pdf`.
This was recommended in the 2nd round of review feedback.

@@ -364,13 +364,13 @@ def main_jlcs10(filename=None):
     )
 
     # Make sure the ``y``-limit stays set (the bounds lines exceed).
-    ax.set_ylim(min_y, 10.0 ** 18)
+    ax.set_ylim(min_y, 10.0)
     ax.set_xlim(min_x, max_x)
     # Add the legend.
     ax.legend(loc="lower right", framealpha=1.0, frameon=True)
     # Set "nice" ticks.
     ax.set_xticks([10.0 ** n for n in range(5, 45 + 5, 5)])
-    ax.set_yticks([10.0 ** n for n in range(-18, 14 + 4, 4)])
+    ax.set_yticks([10.0 ** n for n in range(-20, 0 + 4, 4)])
     # Set special ``xticks`` for ``1/u`` and ``1/u^2``.
     ax.set_xticks([1.0 / float(U), 1.0 / float(U) ** 2], minor=True)
     ax.set_xticklabels([r"$1/\mathbf{u}$", r"$1/\mathbf{u}^2$"], minor=True)

### 2019-04-06 doc/paper.tex
Adding a future work section per reviewer feedback.
Noted the existing C/C++/Python implementations and mention a more
comprehensive library that could include them as well as referenced
a related library / paper.
Adding note about published version to `paper.pdf`.

@@
 
+(This has been published as \cite{Hermes2019}.)
 In computer aided geometric design, polynomials are usually expressed in
@@
 
+\section{Future Work}
+
+The family of algorithms described in this paper have been implemented in
+C, C++ and Python by the author (\cite{KCompensatedGitHub}). A more complete
+compensated algorithms library (similar to \cite{Barrio2018}) could be quite
+useful. For example, such a library could include the algorithms in the
+existing literature such as the \(K\)-compensated algorithm for Horner's method
+from \cite{Graillat2009}.
+
 \section{Acknowledgements}
@@
 forth in \cite{Babuska1968}, \cite{Knuth1969} and \cite{Dekker1971}.
-
-The family of algorithms described in this paper have been implemented in
-C, C++ and Python by the author (\cite{KCompensatedGitHub}).
 

### 2019-04-06 doc/paper.tex
Updating `paper.tex` to put "published in" in footnote on 1st page.
This required modifying `\ps@pprintTitle` from the original `elsarticle.cls`
source. This was located at `/usr/share/texlive/texmf-dist/tex/latex/elsarticle/elsarticle.cls`
on my local machine.

Also making sure `de Casteljau` and `$K$` are correctly capitalized
in the bibliography.

@@
 %% H/T: https://tex.stackexchange.com/a/23242/32270
-\renewcommand*{\today}{May 16, 2018}
+%% Initial Submission: May 16, 2018
+%% First Revision: August 29, 2018
+%% Second Revision: March 11, 2019
+%% Accepted: March 19, 2019
+%% Published: April 5, 2019
 
@@
 
-\journal{Applied Mathematics and Computation}
+%% \journal{Applied Mathematics and Computation}
+\makeatletter
+\def\ps@pprintTitle{%
+  \let\@oddhead\@empty
+  \let\@evenhead\@empty
+  \def\@oddfoot{\footnotesize\itshape
+    Published in Applied Mathematics and Computation (\cite{Hermes2019})
+      \hfill April 5, 2019}%
+  \let\@evenfoot\@oddfoot}
+\makeatother
 
@@
 
-(This has been published as \cite{Hermes2019}.)
 In computer aided geometric design, polynomials are usually expressed in
